---
title: "1_loading_and_preparing_data_from_excel"
author: "Kamau Lindhardt, lbk125"
date: "2022-09-11"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/Library/Mobile Documents/com~apple~CloudDocs/Temperate_Agroforestry_PAPER/META_ANALYSIS/R/silvoarable_review")
```

# Aim & Research questions

The **aim** of this systematic mapping review is to thoroughly catalogue
and assess the current research themes of knowledge abundance and
scarcity and provide the first global systematic synthesis of ES
research in relation to temperate SAF. We specifically addressed the
following four **research questions**:

1.  What temperate SAF systems and ecosystem services have been studied?

2.  What tree and crop species composition and associations are used in
    temperate SAF systems?

3.  How are temperate SAF systems and ecosystem services distributed
    across geographic and pedoclimatic gradients?

4.  To what extend does pedoclimatic properties, such as soil quality,
    temperature and precipitation drive inter-site differences? 

Descriptive statistics, e.g. frequency of occurrences is used to
visualise data

# Preparing R script

Packages to be installed first

```{r Installing the pacman package, warning=FALSE}
# install.packages("pacman")
# install.packages("readxl")
# install.packages("pacman")
# install.packages("tidygeocoder")
# install.packages("ggdensity")
# install.packages("rnaturalearth")
# install.packages("rnaturalearthdata")
# install.packages("gridExtra")
# install.packages("soilDB")
# install.packages("aqp")
# install.packages("geodata")
```

## Loading packages and libraries

```{r Loading other needed packages, warning=FALSE}
suppressWarnings({

  # Applying the p_load function to load multiple add-on packages in one line of code.

  pacman::p_load(
    tidyverse, # general package for everything
    readxl,
    vroom, # for crazy fast method of loading/reading datasets from local disc
    janitor, # for cleaning data column names
    #styler, # for neatly organising code
    tidygeocoder, # for making geospatial analysis (e.g. derive coordinates from location names)
    rnaturalearth, # facilitates world mapping by making Natural Earth map data more easily available
    rnaturalearthdata, # holds data used by rnaturalearth package from Natural Earth.

    raster, # for general spatial analysis and in particular for downloading BioClim variables from WorldClim
    # geodata     # use geodata::worldclim_global() instead of raster::getData() if encounting issues
    sp,
    sf,
    soilDB, # for downloading predicted soil properties from ISRIC SoilGrids
    aqp
  )
})
```

## Update GitHub credentials for efficient 'push' and 'pull' operations

```{r Update GitHub credentials for efficient 'push' and 'pull' operations}
# Run this function in order to update GitHub credentials for efficient 'push' and 'pull' operations

# gitcreds_set()
```

## Loading main database

```{r Loading database 2.0, warning=FALSE, message=FALSE}
# Get the current working directory
place_of_database <- "~/Library/Mobile Documents/com~apple~CloudDocs/Temperate_Agroforestry_PAPER/META_ANALYSIS/R/silvoarable_review/DATA/DATABASE_2.0.xlsx"

suppressWarnings({
  database <- readxl::read_excel(path = place_of_database,
    sheet = "MAIN_DATABASE_2.0"
  ) %>%
    filter(INCLUDED == TRUE) # include only data entries that have been assesed and deemed included
})
```

**Glimpse (taking a look at the data)**
```{r Glimpse the dataset, eval=FALSE}
database %>% dplyr::glimpse() 
```

```{r}
database %>% 
  dplyr::select(ID.P, ID.S,
                STUDY.LOC.COUNTRY,
                TITLE,
                ST.TYPE) %>%
  dplyr::filter(STUDY.LOC.COUNTRY == "Netherlands")
```

**Number of independent studies and distinct sites**
Defining number of independent studies and number of observations based on individual study sites
```{r Number of studies and study sites}
n_independent_studies <- database %>%
  distinct(ID.P) %>%
  nrow()

# n_independent_studies = 73 (16 aug. 2022)

n_individual_study_sites <- database %>%
  distinct(ID.P, ID.S) %>%
  nrow()

# n_individual_study_sites = 213 (16 aug. 2022)
```

# Structure of this script is as follows:

1.  Single variate data dimensions are visualised (everything that is not matrices)

2.  Multivariate data dimensions are visualised (mostly matrices)

3.  Geospatially-related data dimensions (pedoclimatic data) are visualised

# Single datasets (not matrices)

## Temporal scale of studies

Creating the dataset for assessing the **temporal scale of studies** using Pie-Donut plot

```{r}
temp.scale <- database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::select(
    ID.P, ID.S,
    DUR.SAMPLING
  ) %>%
  # na.omit() %>%
  # filter(!DUR.SAMPLING == "NA")
  mutate_if(is.character, as.factor) %>%
  group_by(DUR.SAMPLING) %>%
  summarise(count = n()) %>%
  mutate(pct = count / sum(count)) %>%
  mutate(across(where(is.numeric), round, 3)) %>%
  # rounding no of decimals
  # filter(DUR.SAMPLING != "NA") %>%                           # removing one entry that has NA
  mutate(Tier_2_scale = DUR.SAMPLING) %>%
  mutate(Tier_2_scale = recode(Tier_2_scale, # rename dataframe rows
    `< 2 years` = "< 2 years",
    `2 - 4 years` = "2 - 4 years",
    `4 - 10 years` = "4 - 10 years",
    `> 10 years` = "> 10 years",
    .keep = "all"
  )) %>%
  mutate(Tier_1_scale = recode(Tier_2_scale, # rename dataframe rows
    `< 2 years` = "Short",
    `2 - 4 years` = "Medium",
    `4 - 10 years` = "Long",
    `> 10 years` = "Very long",
    .keep = "all"
  )) %>%
  relocate(Tier_1_scale, Tier_2_scale)

# ------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(temp.scale, "./DATASET.FROM.SCRIPT/temp.scale.csv")
# -----------------------------------------------------------------------------------------------
glimpse(temp.scale)
```

## Spatial scale of studies

Creating the dataset for assessing the **spatial scale of studies** using Pie-Donut plot

```{r}
spatial.scale <- database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::select(
    ID.P, ID.S,
    SCALE
  ) %>%
  # na.omit() %>%
  # filter(!SCALE == "NA")
  mutate_if(is.character, as.factor) %>%
  group_by(SCALE) %>%
  summarise(count = n()) %>%
  mutate(pct = count / sum(count)) %>%
  mutate(across(where(is.numeric), round, 3)) %>%
  mutate(Tier_2_scale = SCALE) %>%
  dplyr::select(!SCALE) %>%
  mutate(Tier_2_scale = recode(Tier_2_scale, # rename dataframe rows
    PLOT = "Plot",
    PLOT.FARM = "Plot and farm",
    FARM = "Farm",
    LAND = "Landscape",
    PLOT.REGI = "Plot and regional",
    REGI = "Regional",
    NATI = "National",
    INTE = "International",
    .keep = "all"
  )) %>%
  mutate(Tier_1_scale = recode(Tier_2_scale, # rename dataframe rows
    Plot = "Small scale",
    `Plot and farm` = "Small scale",
    Farm = "Small scale",
    Landscape = "Medium scale",
    `Plot and regional` = "Medium scale",
    Regional = "Medium scale",
    National = "Large scale",
    International = "Large scale",
    .keep = "all"
  )) %>%
  relocate(Tier_1_scale, Tier_2_scale)

# -----------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(spatial.scale, "./DATASET.FROM.SCRIPT/spatial.scale.csv")
# -----------------------------------------------------------------------------------------------
glimpse(spatial.scale)
```

## Ecosystem services

Creating **ecosystem services** data by renaming the variables (indicators)

```{r Ecosystem service}

ecosystem <-
  database %>%
  filter(INCLUDED == TRUE) %>%
  # selecting Ecosystem services columns
  dplyr::select(
    ID.P, ID.S, # ST.TYPE, MANAGEMENT
    starts_with("ES")
  ) %>%
  mutate(across(matches("ES"), as.logical)) %>%
  mutate_at(vars(-c(ID.P, ID.S)), tidyr::replace_na, 0) %>%
  # <- Including NA's as 0 (FALSE)
  mutate_if(is.logical, as.numeric) %>%
  mutate_if(is.character, as.numeric) %>%
  dplyr::select(-c(ID.S)) %>%
  distinct(ID.P, across(contains("ES"))) # including only independent studies and not separate study sites

# ---------------------------------------------------------------------------------------------------
# total number of peer-reviewed papers reporting on ecosystem service-related indicators


# sum the data using apply
ecosystem.sum <-
  data.frame(value = apply(ecosystem, MARGIN = 2, FUN = sum))

ecosystem.sum$key <- rownames(ecosystem.sum)


total_n_ES <- ecosystem.sum %>%
  filter(key == "ES")

# Saving the saving dataset
readr::write_csv(total_n_ES, "./DATASET.FROM.SCRIPT/total_n_ES.csv")
# ---------------------------------------------------------------------------------------------------

ecosystem.data <- ecosystem %>%
  rename(
    "Food (yield)" = ES.SUB.PROV.FOOD,
    "Feed" = ES.SUB.PROV.FEED,
    "Fibre" = ES.SUB.PROV.FIBRE,
    "Fuel (firewood etc.)" = ES.SUB.PROV.FUEL,
    "Genetic resources" = ES.SUB.PROV.GENRES,
    "Biochemicals" = ES.SUB.PROV.BIOCHEM,
    "Ornamentals" = ES.SUB.PROV.ORNA,
    "Fresh water" = ES.SUB.PROV.WATFRESH,
    "Air regulation" = ES.SUB.REGU.AIR,
    "Climate regulation" = ES.SUB.REGU.CLIMREG,
    "Carbon sequestration" = ES.SUB.REGU.CSEQREG,
    "Water regulation" = ES.SUB.REGU.WATREG,
    "Nutrient leaching regulation" = ES.SUB.REGU.NUTRILE,
    "Erosion regulation" = ES.SUB.REGU.EROREG,
    "Water purification" = ES.SUB.REGU.WATPUR,
    "Disease regulation" = ES.SUB.REGU.DISREG,
    "Pest regulation" = ES.SUB.REGU.PESTREG,
    "Polination regulation" = ES.SUB.REGU.POLREG,
    "Natural hazards regulation" = ES.SUB.REGU.NATHAZREG,
    "Soil formation" = ES.SUB.SUPP.SOILFORM,
    "Photosynthesis" = ES.SUB.SUPP.PHOTSYNT,
    "Primary production" = ES.SUB.SUPP.PRIMPROD,
    "Nutrient cycling" = ES.SUB.SUPP.NUTRICYC,
    "Water cycling" = ES.SUB.SUPP.WATCYC,
    "Cultural diversity" = ES.SUB.CULT.CULTDIV,
    "Spiritual value" = ES.SUB.CULT.SPIRITVAL,
    "Knowledge systems" = ES.SUB.CULT.KNOWLSYS,
    "Educational value" = ES.SUB.CULT.EDUVAL,
    "Inspirational value" = ES.SUB.CULT.INSPIR,
    "Aesthetic value" = ES.SUB.CULT.AESTVAL,
    "Social relations" = ES.SUB.CULT.SOCREL,
    "Sense of place" = ES.SUB.CULT.SENPLACE,
    "Cultural heritage value" = ES.SUB.CULT.CULTHERIVAL,
    "Recreational and ecotourism" = ES.SUB.CULT.RECRECOTOUR,
    "Provisioning ES" = ES.PROV,
    "Regulating ES" = ES.REGU,
    "Supporting ES" = ES.SUPP,
    "Cultural ES" = ES.CULT,
    "Ecosystem Services" = ES
  )
# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(ecosystem.data, "./DATASET.FROM.SCRIPT/ecosystem.data.csv")
# -----------------------------------------------------------------------------------------------
glimpse(ecosystem.data)
```

## Approach used by studies to assess ecosystem services

Creating the dataset; **approach used by studies for ES assessment**
using Pie-Donut plot

```{r}
study.approach <- database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::select(
    ID.P,
    STUDY.FRAME
  ) %>%
  mutate_at(vars(-c(STUDY.FRAME)), as.numeric)

ecosystem.data.DP <-
  ecosystem.data %>%
  dplyr::select(
    ID.P,
    `Provisioning ES`,
    `Regulating ES`,
    `Supporting ES`,
    `Cultural ES`
  ) %>%
  pivot_longer(
    cols = ends_with("ES"),
    names_to = "ES",
    values_to = "count"
  )

# --------------------------------------------------------------------------------------------------- Merging the two datasets

study.approach.ecosystem <- left_join(
  x = study.approach, y = ecosystem.data.DP,
  by = c("ID.P" = "ID.P")
) %>%
  relocate(ID.P) %>%
  mutate_if(is.character, as.factor) %>%
  dplyr::select(!ID.P)

# ---------------------------------------------------------------------------------------------------

# Building a table with the data for the plot
ES.study.approach.PD.data <- study.approach.ecosystem %>%
  group_by(STUDY.FRAME, ES) %>%
  rename("Approach" = STUDY.FRAME) %>%
  rename("Ecosystem_Service" = ES) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  mutate(Type = "Approach") %>%
  mutate(Approach = recode(Approach, # rename dataframe rows
    BIO.PHYS = "pedoclimatic",
    MIXED = "Mixed approach",
    MONETARY = "Monetary",
    SOCIO.CUL = "Socio-cultural"
  ))

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(ES.study.approach.PD.data, "./DATASET.FROM.SCRIPT/ES.study.approach.PD.data.csv")
# -----------------------------------------------------------------------------------------------
glimpse(ES.study.approach.PD.data)
```


## Experimental sites of SAF - countries

```{r}
sites.countries.data <- database %>%
  filter(INCLUDED == TRUE) %>%
  
  dplyr::select(#ID.P, ID.S,
                STUDY.LOC.COUNTRY) %>%
  distinct()
```



## SAF System Trait Types

Creating the dataset for **SAF system trait types**

```{r SAF system trait types, tidy=TRUE}
system.trait.data <- database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::mutate(ST.TYPE = case_when(
    str_detect(ST.TYPE, "RBS") ~ "Riparian buffer strips",
    str_detect(ST.TYPE, "SHB") ~ "Line belts *",
    str_detect(ST.TYPE, "MIXED") ~ "Mixed systems",
    str_detect(ST.TYPE, "TIMBER.ARABLE") ~ "Timber - Arable",
    str_detect(ST.TYPE, "BIOMASS.ARABLE") ~ "Biomass - Arable",
    str_detect(ST.TYPE, "NUT.ARABLE") ~ "Nut - Arable",
    str_detect(ST.TYPE, "FRUIT.ARABLE") ~ "Fruit - Arable",
    TRUE ~ as.character(as.character(ST.TYPE))
  )) %>%
  dplyr::mutate(ST.TYPE.PROD.TYPE = "x") %>%
  dplyr::mutate(ST.TYPE.PROD.TYPE == case_when(
    str_detect(ST.TYPE, "Timber - Arable") ~ "Biomass",
    str_detect(ST.TYPE, "Biomass - Arable") ~ "Biomass",
    str_detect(ST.TYPE, "Nut - Arable") ~ "Food",
    str_detect(ST.TYPE, "Fruit - Arable") ~ "Food",
    str_detect(ST.TYPE, "Riparian buffer strips") ~ "Multiple",
    str_detect(ST.TYPE, "Line belts *") ~ "Multiple",
    str_detect(ST.TYPE, "Mixed systems") ~ "Multiple",
    TRUE ~ as.character(as.character(ST.TYPE.PROD.TYPE))
  )) %>%
  group_by(ST.TYPE) %>%
  summarise(count = n()) %>%
  mutate(pct = (count / sum(count))) %>%
  arrange(desc(-count)) %>%
  # summarise(count = sum(count)) %>%
  # filter(ST.TYPE != "NA")
  # Order factor levels by number, putting "Other" to end
  # Order factor levels by number, putting "NA" to next to end
  dplyr::mutate(
    ST.TYPE = forcats::fct_rev(forcats::fct_inorder(ST.TYPE)),
    ST.TYPE = forcats::fct_relevel(ST.TYPE, "NA", after = 8)
  )
#  dplyr::mutate(ST.TYPE = forcats::fct_rev(forcats::fct_inorder(ST.TYPE)),
#              ST.TYPE = forcats::fct_relevel(ST.TYPE, "Mixed and other systems", after = 4))

# Checking that the reordering worked
# levels(system.trait.data$ST.TYPE)

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(system.trait.data, "./DATASET.FROM.SCRIPT/system.trait.data.csv")
# ---------------------------------------------------------------------------------------------------------------------------------
glimpse(system.trait.data)
```

## Functional role (product type) of trees

Creating the dataset for the **functional role (product type) of tree
component** of SAF system trait types

```{r Functional role (product type) of SAF system trait types}

system.trait.products <- database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::mutate(ST.TYPE = case_when(
    str_detect(ST.TYPE, "RBS") ~ "Riparian buffer strips",
    str_detect(ST.TYPE, "SHB") ~ "Line belts *",
    str_detect(ST.TYPE, "MIXED") ~ "Mixed systems",
    str_detect(ST.TYPE, "TIMBER.ARABLE") ~ "Timber - Arable",
    str_detect(ST.TYPE, "BIOMASS.ARABLE") ~ "Biomass - Arable",
    str_detect(ST.TYPE, "NUT.ARABLE") ~ "Nut - Arable",
    str_detect(ST.TYPE, "FRUIT.ARABLE") ~ "Fruit - Arable",
    TRUE ~ as.character(as.character(ST.TYPE))
  )) %>%
  mutate(ST.TYPE.PROD.TYPE = case_when(
    ST.TYPE == "Timber - Arable" ~ "Biomass",
    ST.TYPE == "Biomass - Arable" ~ "Biomass",
    ST.TYPE == "Riparian buffer strips" ~ "Biomass",
    ST.TYPE == "Nut - Arable" ~ "Food",
    ST.TYPE == "Fruit - Arable" ~ "Food",
    ST.TYPE == "Line belts *" ~ "Nature + wind shelter",
    ST.TYPE == "Mixed systems" ~ "Mixed"
  )) %>%
  dplyr::group_by(ST.TYPE.PROD.TYPE) %>%
  summarise(count = n()) %>%
  add_tally(wt = count) %>%
  mutate(pct = (count / n)) %>%
  arrange(desc(-count)) %>%
  filter(ST.TYPE.PROD.TYPE != "NA")

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(system.trait.products, "./DATASET.FROM.SCRIPT/system.trait.products.csv")
# -----------------------------------------------------------------------------------------------

glimpse(system.trait.products)
```

## Spatial arrangements (interrow distance)

Creating the dataset for **interrow distance**

```{r Interrow spatial arrangements, tidy=TRUE}

spatial.arrange.data <- database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::select(
    ID.P, ID.S,
    CA.DIST.INTERROW, CA.DIST.INTERROW.MIN, CA.DIST.INTERROW.MAX,
    CA.DIST.INTRAROW, ST.TYPE
  ) %>%
  mutate_at(vars(-c(ST.TYPE)), as.numeric) %>%
  # <- converting characters to numeric
  # where there is no information in the single value interrow distance, then use the mean value of interrow min and interrow max
  rowwise() %>%
  mutate(avr_interrow = mean(c(CA.DIST.INTERROW.MIN, CA.DIST.INTERROW.MAX), na.rm = FALSE))

# mutate(CA.DIST.INTERROW = ifelse(CA.DIST.INTERROW == "NA", avr_interrow, CA.DIST.INTERROW), .keep = "all")

# Using base R commands
# Duplicate data
spatial.arrange.data_new <- spatial.arrange.data
# Replace NA values
spatial.arrange.data_new$CA.DIST.INTERROW[is.na(spatial.arrange.data_new$CA.DIST.INTERROW)] <-
  spatial.arrange.data_new$avr_interrow[is.na(spatial.arrange.data_new$CA.DIST.INTERROW)]

# Renaming to original
spatial.arrange.data <- spatial.arrange.data_new

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(spatial.arrange.data, "./DATASET.FROM.SCRIPT/spatial.arrange.data.csv")
# -----------------------------------------------------------------------------------------------
glimpse(spatial.arrange.data)
```

## Crop species groups

Creating the dataset for **crop species groups**

```{r Crop groups, message=FALSE, tidy=TRUE}
# using n_individual_study_sites because crop species vary between sites, as some studies include multiple sites with different species compositions and arrangements.

crops.data <-
  read_excel("DATA/DATABASE_2.0.xlsx",
    sheet = "CROP_SPECIES"
  ) %>%
  # Generating broad categories of crop species - based on broad output product
  # based on FAO, 2019. FAOSTAT database. http://www.fao.org/faostat/en/#data.
  mutate(CROP.OUTPUT.GROUP = case_when(
    str_detect(CROP.SPECIES, "Triticum") ~ "Cereals",
    str_detect(CROP.SPECIES, "Zea") ~ "Cereals",
    str_detect(CROP.SPECIES, "Hordeum") ~ "Cereals",
    str_detect(CROP.SPECIES, "Secale") ~ "Cereals",
    str_detect(CROP.SPECIES, "Avena") ~ "Cereals",
    str_detect(CROP.SPECIES, "Sorghum") ~ "Cereals",
    str_detect(CROP.SPECIES, "Fagopyrum") ~ "Cereals",
    str_detect(CROP.SPECIES, "Glycine") ~ "Pulses",
    str_detect(CROP.SPECIES, "Pisum") ~ "Pulses",
    str_detect(CROP.SPECIES, "Cicer") ~ "Pulses",
    str_detect(CROP.SPECIES, "Lupinus") ~ "Pulses",
    str_detect(CROP.SPECIES, "Beta") ~ "Sugar crops",
    str_detect(CROP.SPECIES, "Cucurbita") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "UNSPECIFIED.VEGETABLES") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "Asparagus") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "Brassica oleracea") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "Capsicum") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "Citrullus") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "Cucumis") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "Lactuca") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "Solanum lycopersicum") ~ "Vegetable crops",
    str_detect(CROP.SPECIES, "Brassica napus subsp. napus") ~ "Oil crops",
    str_detect(CROP.SPECIES, "Helianthus") ~ "Oil crops",
    str_detect(CROP.SPECIES, "Sinapsis") ~ "Oil crops",
    str_detect(CROP.SPECIES, "Brassica alba") ~ "Oil crops",
    str_detect(CROP.SPECIES, "Ribes") ~ "Fruit crops",
    str_detect(CROP.SPECIES, "Rubus") ~ "Fruit crops",
    str_detect(CROP.SPECIES, "Aronia") ~ "Fruit crops",
    str_detect(CROP.SPECIES, "Fragaria") ~ "Fruit crops",
    str_detect(CROP.SPECIES, "Solanum tuberosum") ~ "Roots and tuber crops",
    str_detect(CROP.SPECIES, "HERBS") ~ "OTHER CROPS FOR HUMAN CONS.",
    str_detect(CROP.SPECIES, "	SP.BEEF.CATTLE") ~ "OTHER CROPS FOR HUMAN CONS.",
    str_detect(CROP.SPECIES, "	SP.DAIRY.CATTLE") ~ "OTHER CROPS FOR HUMAN CONS.",
    str_detect(CROP.SPECIES, "	SP.PIG") ~ "OTHER CROPS FOR HUMAN CONS.",
    str_detect(CROP.SPECIES, "Lolium") ~ "Forage crops",
    str_detect(CROP.SPECIES, "Vicia") ~ "Forage crops",
    str_detect(CROP.SPECIES, "Poa") ~ "Forage crops",
    str_detect(CROP.SPECIES, "Festuca") ~ "Forage crops",
    str_detect(CROP.SPECIES, "Medicago") ~ "Forage crops",
    str_detect(CROP.SPECIES, "Panicum") ~ "Forage crops",
    str_detect(CROP.SPECIES, "Trifolium") ~ "Forage crops",
    str_detect(CROP.SPECIES, "HAY.FODDER") ~ "Forage crops",
    str_detect(CROP.SPECIES, "UNSPECIFIED.CLOVERLAY") ~ "Forage crops",
    str_detect(CROP.SPECIES, "UNSPECIFIED.Fallow") ~ "Fallow"
  )) %>%
  # removing 'unreal' species
  filter(!grepl("HAY|UNSPECIFIED|HERBS|SP|NA|NS", CROP.SPECIES)) %>%
  # Generating broad categories of crop species - based on genus
  separate(CROP.SPECIES, c("CROP.GENUS.GROUP", "CROP.SPECIES.GROUP"), "^\\S*\\K\\s+", remove = FALSE)

crops.data.aggr <- crops.data %>%
  filter(CROP.OUTPUT.GROUP != "NA") %>%
  # excluding 41 records of 'NA'
  group_by(CROP.OUTPUT.GROUP) %>%
  summarise(count = n()) %>%
  add_tally(wt = count) %>%
  mutate(pct = (count / n) * 100) %>%
  mutate(across(where(is.numeric), round, 1)) %>%
  # rounding no of decimals
  arrange(desc(-count))

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(crops.data, "./DATASET.FROM.SCRIPT/crops.data.csv")
# -----------------------------------------------------------------------------------------------
glimpse(crops.data)
```

## Tree species genera

Creating the dataset for **tree species and tree genera groups**

```{r Tree genera groups, message=FALSE, warning=FALSE, tidy=TRUE}
# using n_individual_study_sites because tree species vary between sites, as some studies include multiple sites with different species compositions and arrangements.

trees.data <-
  read_excel("DATA/DATABASE_2.0.xlsx",
    sheet = "TREE_SPECIES"
  ) %>%
  # Generating broad categories of tree species - based on families, genus and hybrids
  separate(TREE.SPECIES, c("TREE.GENUS.GROUP", "TREE.SPECIES.GROUP"), "^\\S*\\K\\s+", remove = FALSE) %>%
  separate(TREE.GENUS.GROUP, c("TREE.HYBRID.SPEC.1", "TREE.HYBRID.SPEC.2"), "^\\S*\\K\\s+", remove = FALSE) %>%
  dplyr::mutate(TREE.GENUS.GROUP = case_when(
    str_detect(TREE.GENUS.GROUP, "(?<![:alpha:])UNSPECIFIED(?![:alpha:])") ~ "Unspecified trees",
    TRUE ~ as.character(as.character(TREE.GENUS.GROUP))
  ))

trees.data.aggr <- trees.data %>%
  # Grouping and preparing the trees species group data for plotting
  group_by(TREE.GENUS.GROUP) %>%
  summarise(count = n()) %>%
  group_by(TREE.GENUS.GROUP = replace(TREE.GENUS.GROUP, count < 5, "Other species")) %>%
  summarise(count = sum(count)) %>%
  mutate(pct = (count / sum(count))) %>%
  filter(TREE.GENUS.GROUP != "NA") %>%
  arrange(desc(-count)) %>%
  ## order factor levels by number, put "Other" to end
  dplyr::mutate(
    TREE.GENUS.GROUP = forcats::fct_rev(forcats::fct_inorder(TREE.GENUS.GROUP)),
    TREE.GENUS.GROUP = forcats::fct_relevel(TREE.GENUS.GROUP, "Other species", after = 16)
  )

# Checking that the reordering worked
# levels(trees.data.aggr$TREE.GENUS.GROUP)

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(trees.data, "./DATASET.FROM.SCRIPT/trees.data.csv")
# -----------------------------------------------------------------------------------------------
glimpse(trees.data)
```

## Biodiversity indicators

Creating the dataset for **biodiversity indicators**
```{r Biodiversity indicators}
# using n_independent_studies because biodiversity and ecosystem service related indicators do not (normally) vary between sites, because independent studies look at the same variables (biodiversity and ecosystem service) even if several individual study sites are used for the study.

biodiversity <-
  database %>%
  filter(INCLUDED == TRUE) %>%
  # selecting Biodiversity columns
  dplyr::select(
    ID.P, ID.S, # ST.TYPE, MANAGEMENT
    starts_with("BD")
  ) %>%
  mutate(across(matches("BD"), as.logical)) %>%
  mutate_at(vars(-c(ID.P, ID.S)), tidyr::replace_na, 0) %>%
  # <- Including NA's as 0 (FALSE)
  mutate_if(is.logical, as.numeric) %>%
  mutate_if(is.character, as.numeric) %>%
  dplyr::select(-c(ID.S)) %>%
  distinct(ID.P, across(contains("BD"))) # including only independent studies and not separate study sites

# sum the data using apply
biodiversity.sum <-
  data.frame(value = apply(biodiversity, MARGIN = 2, FUN = sum))


biodiversity.sum$key <- rownames(biodiversity.sum)

biodiversity.data <-
  biodiversity.sum %>%
  rename(
    count = value,
    BD.SUB.INDICATORS = key
  ) %>%
  filter(BD.SUB.INDICATORS == "BD.SUB.FLORAMACRO" |
    BD.SUB.INDICATORS == "BD.SUB.FLORAMICRO" |
    BD.SUB.INDICATORS == "BD.SUB.FAUNAMACRO" |
    BD.SUB.INDICATORS == "BD.SUB.FAUNAMICRO" |
    BD.SUB.INDICATORS == "BD.SUB.FUNGI" |
    BD.SUB.INDICATORS == "BD.SUB.BACTERIA") %>%
  # grouping of flora and fauna categories according to the defined data structures and data analysis framework
  mutate(BD.GROUPS = case_when(
    str_detect(BD.SUB.INDICATORS, "BD.SUB.FLORAMACRO") ~ "FLORA",
    str_detect(BD.SUB.INDICATORS, "BD.SUB.FLORAMICRO") ~ "FLORA",
    str_detect(BD.SUB.INDICATORS, "BD.SUB.FAUNAMACRO") ~ "FAUNA",
    str_detect(BD.SUB.INDICATORS, "BD.SUB.FAUNAMICRO") ~ "FAUNA",
    str_detect(BD.SUB.INDICATORS, "BD.SUB.FUNGI") ~ "FUNGI",
    str_detect(BD.SUB.INDICATORS, "BD.SUB.BACTERIA") ~ "BACTERIA"
  )) %>%
  # grouping, counting and preparing the biodiversity groups for plotting
  group_by(BD.GROUPS) %>%
  summarise(count = sum(count)) %>%
  # the proportion of individual biodiversity-related indicators relative to the total number of papers: How many papers report in        biodiversity-related indicators
  mutate(pct_of_total = (count / n_independent_studies)) %>%
  arrange(desc(-count))

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(biodiversity.data, "./DATASET.FROM.SCRIPT/biodiversity.data.csv")
# -----------------------------------------------------------------------------------------------
glimpse(biodiversity.data)
```

## Ecosystem services (frequency of occurrence of ecosystem services)

Creating the dataset for assessing the **frequency of occurrence of
ecosystem services** using Pie-Donut plot

```{r}
# ---------------------------------------------------------------------------------------------------
# summarising ES Tier 2 data using the base R apply function

ecosystem.sum.data.PD <-
  data.frame(value = apply(ecosystem.data, # the dataset ecosystem.data has been generated earlier in the script
    MARGIN = 2, FUN = sum
  ))


ecosystem.sum.data.PD$ES.TIER2.GROUP <- rownames(ecosystem.sum.data.PD) # providing the row names from the original dataset

# ---------------------------------------------------------------------------------------------------
# Adding a column for Tier 1 ES ()

ecosystem.sum.data.PD.tier1.tier2 <-
  ecosystem.sum.data.PD %>%
  # dplyr::select(!key) %>%
  dplyr::filter(!ES.TIER2.GROUP %in% c(
    "ID.P", # excluding ID.P and Tier 1 ES rows as that would not make sense in the Pie-Donut chart
    "Provisioning ES",
    "Regulating ES",
    "Supporting ES",
    "Cultural ES",
    "Ecosystem Services"
  )) %>%
  as.tibble() %>%
  # ---------------------------------------------------------------------------------------------------
  # Specifying ES + CAT groups into ONLY ES = prov, regu, supp and cult. (CAT = agronomy, environment, economic and socio-cultural)
  # ---------------------------------------------------------------------------------------------------
  mutate(ES.TIER1.GROUP = "x") %>%
  mutate(ES.TIER1.GROUP = case_when(
    str_detect(ES.TIER2.GROUP, "Food") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Feed") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Fibre") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Fuel") ~ "Povisioning ES", # "Fuel (firewood etc.)"
    str_detect(ES.TIER2.GROUP, "Genetic resources") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Biochemicals") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Ornamentals") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Fresh water") ~ "Povisioning ES",

    # ---------------------------------------------------------------------------------------------------
    str_detect(ES.TIER2.GROUP, "Air regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Climate regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Carbon sequestration") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Water regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Water purification") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Nutrient leaching regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Erosion regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Water availability") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Disease regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Pest regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Polination regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Polination regulation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Natural hazards regulation") ~ "Regulating ES",
    # ---------------------------------------------------------------------------------------------------

    str_detect(ES.TIER2.GROUP, "Soil formation") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Photosynthesis") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Primary production") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Nutrient cycling") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Water cycling") ~ "Supporting ES",
    # ---------------------------------------------------------------------------------------------------

    str_detect(ES.TIER2.GROUP, "Cultural diversity") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Spiritual value") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Knowledge systems") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Educational value") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Inspirational value") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Aesthetic value") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Social relations") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Sense of place") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Cultural heritage value") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Recreational and ecotourism") ~ "Cultural ES",
    # ---------------------------------------------------------------------------------------------------

    str_detect(ES.TIER2.GROUP, "Food (yield)") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Biomass") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Land equivalent ratio") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Pests") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Diseases") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Weeds") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Product quality") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Product stability") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Mechanisa") ~ "Povisioning ES", # "Mechanisation"
    str_detect(ES.TIER2.GROUP, "Belowground") ~ "Povisioning ES", # "Belowground (crop-tree) competition"
    str_detect(ES.TIER2.GROUP, "Aboveground") ~ "Povisioning ES", # "Aboveground (crop-tree) light competition"
    str_detect(ES.TIER2.GROUP, "Nutrient use") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Nutrient use effeciency") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Breeding and selection") ~ "Povisioning ES",
    # ---------------------------------------------------------------------------------------------------

    str_detect(ES.TIER2.GROUP, "Soil quality") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Biodiversity") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Carbon stock") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Greenhouse gas emissions") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Pollination") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Soil structure") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Chemical soil components") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Biological soil components") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Soil water") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Soil erosion") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Soil formations") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Water quality") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Water perculation") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Water cyclings") ~ "Supporting ES",
    str_detect(ES.TIER2.GROUP, "Water availability") ~ "Regulating ES",
    str_detect(ES.TIER2.GROUP, "Hydraulic lift") ~ "Regulating ES",
    # ---------------------------------------------------------------------------------------------------

    str_detect(ES.TIER2.GROUP, "Economic profit related") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Economic management related") ~ "Povisioning ES",
    str_detect(ES.TIER2.GROUP, "Economic market related") ~ "Povisioning ES",
    # ---------------------------------------------------------------------------------------------------

    str_detect(ES.TIER2.GROUP, "Landscape aesthetics") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Recreational and ecotourism") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Spiritual values") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Perception of adoptation") ~ "Cultural ES",
    str_detect(ES.TIER2.GROUP, "Policy information") ~ "Cultural ES"
  )) %>%
  relocate(ES.TIER1.GROUP, ES.TIER2.GROUP, value)

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(ecosystem.sum.data.PD.tier1.tier2, "./DATASET.FROM.SCRIPT/ecosystem.sum.data.PD.tier1.tier2.csv")
# -----------------------------------------------------------------------------------------------
glimpse(ecosystem.sum.data.PD.tier1.tier2)
```

## Categories of Approach

Creating dataset for **categories of approach**

```{r Categories of Approach}
# using n_independent_studies because categories of approach do not (normally) vary between sites, because independent studies look at the same variables (categories of approach) even if several individual study sites are used for the study.

categories <-
  database %>%
  filter(INCLUDED == TRUE) %>%
  # selecting Ecosystem services columns
  dplyr::select(
    ID.P, ID.S, # ST.TYPE, MANAGEMENT
    starts_with("CAT")
  ) %>%
  mutate(across(matches("CAT"), as.logical)) %>%
  mutate_at(vars(-c(ID.P, ID.S)), tidyr::replace_na, 0) %>%
  # <- Including NA's as 0 (FALSE)
  mutate_if(is.logical, as.numeric) %>%
  mutate_if(is.character, as.numeric) %>%
  dplyr::select(-c(ID.S)) %>%
  distinct(ID.P, across(contains("CAT"))) # including only independent studies and not separate study sites


# ---------------------------------------------------------------------------------------------------
# Renaming data columns

categories.data <- categories %>%
  # rename_with(~ gsub("CAT.SUB.AGRO.", "", .x, fixed = TRUE))
  rename(
    "Food" = CAT.SUB.AGRO.YIELD,
    "Biomass" = CAT.SUB.AGRO.BIOMASS,
    "Land equivalent ratio" = CAT.SUB.AGRO.LER,
    "Pests" = CAT.SUB.AGRO.PESTS,
    "Diseases" = CAT.SUB.AGRO.DISEASE,
    "Weeds" = CAT.SUB.AGRO.WEEDS,
    "Product quality" = CAT.SUB.AGRO.PRODQUAL,
    "Product stability" = CAT.SUB.AGRO.PRODSTAB,
    "Mechanisation" = CAT.SUB.AGRO.MECHAN,
    "Belowground (crop-tree) competition" = CAT.SUB.AGRO.CRTRBLOW,
    "Aboveground (crop-tree) light competition" = CAT.SUB.AGRO.LIGHTCMP,
    "Nutrient use" = CAT.SUB.AGRO.NUTRIUSE,
    "Nutrient use effeciency" = CAT.SUB.AGRO.NUTRIEF,
    "Breeding and selection" = CAT.SUB.AGRO.BRDSELECT,
    "Soil quality" = CAT.SUB.ENVI.SOILQUAL,
    "Biodiversity" = CAT.SUB.ENVI.BIODIV,
    "Carbon stock" = CAT.SUB.ENVI.CSTOCK,
    "Greenhouse gas emissions" = CAT.SUB.ENVI.GHGEMS,
    "Pollination" = CAT.SUB.ENVI.POLLCON,
    "Soil structure" = CAT.SUB.ENVI.SOILSTRUCT,
    "Chemical soil components" = CAT.SUB.ENVI.SOILCHEM,
    "Biological soil components" = CAT.SUB.ENVI.SOILBIO,
    "Soil water" = CAT.SUB.ENVI.SOILWAT,
    "Soil erosion" = CAT.SUB.ENVI.SOILEROCO,
    "Soil formations" = CAT.SUB.ENVI.SOILFORM,
    "Water quality" = CAT.SUB.ENVI.WATQUAL,
    "Water perculation" = CAT.SUB.ENVI.WATPERC,
    "Water cyclings" = CAT.SUB.ENVI.WATCYC,
    "Water availability" = CAT.SUB.ENVI.WATAVAIL,
    "Hydraulic lift" = CAT.SUB.ENVI.HYDRAULIFT,
    "Economic profit related" = CAT.SUB.ECON.PROFIT,
    "Economic management related" = CAT.SUB.ECON.MANAGE,
    "Economic market related" = CAT.SUB.ECON.MARKET,
    "Landscape aesthetics" = CAT.SUB.SOCUL.LANAES,
    "Recreational and ecotourisms" = CAT.SUB.SOCUL.RECRECOTOUR,
    "Spiritual values" = CAT.SUB.SOCUL.SPIRITVAL,
    "Perception of adoptation" = CAT.SUB.SOCUL.PERCEPADOPT,
    "Policy information" = CAT.SUB.SOCUL.POLICYINFR,
    "Agronomic category" = CAT.AGRO,
    "Environmental category" = CAT.ENVIRON,
    "Economic category" = CAT.ECONOM,
    "Socio-cultural category" = CAT.SOCIOCUL
  )


# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(categories.data, "./DATASET.FROM.SCRIPT/categories.data.csv")
# -----------------------------------------------------------------------------------------------
glimpse(categories.data)
```

## Combined ES + CAT dataset

To create (1) broad focus groups of ecosystem services as well as (2)
selected ecosystem services we use information from the *data
dimensions*; **ecosystem services**, **biodiversity** and **categories
of approach**. Following the style of *Ditzler et al. (2021)*, *Duru et
al. (2015)* and *Tamburini et al. (2020)*

Also, to create a combined dataset for assessing **input / output ES**
we base the grouping of ecosystem services on the *data dimensions*;
**ecosystem services** and **categories of approach** combined.
Following the methodology of *Ditzler et al. (2021)* and *Duru et al.
(2015)*

```{r Combined ES + CAT dataset input / output ES and broad groups of ES, eval=FALSE}
# based on the datasets
# category.data.renamed AND ecosystem.data

joined.category.ecosystem <- left_join(
  x = categories.data, y = ecosystem.data,
  by = c("ID.P" = "ID.P")
) %>%
  relocate(ID.P)

category.ecosystem.long <- joined.category.ecosystem %>%
  pivot_longer(!ID.P,
    names_to = "name",
    values_to = "value"
  ) %>%
  # renaming dataframe rows and assigning them to input/output services, following the methodology of *Ditzler et al. (2021)* and
  # *Duru et al. (2015)*
  mutate(INPUT.OUTPUT.ES = recode(name,

    # --------------------------------------------------------------------------------------------------- Regular INPUT services
    `Erosion regulation` = "INPUT.ES",
    `Disease regulation` = "INPUT.ES",
    `Pest regulation` = "INPUT.ES",
    `Polination regulation` = "INPUT.ES",
    `Soil formation` = "INPUT.ES",
    `Nutrient use effeciency` = "INPUT.ES",
    `Land equivalent ratio` = "INPUT.ES",
    `Economic profit related` = "INPUT.ES",
    `Biodiversity` = "INPUT.ES",
    # ---------------------------------------------------------------------------------------------------

    `Soil structure` = "INPUT.ES",
    `Water availability` = "INPUT.ES",
    `Hydraulic lift` = "INPUT.ES",
    `Water cycling` = "INPUT.ES",
    `Nutrient cycling` = "INPUT.ES",
    # --------------------------------------------------------------------------------------------------- Regular OUTPUT services

    `Economic profit related` = "OUTPUT.ES",
    `Food` = "OUTPUT.ES",
    `Fibre` = "OUTPUT.ES",
    `Fuel (firewood etc.)` = "OUTPUT.ES",
    `Climate regulation` = "OUTPUT.ES",
    `Carbon sequestration` = "OUTPUT.ES",
    `Water regulation` = "OUTPUT.ES",
    `Nutrient leaching regulation` = "OUTPUT.ES",
    `Landscape aesthetics` = "OUTPUT.ES",
    `Recreational and ecotourism` = "OUTPUT.ES",
    `Product quality` = "OUTPUT.ES",
    `Pollination` = "OUTPUT.ES",
    `Product stability` = "OUTPUT.ES",
    `Recreational and ecotourisms` = "OUTPUT.ES",
    `Spiritual values` = "OUTPUT.ES",
    `Spiritual value` = "OUTPUT.ES",
    `Educational value` = "OUTPUT.ES",
    # ---------------------------------------------------------------------------------------------------

    `Economic management related` = "INPUT.ES",
    `Food (yield)` = "OUTPUT.ES",
    `Feed` = "OUTPUT.ES",
    `Biomass` = "OUTPUT.ES",
    `Biochemicals` = "OUTPUT.ES",
    `Ornamentals` = "OUTPUT.ES",
    .default = "NA"
  )) %>%
  # ---------------------------------------------------------------------------------------------------
  # Specifying ES + CAT groups into ONLY ES = prov, regu, supp and cult. (CAT = agronomy, environment, economic and socio-cultural)
  # ---------------------------------------------------------------------------------------------------
  mutate(ES.CAT.GROUPS = recode(name,
    `Food` = "Povisioning ES",
    `Feed` = "Povisioning ES",
    `Fibre` = "Povisioning ES",
    `Fuel` = "Povisioning ES", # "Fuel (firewood etc.)"
    `Genetic resources` = "Povisioning ES",
    `Biochemicals` = "Povisioning ES",
    `Ornamentals` = "Povisioning ES",
    `Fresh water` = "Povisioning ES",
    # `Fuel (firewood etc.)` = "Povisioning ES",
    # ---------------------------------------------------------------------------------------------------

    `Air regulation` = "Regulating ES",
    `Climate regulation` = "Regulating ES",
    `Carbon sequestration` = "Regulating ES",
    `Water regulation` = "Regulating ES",
    `Water purification` = "Regulating ES",
    `Nutrient leaching regulation` = "Regulating ES",
    `Erosion regulation` = "Regulating ES",
    `Water availability` = "Regulating ES",
    `Disease regulation` = "Regulating ES",
    `Pest regulation` = "Regulating ES",
    `Polination regulation` = "Regulating ES",
    `Natural hazards regulation` = "Regulating ES",
    # ---------------------------------------------------------------------------------------------------

    `Soil formation` = "Supporting ES",
    `Photosynthesis` = "Supporting ES",
    `Primary production` = "Supporting ES",
    `Nutrient cycling` = "Supporting ES",
    `Water cycling` = "Supporting ES",
    # ---------------------------------------------------------------------------------------------------

    `Cultural diversity` = "Cultural ES",
    `Spiritual value` = "Cultural ES",
    `Knowledge systems` = "Cultural ES",
    `Educational value` = "Cultural ES",
    `Inspirational value` = "Cultural ES",
    `Aesthetic value` = "Cultural ES",
    `Social relations` = "Cultural ES",
    `Sense of place` = "Cultural ES",
    `Cultural heritage value` = "Cultural ES",
    `Recreational and ecotourism` = "Cultural ES",
    # ---------------------------------------------------------------------------------------------------

    `Food (yield)` = "Povisioning ES",
    `Biomass` = "Povisioning ES",
    `Land equivalent ratio` = "Povisioning ES",
    `Pests` = "Regulating ES",
    `Diseases` = "Regulating ES",
    `Weeds` = "Regulating ES",
    `Product quality` = "Povisioning ES",
    `Product stability` = "Povisioning ES",
    `Mechanisation` = "Povisioning ES",
    `Belowground (crop-tree) competition` = "Povisioning ES",
    `Aboveground (crop-tree) light competition` = "Povisioning ES",
    `Nutrient use` = "Povisioning ES",
    `Nutrient use effeciency` = "Povisioning ES",
    `Breeding and selection` = "Povisioning ES",
    # ---------------------------------------------------------------------------------------------------

    `Soil quality` = "Supporting ES",
    `Biodiversity` = "Regulating ES",
    `Carbon stock` = "Regulating ES",
    `Greenhouse gas emissions` = "Regulating ES",
    `Pollination` = "Regulating ES",
    `Soil structure` = "Supporting ES",
    `Chemical soil components` = "Supporting ES",
    `Biological soil components` = "Supporting ES",
    `Soil water` = "Supporting ES",
    `Soil erosion` = "Regulating ES",
    `Soil formations` = "Supporting ES",
    `Water quality` = "Regulating ES",
    `Water perculation` = "Regulating ES",
    `Water cyclings` = "Supporting ES",
    `Water availability` = "Regulating ES",
    `Hydraulic lift` = "Regulating ES",
    # ---------------------------------------------------------------------------------------------------

    `Economic profit related` = "Povisioning ES",
    `Economic management related` = "Povisioning ES",
    `Economic market related` = "Povisioning ES",
    # ---------------------------------------------------------------------------------------------------

    `Landscape aesthetics` = "Cultural ES",
    `Recreational and ecotourism` = "Cultural ES",
    `Spiritual values` = "Cultural ES",
    `Perception of adoptation` = "Cultural ES",
    `Policy information` = "Cultural ES",
    # `Recreational and ecotourisms` = "Cultural ES",
    .default = "NA"
  )) %>%
  # ---------------------------------------------------------------------------------------------------
  # Specifying groups of ES + CAT
  # ---------------------------------------------------------------------------------------------------
  # # renaming dataframe rows and assigning them to broad focus groups of ES, following the methodology of Ditzler et al. (2021) and         # Tamburini et al. (2020)
  mutate(FOCUS.GROUPS.BD.ES.CAT = recode(name,

    # --------------------------------------------------------------------------------------------------- Production-related
    `Food` = "Production",
    `Feed` = "Production",
    `Fibre` = "Production",
    `Fuel` = "Production",
    `Genetic resources` = "Production",
    `Biochemicals` = "Production",
    `Ornamentals` = "Production",
    `Fuel (firewood etc.)` = "Production",
    `Primary production` = "Production",
    `Food (yield)` = "Production",
    `Biomass` = "Production",
    `Land equivalent ratio` = "Production",
    `Product quality` = "Production",
    `Product stability` = "Production",
    `Mechanisation` = "Production",
    `Belowground (crop-tree) competition` = "Production",
    `Aboveground (crop-tree) light competition` = "Production",
    `Nutrient use` = "Production",
    `Nutrient use effeciency` = "Production",
    `Breeding and selection` = "Production",
    # --------------------------------------------------------------------------------------------------- Nat. process regulation related

    `Air regulation` = "Nat. process regulation",
    `Climate regulation` = "Nat. process regulation",
    `Erosion regulation` = "Nat. process regulation",
    `Polination regulation` = "Nat. process regulation",
    `Pollination` = "Nat. process regulation",
    `Natural hazards regulation` = "Nat. process regulation",
    `Photosynthesis` = "Nat. process regulation",
    `Nutrient cycling` = "Nat. process regulation",
    # --------------------------------------------------------------------------------------------------- Water related

    `Water regulation` = "Water",
    `Water purification` = "Water",
    `Water availability` = "Water",
    `Fresh water` = "Water",
    `Water cycling` = "Water",
    `Water quality` = "Water",
    `Water perculation` = "Water",
    `Water cyclings` = "Water",
    `Water availability` = "Water",
    `Hydraulic lift` = "Water",
    # ------------------------------------------------------------------------------------  Soil quality, carbon and nutrients related

    `Carbon sequestration` = "Soil quality, carbon and nutrients",
    `Nutrient leaching regulation` = "Soil quality, carbon and nutrients",
    `Soil formation` = "Soil quality, carbon and nutrients",
    `Soil quality` = "Soil quality, carbon and nutrients",
    `Carbon stock` = "Soil quality, carbon and nutrients",
    `Greenhouse gas emissions` = "Soil quality, carbon and nutrients",
    `Soil structure` = "Soil quality, carbon and nutrients",
    `Chemical soil components` = "Soil quality, carbon and nutrients",
    `Biological soil components` = "Soil quality, carbon and nutrients",
    `Soil water` = "Soil quality, carbon and nutrients",
    `Soil erosion` = "Soil quality, carbon and nutrients",
    `Soil formations` = "Soil quality, carbon and nutrients",
    # --------------------------------------------------------------------------------------------------- Disease and pest related

    `Disease regulation` = "Disease and pest",
    `Pest regulation` = "Disease and pest",
    `Pests` = "Disease and pest",
    `Diseases` = "Disease and pest",
    `Weeds` = "Disease and pest",
    # --------------------------------------------------------------------------------------------------- Biodiversity related
    `Biodiversity` = "Biodiversity",
    # ------------------------------------------------------------------------------------------------- Socio-cultural-economic related

    `Cultural diversity` = "Socio-cultural and economic",
    `Spiritual value` = "Socio-cultural and economic",
    `Knowledge systems` = "Socio-cultural and economic",
    `Educational value` = "Socio-cultural and economic",
    `Inspirational value` = "Socio-cultural and economic",
    `Aesthetic value` = "Socio-cultural and economic",
    `Social relations` = "Socio-cultural and economic",
    `Sense of place` = "Socio-cultural and economic",
    `Cultural heritage value` = "Socio-cultural and economic",
    `Recreational and ecotourism` = "Socio-cultural and economic",
    `Economic profit related` = "Socio-cultural and economic",
    `Economic management related` = "Socio-cultural and economic",
    `Economic market related` = "Socio-cultural and economic",
    `Landscape aesthetics` = "Socio-cultural and economic",
    `Recreational and ecotourism` = "Socio-cultural and economic",
    `Spiritual values` = "Socio-cultural and economic",
    `Perception of adoptation` = "Socio-cultural and economic",
    `Policy information` = "Socio-cultural and economic",
    .default = "NA"
  ))

# ---------------------------------------------------------------------------------------------------------------------------------
# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(joined.category.ecosystem, "./DATASET.FROM.SCRIPT/joined.category.ecosystem.csv")
# -----------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(category.ecosystem.long, "./DATASET.FROM.SCRIPT/category.ecosystem.long.csv")
# -----------------------------------------------------------------------------------------------
glimpse(category.ecosystem.long)
```

## Average number of ecosystem services (ES + CAT) reported by papers

```{r}
no.ecosystem.paper.data <- ecosystem.data %>%
  dplyr::select(!c(`Provisioning ES`, `Regulating ES`, `Supporting ES`, `Cultural ES`, `Ecosystem Services`)) %>%
  drop_na() %>%
  mutate(ES_sum_per_paper = rowSums(.[2:35])) %>%
  mutate(avr_ES_per_paper = sum(ES_sum_per_paper) / n_independent_studies) %>%
  relocate(ES_sum_per_paper, avr_ES_per_paper) %>%
  arrange(desc(ES_sum_per_paper)) %>%
  mutate(across(where(is.numeric), round, 2))
  #dplyr::summarise(pct = mean(ES_sum_per_paper < 3, na.rm = TRUE)) # % observation > a value

# Defining quantiles 
q <- quantile(no.ecosystem.paper.data$ES_sum_per_paper)  

# Plotting histogram
no.ecosystem.paper.plot <- no.ecosystem.paper.data %>%
  ggplot(aes(x = ES_sum_per_paper)) +
  geom_histogram(aes(y = ..density..),
    colour = "black", fill = "white",
    binwidth = 1
  ) +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  geom_vline(aes(xintercept = avr_ES_per_paper),
             linetype ="dashed", size = 1, show.legend = FALSE, col = "black") + 
  geom_text(aes(x = avr_ES_per_paper + 0.6, label = paste0("Mean\n",avr_ES_per_paper), y = 0.25)) +
  annotate(geom = "text", x = q + 0.45, y = 0.2, label = names(q)) +
  theme(text = element_text(size = 10)) +
  geom_vline(xintercept = q, linetype = "dotted") +
  theme_bw() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        text = element_text(size = 18)) +
  scale_x_continuous(breaks = seq(0, 14, 1), limits = c(0, 14)) +
  labs(x = "Number of ecosystem services studied per paper",
       y = "") 

no.ecosystem.paper.plot
```

*Saving last plot*

```{r}
ggsave(
  filename = "no.ecosystem.paper.plot.png",
  plot = last_plot(),
  path = "./FIGURES.OUTPUT.FROM.SCRIPT",
  width = 35, 
  height = 15, 
  units = "cm",
  dpi = 500
)
```

# Multivariate datasets; matrices

## Tree - crop

Creating dataset for assessing **tree - crop associations**

```{r tree - crop associations datasets}
# -----------------------------------------------------------------Merging (joining) the crop and tree datasets based on ID.P and ID.S
joined.tree.crop <-
  left_join(
    x = trees.data, y = crops.data,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  dplyr::select(-(ends_with(".x") | ends_with(".y")))

# ---------------------------------------------------------------------------------------------------------------------------------
# Preparing dataset on crop groups - tree dataset
matrix.crop.groups.tree <-
  joined.tree.crop %>%
  dplyr::select(CROP.OUTPUT.GROUP, TREE.GENUS.GROUP) %>%
  filter(TREE.GENUS.GROUP != "NA") %>%
  # excluding rows with 'NA' in either crop or tree data
  filter(CROP.OUTPUT.GROUP != "NA") %>%
  filter(!str_detect(TREE.GENUS.GROUP, "Unspecified"))

matrix.crop.groups.tree.all <-
  matrix.crop.groups.tree %>%
  group_by(CROP.OUTPUT.GROUP, TREE.GENUS.GROUP) %>%
  dplyr::summarise(count = n()) %>%
  ungroup()

# ----------------------------------------------------------------------------------------------------------------------------------
# Preparing dataset on crop species - tree dataset

matrix.crop.species.tree <-
  joined.tree.crop %>%
  dplyr::select(CROP.GENUS.GROUP, TREE.GENUS.GROUP) %>%
  filter(TREE.GENUS.GROUP != "NA") %>%
  # excluding rows with 'NA' in either crop or tree data
  filter(CROP.GENUS.GROUP != "NA") %>%
  filter(!str_detect(TREE.GENUS.GROUP, "Unspecified")) %>%
  filter(!str_detect(CROP.GENUS.GROUP, "UNSPECIFIED"))

# ----------------------------------------------------------------------------------------------------------------------------------
matrix.crop.species.tree.all <-
  matrix.crop.species.tree %>%
  group_by(CROP.GENUS.GROUP, TREE.GENUS.GROUP) %>%
  dplyr::summarise(count = n()) %>%
  ungroup() %>%
  arrange(desc(count))

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(matrix.crop.species.tree.all, "./DATASET.FROM.SCRIPT/matrix.crop.species.tree.all.csv")
# Saving the saving dataset
readr::write_csv(matrix.crop.groups.tree.all, "./DATASET.FROM.SCRIPT/matrix.crop.groups.tree.all.csv")
# -----------------------------------------------------------------------------------------------
# -----------------------------------------------------------------------------------------------
glimpse(matrix.crop.species.tree.all)
```

## Biodiversity indicators - crop groups

Creating the dataset for assessing **biodiversity indicators and crops**

```{r }
# -----------------------------------------------------------------------Preparing biodiversity data with individual study sites
# using n_individual_study_sites because crop species vary between sites, as some studies include multiple sites with different species compositions and arrangements.

biodiversity.data.raw <-
  database %>%
  filter(INCLUDED == TRUE) %>%
  # selecting Biodiversity columns
  dplyr::select(
    ID.P, ID.S, # ST.TYPE, MANAGEMENT
    starts_with("BD")
  ) %>%
  mutate(across(matches("BD"), as.logical)) %>%
  mutate_at(vars(-c(ID.P, ID.S)), tidyr::replace_na, 0) %>%
  # <- Including NA's as 0 (FALSE)
  mutate_if(is.logical, as.numeric) %>%
  mutate_if(is.character, as.numeric) %>%
  # select(-c(ID.S)) %>%
  # distinct(ID.P, across(contains("BD"))) # including individual study sites and not only independent papers


  #  filter(INCLUDED == TRUE) %>%
  # grouping of flora and fauna categories according to the defined data structures and data analysis framework
  rename(
    "FLORA.1" = BD.SUB.FLORAMACRO,
    "FLORA.2" = BD.SUB.FLORAMICRO,
    "FAUNA.1" = BD.SUB.FAUNAMACRO,
    "FAUNA.2" = BD.SUB.FAUNAMICRO,
    "BACTERIA" = BD.SUB.BACTERIA,
    "FUNGI" = BD.SUB.FUNGI
  ) %>%
  mutate(
    FLORA = FLORA.1 + FLORA.2,
    FAUNA = FAUNA.1 + FAUNA.2
  ) %>%
  relocate(
    ID.P, ID.S,
    FLORA, FAUNA, BACTERIA, FUNGI
  ) %>%
  dplyr::select(-c(
    BD.SUB.SPEC, BD,
    FLORA.1, FAUNA.1, FLORA.2, FAUNA.2
  ))

# -------------------------------------------------------------------------------Preparing crops data with individual study sites
crop.data.raw <-
  crops.data %>%
  group_by(CROP.OUTPUT.GROUP) %>%
  relocate(ID.S, ID.P, CROP.SPECIES, CROP.OUTPUT.GROUP)

# ----------------------------------------------------- Merging (joining) the biodiversity and crops datasets based on ID.S and ID.P

joined.crop.biodiversity <-
  left_join(crop.data.raw, biodiversity.data.raw,
    by = c(
      "ID.S" = "ID.S",
      "ID.P" = "ID.P"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    CROP.SPECIES, CROP.OUTPUT.GROUP,
    FLORA, FAUNA, BACTERIA, FUNGI
  )

# ----------------------------------------------------------------------------------------------------------------------------------
# ------------------------------------------------------------------------- Preparing biodiversity-crops data for matrix plot

matrix.crop.biodiversity <-
  joined.crop.biodiversity %>%
  filter(
    CROP.OUTPUT.GROUP != "NA",
    FLORA != "NA",
    FAUNA != "NA",
    BACTERIA != "NA",
    FUNGI != "NA"
  )

matrix.crop.biodiversity[c("FLORA", "FAUNA", "BACTERIA", "FUNGI")] <-
  lapply(matrix.crop.biodiversity[c("FLORA", "FAUNA", "BACTERIA", "FUNGI")], as.logical)

matrix.crop.biodiversity.all <-
  matrix.crop.biodiversity %>%
  dplyr::select(c(CROP.OUTPUT.GROUP, FLORA, FAUNA, BACTERIA, FUNGI)) %>%
  mutate_if(is.logical, as.character) %>%
  pivot_longer(cols = c(-CROP.OUTPUT.GROUP)) %>%
  mutate(value = case_when(
    value == "TRUE" ~ 1,
    value == "FALSE" ~ 0
  )) %>%
  group_by(CROP.OUTPUT.GROUP, name) %>%
  summarize(count = sum(value)) %>%
  arrange(desc(count), CROP.OUTPUT.GROUP, name)

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the saving dataset
readr::write_csv(matrix.crop.biodiversity.all, "./DATASET.FROM.SCRIPT/matrix.crop.biodiversity.all.csv")
# -----------------------------------------------------------------------------------------------
glimpse(matrix.crop.biodiversity.all)
```

## Biodiversity indicators - SAF system trait types

Creating dataset for assessing the associations between **biodiversity
and SAF system trait types**

```{r Preparing system traits type data}
# using n_individual_study_sites because system trait types vary between sites, as some studies include multiple sites with different species compositions and arrangements.

system.traits.data.raw <-
  database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::mutate(ST.TYPE = case_when(
    str_detect(ST.TYPE, "RBS") ~ "Riparian buffer strips",
    str_detect(ST.TYPE, "SHB") ~ "Line belts *",
    str_detect(ST.TYPE, "MIXED") ~ "Mixed systems",
    str_detect(ST.TYPE, "TIMBER.ARABLE") ~ "Timber - Arable",
    str_detect(ST.TYPE, "BIOMASS.ARABLE") ~ "Biomass - Arable",
    str_detect(ST.TYPE, "NUT.ARABLE") ~ "Nut - Arable",
    str_detect(ST.TYPE, "FRUIT.ARABLE") ~ "Fruit - Arable",
    TRUE ~ as.character(as.character(ST.TYPE))
  )) %>%
  dplyr::mutate(ST.TYPE.PROD.TYPE = "x") %>%
  dplyr::mutate(ST.TYPE.PROD.TYPE == case_when(
    str_detect(ST.TYPE, "Timber - Arable") ~ "Biomass",
    str_detect(ST.TYPE, "Biomass - Arable") ~ "Biomass",
    str_detect(ST.TYPE, "Nut - Arable") ~ "Food",
    str_detect(ST.TYPE, "Fruit - Arable") ~ "Food",
    str_detect(ST.TYPE, "Riparian buffer strips") ~ "Multiple",
    str_detect(ST.TYPE, "Line belts *") ~ "Multiple",
    str_detect(ST.TYPE, "Mixed systems") ~ "Multiple",
    TRUE ~ as.character(as.character(ST.TYPE.PROD.TYPE))
  )) %>%
  # selecting system trait columns
  group_by(ST.TYPE) %>%
  dplyr::select(ID.P, ID.S, ST.TYPE, MANAGEMENT) %>%
  mutate_at(c("ID.P", "ID.S"), as.numeric) # <- converting ID columns from characters to numeric

# ---------------------------------------------------------- Merging (joining) the biodiversity and system trait datasets based on ID

joined.biodiversity.systemtrait <-
  left_join(biodiversity.data.raw, system.traits.data.raw,
    by = c(
      "ID.S" = "ID.S",
      "ID.P" = "ID.P"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ST.TYPE, MANAGEMENT, FLORA, FAUNA, BACTERIA, FUNGI
  )

# ---------------------------------------------------------- Preparing the biodiversity and system trait for matrix plot

matrix.biodiversity.systrait <-
  joined.biodiversity.systemtrait %>%
  dplyr::select(ST.TYPE, FLORA, FAUNA, BACTERIA, FUNGI) %>%
  pivot_longer(cols = c(-ST.TYPE)) %>%
  group_by(ST.TYPE, name) %>%
  summarize(count = sum(value)) %>%
  arrange(desc(count), ST.TYPE, name) %>%
  filter(!ST.TYPE == "NA")

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(matrix.biodiversity.systrait, "./DATASET.FROM.SCRIPT/matrix.biodiversity.systrait.csv")
# -----------------------------------------------------------------------------------------------
glimpse(matrix.biodiversity.systrait)
```

## Ecosystem services (broad ecosystem services) - crop groups

Creating the dataset for assessing the associations between **broad
ecosystem services and crop groups**

```{r Preparing ecosystems services data}
# using n_individual_study_sites because crop species vary between sites, as some studies include multiple sites with different species compositions and arrangements.

ecosystem.data.raw <-
  database %>%
  filter(INCLUDED == TRUE) %>%
  # selecting Ecosystem services columns
  dplyr::select(
    ID.P, ID.S, # ST.TYPE, MANAGEMENT
    starts_with("ES")
  ) %>%
  mutate(across(matches("ES"), as.logical)) %>%
  mutate_at(vars(-c(ID.P, ID.S)), tidyr::replace_na, 0) %>%
  # <- Including NA's as 0 (FALSE)
  mutate_if(is.logical, as.numeric) %>%
  mutate_if(is.character, as.numeric) %>%
  # including individual study sites and not only independent papers
  # select(-c(ID.S)) %>%
  # distinct(ID.P, across(contains("ES"))) # including only independent studies and not separate study sites

  rename(
    "Food (yield)" = ES.SUB.PROV.FOOD,
    "Feed" = ES.SUB.PROV.FEED,
    "Fibre" = ES.SUB.PROV.FIBRE,
    "Fuel (firewood etc.)" = ES.SUB.PROV.FUEL,
    "Genetic resources" = ES.SUB.PROV.GENRES,
    "Biochemicals" = ES.SUB.PROV.BIOCHEM,
    "Ornamentals" = ES.SUB.PROV.ORNA,
    "Fresh water" = ES.SUB.PROV.WATFRESH,
    "Air regulation" = ES.SUB.REGU.AIR,
    "Climate regulation" = ES.SUB.REGU.CLIMREG,
    "Carbon sequestration" = ES.SUB.REGU.CSEQREG,
    "Water regulation" = ES.SUB.REGU.WATREG,
    "Erosion regulation" = ES.SUB.REGU.EROREG,
    "Water purification" = ES.SUB.REGU.WATPUR,
    "Disease regulation" = ES.SUB.REGU.DISREG,
    "Pest regulation" = ES.SUB.REGU.PESTREG,
    "Polination regulation" = ES.SUB.REGU.POLREG,
    "Natural hazards regulation" = ES.SUB.REGU.NATHAZREG,
    "Soil formation" = ES.SUB.SUPP.SOILFORM,
    "Photosynthesis" = ES.SUB.SUPP.PHOTSYNT,
    "Primary production" = ES.SUB.SUPP.PRIMPROD,
    "Nutrient cycling" = ES.SUB.SUPP.NUTRICYC,
    "Water cycling" = ES.SUB.SUPP.WATCYC,
    "Cultural diversity" = ES.SUB.CULT.CULTDIV,
    "Spiritual value" = ES.SUB.CULT.SPIRITVAL,
    "Knowledge systems" = ES.SUB.CULT.KNOWLSYS,
    "Educational value" = ES.SUB.CULT.EDUVAL,
    "Inspirational value" = ES.SUB.CULT.INSPIR,
    "Aesthetic value" = ES.SUB.CULT.AESTVAL,
    "Social relations" = ES.SUB.CULT.SOCREL,
    "Sense of place" = ES.SUB.CULT.SENPLACE,
    "Cultural heritage value" = ES.SUB.CULT.CULTHERIVAL,
    "Recreational and ecotourism" = ES.SUB.CULT.RECRECOTOUR,
    "Provisioning ES" = ES.PROV,
    "Regulating ES" = ES.REGU,
    "Supporting ES" = ES.SUPP,
    "Cultural ES" = ES.CULT,
    "Ecosystem Services" = ES
  )

# ------------------------------------------------- Merging (joining) the ecosystem services and crops datasets based on ID

joined.crop.ecosystem <-
  left_join(crop.data.raw, ecosystem.data.raw,
    by = c(
      "ID.S" = "ID.S",
      "ID.P" = "ID.P"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    CROP.SPECIES, CROP.OUTPUT.GROUP,
    `Provisioning ES`, `Regulating ES`, `Supporting ES`, `Cultural ES`
  )

# ------------------------------------------------------------ Preparing ecosystems services - crops data for matrix plot*

matrix.crop.ecosystem <-
  joined.crop.ecosystem %>%
  ungroup() %>%
  dplyr::select(c(CROP.OUTPUT.GROUP, contains("ES"))) %>%
  dplyr::select(c(
    CROP.OUTPUT.GROUP,
    `Provisioning ES`, `Regulating ES`, `Supporting ES`, `Cultural ES`
  )) %>%
  filter(
    CROP.OUTPUT.GROUP != "NA",
    `Provisioning ES` != "NA",
    `Regulating ES` != "NA",
    `Supporting ES` != "NA",
    `Cultural ES` != "NA"
  ) %>%
  mutate_if(is.numeric, as.logical) %>%
  mutate_if(is.logical, as.character) %>%
  pivot_longer(cols = c(-CROP.OUTPUT.GROUP)) %>%
  mutate(value = case_when(
    value == "TRUE" ~ 1,
    value == "FALSE" ~ 0
  )) %>%
  group_by(CROP.OUTPUT.GROUP, name) %>%
  summarize(count = sum(value)) %>%
  arrange(desc(count), CROP.OUTPUT.GROUP, name)

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(matrix.crop.ecosystem, "./DATASET.FROM.SCRIPT/matrix.crop.ecosystem.csv")
# -----------------------------------------------------------------------------------------------
glimpse(matrix.crop.ecosystem)
```

## Ecosystem services - SAF system trait types

Creating the dataset for assessing the associations between **ecosystem
services and SAF system trait types**

```{r }
# ------------------------------------ Merging (joining) the ecosystem services and system trait types datasets based on ID
joined.ecosystem.systemtrait <-
  left_join(ecosystem.data.raw, system.traits.data.raw,
    by = c(
      "ID.S" = "ID.S",
      "ID.P" = "ID.P"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ST.TYPE, MANAGEMENT,
    `Provisioning ES`, `Regulating ES`, `Supporting ES`, `Cultural ES`
  )

# -------------------------------------------------------------- Preparing the ecosystem services and system trait for matrix plot
matrix.ecosystem.systrait <-
  joined.ecosystem.systemtrait %>%
  dplyr::select(ST.TYPE, `Provisioning ES`, `Regulating ES`, `Supporting ES`, `Cultural ES`) %>%
  pivot_longer(cols = c(-ST.TYPE)) %>%
  group_by(ST.TYPE, name) %>%
  summarize(count = sum(value)) %>%
  arrange(desc(count), ST.TYPE, name) %>%
  filter(!ST.TYPE == "NA")

# ------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(matrix.ecosystem.systrait, "./DATASET.FROM.SCRIPT/matrix.ecosystem.systrait.csv")
# -----------------------------------------------------------------------------------------------
glimpse(matrix.ecosystem.systrait)
```

## Input/output ES - publication year

Creating dataset for assessing the **developments in the relationship
between input/output ES and publication years**

```{r, warning=FALSE, message=FALSE}
# ------------------------------------------------------------------------- Preparing publication years data
pub.year.data.ID <-
  database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::select(ID.P, PUB.YR) %>%
  mutate_if(is.character, as.numeric) %>%
  distinct(ID.P, PUB.YR) %>%
  group_by(PUB.YR) %>%
  add_tally() %>%
  mutate(no_studies_per_year = n) %>%
  dplyr::select(-n)
# add_tally() %>%

# -------------------------- Merging (joining) the input/output ES and publication years datasets based on ID.P

joined.inoutES.pub.yr <-
  left_join(
    x = category.ecosystem.long, y = pub.year.data.ID,
    by = c("ID.P" = "ID.P")
  ) %>%
  dplyr::select(-value)

# --------------------------------------- Preparing dataset on input/output ES in relation to publication years
matrix.inoutES.pub.yr <-
  joined.inoutES.pub.yr %>%
  filter(INPUT.OUTPUT.ES != "NA")

matrix.inoutES.pub.yr[c("ID.P")] <-
  lapply(matrix.inoutES.pub.yr[c("ID.P")], as.numeric)

matrix.inoutES.pub.yr[c("PUB.YR")] <-
  lapply(matrix.inoutES.pub.yr[c("PUB.YR")], as.numeric)

matrix.inoutES.pub.yr[c("ES.CAT.GROUPS")] <-
  lapply(matrix.inoutES.pub.yr[c("ES.CAT.GROUPS")], as.factor)

matrix.inoutES.pub.yr.all <-
  matrix.inoutES.pub.yr %>%
  group_by(INPUT.OUTPUT.ES, PUB.YR, ES.CAT.GROUPS) %>%
  summarise(count = no_studies_per_year) %>%
  distinct(INPUT.OUTPUT.ES, PUB.YR, ES.CAT.GROUPS, .keep_all = TRUE) %>%
  ungroup()

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(matrix.inoutES.pub.yr.all, "./DATASET.FROM.SCRIPT/matrix.inoutES.pub.yr.all.csv")
# -----------------------------------------------------------------------------------------------
glimpse(matrix.inoutES.pub.yr.all)
```

## Ecosystem services (ES focus groups) - crop groups - trees genera

Creating dataset with **ES focus groups in a association with crop
groups and tree genera** based on ecosystem services and categories of
approach combined, following the methodology of *Ditzler et al. (2021)*
and *Tamburini et al. (2020)*

```{r ES focus groups in a association with crop groups and tree genera}
# ---------------------------------------------------------------- Input output ES merged with crop groups and tree genera dataset
joined.focus.ES.crops.trees <-
  left_join(
    x = category.ecosystem.long, y = joined.tree.crop,
    by = c("ID.P" = "ID.P")
  ) %>%
  dplyr::select(c(
    ID.P,
    FOCUS.GROUPS.BD.ES.CAT, INPUT.OUTPUT.ES, ES.CAT.GROUPS,
    TREE.GENUS.GROUP,
    CROP.OUTPUT.GROUP, CROP.GENUS.GROUP
  ))
# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(joined.focus.ES.crops.trees, "./DATASET.FROM.SCRIPT/joined.focus.ES.crops.trees.csv")
# -----------------------------------------------------------------------------------------------
glimpse(joined.focus.ES.crops.trees)
```

## Dataset: Selected ecosystem services

## Ecosystem services (selected ecosystem services) - crop groups - SAF system trait types

Creating dataset for assessing the associations between **ecosystem
services (selected ES), crop groups and SAF system trait types**

```{r }
# -------------------------------------------------------------------------------- Preparing combined data on  ES, BD, CAT
es.bd.cat.systrait.data <-
  database %>%
  filter(INCLUDED == TRUE) %>%
  dplyr::select(
    ID.P, ID.S,
    ST.TYPE, # selecting Silvoarable System Trait type
    starts_with("ES"), # selecting ecosystem services, biodiversity and categories data
    starts_with("BD"),
    starts_with("CAT")
  ) %>%
  mutate(across(matches("ES"), as.logical)) %>%
  mutate(across(matches("BD"), as.logical)) %>%
  mutate(across(matches("CAT"), as.logical)) %>%
  mutate(across(matches("ST.TYPE"), as.character)) %>%
  mutate_at(vars(-c(ID.P, ID.S, ST.TYPE)), tidyr::replace_na, 0) %>%
  # <- Including NA's as 0 (FALSE)
  mutate_if(is.logical, as.numeric) %>%
  mutate(across(matches("ID"), as.numeric))
# mutate_if(is.character, as.numeric)

# ------------------------------------------------------------------------------- renaming SAF system trait type variables
es.bd.cat.systrait.data.renamed <-
  es.bd.cat.systrait.data %>%
  dplyr::mutate(ST.TYPE = case_when(
    str_detect(ST.TYPE, "RBS") ~ "Riparian buffer strips",
    str_detect(ST.TYPE, "SHB") ~ "Line belts *",
    str_detect(ST.TYPE, "MIXED") ~ "Mixed systems",
    str_detect(ST.TYPE, "TIMBER.ARABLE") ~ "Timber - Arable",
    str_detect(ST.TYPE, "BIOMASS.ARABLE") ~ "Biomass - Arable",
    str_detect(ST.TYPE, "NUT.ARABLE") ~ "Nut - Arable",
    str_detect(ST.TYPE, "FRUIT.ARABLE") ~ "Fruit - Arable",
    TRUE ~ as.character(as.character(ST.TYPE))
  )) %>%
  dplyr::mutate(ST.TYPE.PROD.TYPE = "x") %>%
  mutate(ST.TYPE.PROD.TYPE = case_when(
    ST.TYPE == "Timber - Arable" ~ "Biomass",
    ST.TYPE == "Biomass - Arable" ~ "Biomass",
    ST.TYPE == "Riparian buffer strips" ~ "Biomass",
    ST.TYPE == "Nut - Arable" ~ "Food",
    ST.TYPE == "Fruit - Arable" ~ "Food",
    ST.TYPE == "Line belts *" ~ "Nature + wind shelter",
    ST.TYPE == "Mixed systems" ~ "Mixed"
  )) %>%
  # -------------------------------------------------------------------------------- renaming ecosystem services variables
  rename(
    "SAF System Trait type" = ST.TYPE,
    "Food (yield)" = ES.SUB.PROV.FOOD,
    "Feed" = ES.SUB.PROV.FEED,
    "Fibre" = ES.SUB.PROV.FIBRE,
    "Fuel (firewood etc.)" = ES.SUB.PROV.FUEL,
    "Genetic resources" = ES.SUB.PROV.GENRES,
    "Biochemicals" = ES.SUB.PROV.BIOCHEM,
    "Ornamentals" = ES.SUB.PROV.ORNA,
    "Fresh water" = ES.SUB.PROV.WATFRESH,
    "Air regulation" = ES.SUB.REGU.AIR,
    "Climate regulation" = ES.SUB.REGU.CLIMREG,
    "Carbon sequestration" = ES.SUB.REGU.CSEQREG,
    "Water regulation" = ES.SUB.REGU.WATREG,
    "Nutrient leaching regulation" = ES.SUB.REGU.NUTRILE,
    "Erosion regulation" = ES.SUB.REGU.EROREG,
    "Water purification" = ES.SUB.REGU.WATPUR,
    "Disease regulation" = ES.SUB.REGU.DISREG,
    "Pest regulation" = ES.SUB.REGU.PESTREG,
    "Polination regulation" = ES.SUB.REGU.POLREG,
    "Natural hazards regulation" = ES.SUB.REGU.NATHAZREG,
    "Soil formation" = ES.SUB.SUPP.SOILFORM,
    "Photosynthesis" = ES.SUB.SUPP.PHOTSYNT,
    "Primary production" = ES.SUB.SUPP.PRIMPROD,
    "Nutrient cycling" = ES.SUB.SUPP.NUTRICYC,
    "Water cycling" = ES.SUB.SUPP.WATCYC,
    "Cultural diversity" = ES.SUB.CULT.CULTDIV,
    "Spiritual value" = ES.SUB.CULT.SPIRITVAL,
    "Knowledge systems" = ES.SUB.CULT.KNOWLSYS,
    "Educational value" = ES.SUB.CULT.EDUVAL,
    "Inspirational value" = ES.SUB.CULT.INSPIR,
    "Aesthetic value" = ES.SUB.CULT.AESTVAL,
    "Social relations" = ES.SUB.CULT.SOCREL,
    "Sense of place" = ES.SUB.CULT.SENPLACE,
    "Cultural heritage value" = ES.SUB.CULT.CULTHERIVAL,
    "Recreational and ecotourism" = ES.SUB.CULT.RECRECOTOUR,
    "Provisioning ES" = ES.PROV,
    "Regulating ES" = ES.REGU,
    "Supporting ES" = ES.SUPP,
    "Cultural ES" = ES.CULT,
    "Ecosystem Services" = ES,
    # ----------------------------------------------------------------------------------- renaming biodiversity variables
    "Associated biodiversity" = BD, # Biodiversity -associated biodiversity (from Beillouin et al. 2021)
    "FLORA.1" = BD.SUB.FLORAMACRO,
    "FLORA.2" = BD.SUB.FLORAMICRO,
    "FAUNA.1" = BD.SUB.FAUNAMACRO,
    "FAUNA.2" = BD.SUB.FAUNAMICRO,
    "Bacteria" = BD.SUB.BACTERIA,
    "Fungi" = BD.SUB.FUNGI,
    # -------------------------------------------------------------------------- renaming categories of approach variables
    "Food" = CAT.SUB.AGRO.YIELD,
    "Biomass" = CAT.SUB.AGRO.BIOMASS,
    "Land equivalent ratio" = CAT.SUB.AGRO.LER,
    "Pests" = CAT.SUB.AGRO.PESTS,
    "Diseases" = CAT.SUB.AGRO.DISEASE,
    "Weeds" = CAT.SUB.AGRO.WEEDS,
    "Product quality" = CAT.SUB.AGRO.PRODQUAL,
    "Production stability" = CAT.SUB.AGRO.PRODSTAB,     # Product stability
    "Mechanisation" = CAT.SUB.AGRO.MECHAN,
    "Belowground competition" = CAT.SUB.AGRO.CRTRBLOW,  # Belowground (crop-tree) competition
    "Aboveground competition" = CAT.SUB.AGRO.LIGHTCMP,  # Aboveground (crop-tree) light competition
    "Nutrient use" = CAT.SUB.AGRO.NUTRIUSE,
    "Nutrient use effeciency" = CAT.SUB.AGRO.NUTRIEF,
    "Breeding and selection" = CAT.SUB.AGRO.BRDSELECT,
    "Soil quality" = CAT.SUB.ENVI.SOILQUAL,
    "Biodiversity" = CAT.SUB.ENVI.BIODIV, # Biodiversity -associated biodiversity (from Beillouin et al. 2021)
    "Carbon stock" = CAT.SUB.ENVI.CSTOCK,
    "Greenhouse gas emissions" = CAT.SUB.ENVI.GHGEMS,
    "Pollination" = CAT.SUB.ENVI.POLLCON,
    "Soil structure" = CAT.SUB.ENVI.SOILSTRUCT,
    "Chemical soil components" = CAT.SUB.ENVI.SOILCHEM,
    "Biological soil components" = CAT.SUB.ENVI.SOILBIO,
    "Soil water" = CAT.SUB.ENVI.SOILWAT,
    "Soil erosion" = CAT.SUB.ENVI.SOILEROCO,
    "Soil formations" = CAT.SUB.ENVI.SOILFORM,
    "Water quality" = CAT.SUB.ENVI.WATQUAL,
    "Water perculation" = CAT.SUB.ENVI.WATPERC,
    "Water cyclings" = CAT.SUB.ENVI.WATCYC,
    "Water availability" = CAT.SUB.ENVI.WATAVAIL,
    "Hydraulic lift" = CAT.SUB.ENVI.HYDRAULIFT,
    "Profitability" = CAT.SUB.ECON.PROFIT,       # Economic profit related
    "Management" = CAT.SUB.ECON.MANAGE,          # Economic management related
    "Market" = CAT.SUB.ECON.MARKET,              # Economic market related
    "Landscape aesthetics" = CAT.SUB.SOCUL.LANAES,
    "Recreational and ecotourisms" = CAT.SUB.SOCUL.RECRECOTOUR,
    "Spiritual values" = CAT.SUB.SOCUL.SPIRITVAL,
    "Perception of adoptation" = CAT.SUB.SOCUL.PERCEPADOPT,
    "Policy information" = CAT.SUB.SOCUL.POLICYINFR,
    "Agronomic category" = CAT.AGRO,
    "Environmental category" = CAT.ENVIRON,
    "Economic category" = CAT.ECONOM,
    "Socio-cultural category" = CAT.SOCIOCUL 
  ) %>%
  # --------------------------------------------------------------------------------------- simplifying flora, fauna groups
  mutate(
    Flora = FLORA.1 + FLORA.2,
    Fauna = FAUNA.1 + FAUNA.2
  )


# -------------------------------------------------------------------------------------------------------------------------
# ------------------------------------------ Merging (joining) ES, BD, CAT dataset with crop datasets based on ID.P and ID.S

joined.es.bd.cat.systrait.crop <-
  left_join(
    x = es.bd.cat.systrait.data.renamed, y = crops.data,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  dplyr::select(-(ends_with(".x") | ends_with(".y"))) %>%
  mutate_at(c("ID.P", "ID.S"), as.numeric) # <- converting ID columns from characters to numeric

```

**Defining the final dataset for selected Ecosystem Services**
```{r}
# -------------------------------------------------------------------------------------------------------------------------
# Defining the final selected Ecosystem Services based on 
# Ditzler et al. (2021), Beillouin et al. 2021 and Fagerholm et al. (2016) and 
# -------------------------------------------------------------------------------------------------------------------------
matrix.es.bd.cat.systrait.crop <-
  joined.es.bd.cat.systrait.crop %>%
  filter(CROP.OUTPUT.GROUP != "NA") %>%
  filter(!is.na(CROP.GENUS.GROUP))

matrix.es.bd.cat.systrait.crop.all <-
  matrix.es.bd.cat.systrait.crop %>%
  dplyr::select(
    # SAF system trait types, crop species info --------------------------
    "SAF System Trait type",
    "CROP.GENUS.GROUP",
    "CROP.OUTPUT.GROUP",
    # --------------------------------------------------------------------
#--# Biodiversity ---------------------------------------------
    "Associated biodiversity",
#--# Supporting ecosystem services ----------------------------
    "Soil formation",
    "Soil structure",
    "Water cycling",
    "Nutrient cycling",
    "Carbon stock",
        # "Chemical soil components",
        # "Biological soil components",
        # "Greenhouse gas emissions",
#--# Regulating ecosystem services ----------------------------
    "Erosion regulation",
    "Carbon sequestration",
    "Water regulation",
    "Water perculation",
    "Climate regulation",    
    "Pest regulation",
    "Disease regulation",
    "Polination regulation",
        # "Water quality",
#--# Provisioning ecosystem services --------------------------
    "Profitability",
    "Aboveground competition",
    "Water availability",
    "Production stability",
    "Feed",
    "Fuel (firewood etc.)",
    "Food (yield)",
        # "Land equivalent ratio",
        # "Management",
#--# Cultural ecosystem services ------------------------------
    "Recreational and ecotourism",
    "Landscape aesthetics",
    "Policy information"
  ) %>%
  mutate_if(is.character, as.factor) %>%
  relocate(`SAF System Trait type`, CROP.GENUS.GROUP, CROP.OUTPUT.GROUP)

```

```{r}
# ----------------------------------------------------------------------------------------------------------------------
# ------------------------------------------------------------------ Merging (joining) with crop based on ID.P and ID.S*

A_ES_ID <-
  matrix.es.bd.cat.systrait.crop %>%
  dplyr::select(
    # ID --------------------------
    "ID.P",
    "ID.S",
     # --------------------------------------------------------------------
#--# Biodiversity ---------------------------------------------
    "Associated biodiversity",
#--# Supporting ecosystem services ----------------------------
    "Soil formation",
    "Soil structure",
    "Water cycling",
    "Nutrient cycling",
    "Carbon stock",
        # "Chemical soil components",
        # "Biological soil components",
        # "Greenhouse gas emissions",
#--# Regulating ecosystem services ----------------------------
    "Erosion regulation",
    "Carbon sequestration",
    "Water regulation",
    "Water perculation",
    "Climate regulation",    
    "Pest regulation",
    "Disease regulation",
    "Polination regulation",
        # "Water quality",
#--# Provisioning ecosystem services --------------------------
    "Profitability",
    "Aboveground competition",
    "Water availability",
    "Production stability",
    "Feed",
    "Fuel (firewood etc.)",
    "Food (yield)",
        # "Land equivalent ratio",
        # "Management",
#--# Cultural ecosystem services ------------------------------
    "Recreational and ecotourism",
    "Landscape aesthetics",
    "Policy information"
  ) %>%
  mutate_if(is.character, as.factor) %>%
  relocate(ID.P, ID.S) %>%
  pivot_longer(
    cols = c(-ID.P, -ID.S),
    names_to = "ES.GROUPS",
    values_to = "count"
  ) %>%
  distinct() %>%
  filter(count == 1)

# ---------------------------------------------------------------------
B_SAFTT_ID <- joined.es.bd.cat.systrait.crop %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "SAF System Trait type"
  ) %>%
  distinct()

# ---------------------------------------------------------------------
C_CROPS_ID <- matrix.es.bd.cat.systrait.crop %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "CROP.OUTPUT.GROUP"
  ) %>%
  distinct()

# -----------------------------------------------------------------------
D_JOINED_A_B <-
  left_join(
    x = A_ES_ID, y = B_SAFTT_ID,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`
  ) %>%
  distinct()

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with crop groups
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C <-
  left_join(
    x = D_JOINED_A_B, y = C_CROPS_ID,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, CROP.OUTPUT.GROUP
  ) %>%
  distinct() %>%
  dplyr::select(-count)

ES.selected.crop.groups.SATF <- E_JOINED_D_C

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(ES.selected.crop.groups.SATF, "./DATASET.FROM.SCRIPT/ES.selected.crop.groups.SATF.csv")
# -----------------------------------------------------------------------------------------------
glimpse(ES.selected.crop.groups.SATF)
```

```{r}
E_JOINED_D_C
# Next, we will identify the duplicated rows by each country and date and save into response2:

#Mutate
response2 <- E_JOINED_D_C %>% 
  mutate_if(is.character, as.factor) %>%
  group_by(ES.GROUPS, `SAF System Trait type`, CROP.OUTPUT.GROUP) %>%
  summarise(count = n())
  #mutate(count_2 = count/4)
  #mutate(across(where(is.numeric), round, 0))
```

## Ecosystem services (selected ecosystem services) - tree genera - functional role (product type) of tree component of SAF system trait types

Creating dataset for assessing the associations between **ecosystem
services (selected ES), tree genera and \## Ecosystem services (selected
ecosystem services) - tree genera - functional role (product type) of
tree component of SAF system trait types**

```{r}
# -----------------------------------------------Merging (joining) ES, BD, CAT dataset with tree datasets based on ID.P and ID.S

joined.es.bd.cat.systrait.tree <-
  left_join(
    x = es.bd.cat.systrait.data.renamed, y = trees.data,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  dplyr::select(-(ends_with(".x") | ends_with(".y"))) %>%
  mutate_at(c("ID.P", "ID.S"), as.numeric) # <- converting ID columns from characters to numeric

# ------------------------------------------------------------------------------------------------------------------------------
matrix.es.bd.cat.systrait.tree <-
  joined.es.bd.cat.systrait.tree %>%
  filter(TREE.GENUS.GROUP != "NA") %>%
  filter(!is.na(TREE.GENUS.GROUP))

matrix.es.bd.cat.systrait.tree.all <-
  matrix.es.bd.cat.systrait.tree %>%
    dplyr::select(
    # SAF system trait types, tree species info --------------------------
    "SAF System Trait type",
    "TREE.GENUS.GROUP",
    "ST.TYPE.PROD.TYPE",
       # --------------------------------------------------------------------
#--# Biodiversity ---------------------------------------------
    "Associated biodiversity",
#--# Supporting ecosystem services ----------------------------
    "Soil formation",
    "Soil structure",
    "Water cycling",
    "Nutrient cycling",
    "Carbon stock",
        # "Chemical soil components",
        # "Biological soil components",
        # "Greenhouse gas emissions",
#--# Regulating ecosystem services ----------------------------
    "Erosion regulation",
    "Carbon sequestration",
    "Water regulation",
    "Water perculation",
    "Climate regulation",    
    "Pest regulation",
    "Disease regulation",
    "Polination regulation",
        # "Water quality",
#--# Provisioning ecosystem services --------------------------
    "Profitability",
    "Aboveground competition",
    "Water availability",
    "Production stability",
    "Feed",
    "Fuel (firewood etc.)",
    "Food (yield)",
        # "Land equivalent ratio",
        # "Management",
#--# Cultural ecosystem services ------------------------------
    "Recreational and ecotourism",
    "Landscape aesthetics",
    "Policy information"
  ) %>%
  mutate_if(is.character, as.factor) %>%
  relocate(`SAF System Trait type`, TREE.GENUS.GROUP, ST.TYPE.PROD.TYPE)

# -----------------------------------------------------------------------------------------------------------------------
# ------------------------------------------------------------------- Merging (joinging) with crop based on ID.P and ID.S*

A_ES_ID_tree <-
  matrix.es.bd.cat.systrait.tree %>%
  dplyr::select(
    # ID --------------------------
    "ID.S",
    "ID.P",
    # --------------------------------------------------------------------
#--# Biodiversity ---------------------------------------------
    "Associated biodiversity",
#--# Supporting ecosystem services ----------------------------
    "Soil formation",
    "Soil structure",
    "Water cycling",
    "Nutrient cycling",
    "Carbon stock",
        # "Chemical soil components",
        # "Biological soil components",
        # "Greenhouse gas emissions",
#--# Regulating ecosystem services ----------------------------
    "Erosion regulation",
    "Carbon sequestration",
    "Water regulation",
    "Water perculation",
    "Climate regulation",    
    "Pest regulation",
    "Disease regulation",
    "Polination regulation",
        # "Water quality",
#--# Provisioning ecosystem services --------------------------
    "Profitability",
    "Aboveground competition",
    "Water availability",
    "Production stability",
    "Feed",
    "Fuel (firewood etc.)",
    "Food (yield)",
        # "Land equivalent ratio",
        # "Management",
#--# Cultural ecosystem services ------------------------------
    "Recreational and ecotourism",
    "Landscape aesthetics",
    "Policy information"
  ) %>%
  mutate_if(is.character, as.factor) %>%
  relocate(ID.P, ID.S) %>%
  pivot_longer(
    cols = c(-ID.P, -ID.S),
    names_to = "ES.GROUPS",
    values_to = "count"
  ) %>%
  distinct() %>%
  filter(count == 1)

# ---------------------------------------------------------------------
B_SAFTT_ID_tree <- joined.es.bd.cat.systrait.tree %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "SAF System Trait type"
  )

# ---------------------------------------------------------------------
C_TREE_ID <- matrix.es.bd.cat.systrait.tree %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "TREE.GENUS.GROUP",
    "ST.TYPE.PROD.TYPE"
  ) %>%
  distinct()

# -----------------------------------------------------------------------
D_JOINED_A_B_tree <-
  left_join(
    x = A_ES_ID_tree, y = B_SAFTT_ID_tree,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`
  ) %>%
  distinct()

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with crop groups
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_tree <-
  left_join(
    x = D_JOINED_A_B_tree, y = C_TREE_ID,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, TREE.GENUS.GROUP, ST.TYPE.PROD.TYPE
  ) %>%
  distinct() %>%
  dplyr::select(-count)

ES.selected.tree.genera.SATF <- E_JOINED_D_C_tree


# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(ES.selected.tree.genera.SATF, "./DATASET.FROM.SCRIPT/ES.selected.tree.genera.SATF.csv")

# -----------------------------------------------------------------------------------------------
glimpse(ES.selected.tree.genera.SATF)
```

%\>% distinct()

```{r}
# ---------------------------------------------------------------------------------------------------------------------------------
ES_TREE <-
  left_join(
    x = C_TREE_ID %>% dplyr::select(-ST.TYPE.PROD.TYPE), y = A_ES_ID_tree,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS
  ) %>%
  distinct()


# Saving the dataset
readr::write_csv(ES_TREE, "./DATASET.FROM.SCRIPT/ES_TREE.csv")
# -----------------------------------------------------------------------------------------------
glimpse(ES_TREE)

```

## Ecosystem services (focus group ecosystem services) - crop groups - SAF system trait types

Creating dataset for assessing the associations between **ecosystem
services (focus group ES), crop groups and SAF system trait types**

```{r focus ecosystem services) - SAF system trait types - crop groups}
# Merging (joining) the Focus BD ES groups and SAF systems trait types datasets based on ID.P

joined.focus.ES.SAFT <-
  left_join(
    x = category.ecosystem.long, y = joined.es.bd.cat.systrait.crop,
    by = c("ID.P" = "ID.P")
  ) %>%
  dplyr::select(c(
    ID.P,
    FOCUS.GROUPS.BD.ES.CAT, INPUT.OUTPUT.ES, ES.CAT.GROUPS,
    CROP.OUTPUT.GROUP,
    `SAF System Trait type`
  )) %>%
  relocate(`SAF System Trait type`, FOCUS.GROUPS.BD.ES.CAT, CROP.OUTPUT.GROUP)

# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(joined.focus.ES.SAFT, "./DATASET.FROM.SCRIPT/joined.focus.ES.SAFT.csv")
# -----------------------------------------------------------------------------------------------
glimpse(joined.focus.ES.SAFT)
```

## Ecosystem services (selected ecosystem services) - countries of publication

Creating the dataset for assessing the associations between **selected
ES and countries of publication**

```{r}
ES_ID <-
  matrix.es.bd.cat.systrait.crop %>%
  dplyr::select(
    # ID --------------------------
    "ID.S",
    "ID.P",
    # --------------------------------------------------------------------
    # Biodiversity ---------------------------------------------
    "Associated biodiversity",
    # Supporting ecosystem services ----------------------------
    "Soil formation",
    "Soil structure",
    "Water cycling",
    "Nutrient cycling",
    "Greenhouse gas emissions",
    # Regulating ecosystem services ----------------------------
    "Carbon sequestration",
    "Climate regulation",
    "Water regulation",
    "Water quality",
    "Pest regulation",
    "Disease regulation",
    "Polination regulation",
    # Provisioning ecosystem services --------------------------
    "Land equivalent ratio",
    "Aboveground competition",
    "Management",
    "Profitability", 
    "Production stability",
    "Feed",
    "Fuel (firewood etc.)",
    "Food (yield)",
    # Cultural ecosystem services ------------------------------
    "Recreational and ecotourism",
    "Landscape aesthetics",
    "Policy information"
  ) %>%
  mutate_if(is.character, as.factor) %>%
  relocate(ID.P, ID.S) %>%
  pivot_longer(
    cols = c(-ID.P, -ID.S),
    names_to = "ES.GROUPS",
    values_to = "count"
  ) %>%
  distinct() %>%
  filter(count == 1)
# --------------------------------------------------------------------------------------------------------------------------
A_COUNTRIES_ID <- database %>%
  mutate_at(c("ID.P", "ID.S"), as.numeric) %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "STUDY.LOC.COUNTRY"
  ) %>%
  distinct()
# --------------------------------------------------------------------------------------------------------------------------
D_JOINED_COUNTRIES_ES <-
  left_join(
    x = A_COUNTRIES_ID, y = ES_ID,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS
  )
# --------------------------------------------------------------------------------------------------------------------------

B_CROPS_ID <- matrix.es.bd.cat.systrait.crop %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "CROP.OUTPUT.GROUP"
  ) %>%
  distinct()
# --------------------------------------------------------------------------------------------------------------------------
D_JOINED_COUNTRIES_CROPS_ES <-
  left_join(
    x = D_JOINED_COUNTRIES_ES, y = B_CROPS_ID,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct()
# --------------------------------------------------------------------------------------------------------------------------
C_SAFTT_ID <- joined.es.bd.cat.systrait.crop %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "SAF System Trait type"
  ) %>%
  distinct()
# --------------------------------------------------------------------------------------------------------------------------
# Merged country, crop and SAF system trait type dataset
# --------------------------------------------------------------------------------------------------------------------------
D_JOINED_COUNTRIES_CROPS_ES_SAFT <-
  left_join(
    x = D_JOINED_COUNTRIES_CROPS_ES, y = C_SAFTT_ID,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  dplyr::select(-count) %>%
  relocate(STUDY.LOC.COUNTRY, ES.GROUPS, CROP.OUTPUT.GROUP, `SAF System Trait type`)

ES.countries.crops.SAFT <- D_JOINED_COUNTRIES_CROPS_ES_SAFT
# ---------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(ES.countries.crops.SAFT, "./DATASET.FROM.SCRIPT/ES.countries.crops.SAFT.csv")
# -----------------------------------------------------------------------------------------------
glimpse(ES.countries.crops.SAFT)
```

*Proportion of studies represented by the top 10 countries*

```{r}
database %>%
  dplyr::select(STUDY.LOC.COUNTRY, ST.TYPE) %>%
  drop_na() %>%
  group_by(STUDY.LOC.COUNTRY, ST.TYPE) %>%
  summarise(count = n()) %>%
  mutate(pct = count / sum(count)) %>%
  mutate(across(where(is.numeric), round, 3)) %>%
  ungroup() 

database %>%
  dplyr::select(STUDY.LOC.COUNTRY, ST.TYPE) %>%
  drop_na() %>%
  dplyr::group_by(STUDY.LOC.COUNTRY) %>%
  summarise(count = n()) %>%
  dplyr::mutate(count_country = sum(count)) %>%
  dplyr::group_by(STUDY.LOC.COUNTRY, add = TRUE) %>%
  dplyr::mutate(pct_per_country = paste0(round(100 * count / count_country, 2), '%')) %>%
  slice_max(count, n = 10) %>%
  arrange(desc(count))

# 19.25+ 18.31+13.62+10.33+9.86+9.39+7.04+3.76+3.76+1.41 ~ 97 %
```

# Geospatial data; single variate data (not matrices)

## Adding coordinates to database

**Deriving latitude and longitude of study site locations** - *be aware:
this piece of code takes a while to run*

```{r Deriving latitude and longitude, warning=FALSE, message=FALSE, eval = TRUE}
# Notice this piece of code takes a ffew minutes, as hundreds of lat long coordinates has to be derived based on names in the STUDY.LOC.SPEC.COORDS column

database.lat.lon <- database %>%
  tidygeocoder::geocode(STUDY.LOC.SPEC.COORDS, method = "osm", lat = latitude, long = longitude)

# ----------------------------------------------------------------------------------------------------------------
# Saving the dataset with derived coordinates unto local disc
readr::write_csv(database.lat.lon, "./DATASET.FROM.SCRIPT/database.lat.lon.csv")
# --------------------------------------------------------------------
glimpse(database.lat.lon)
```





# Deriving pedoclimatic variables from BioClim and SoilGrids



-   RQ2: How are the (major and minor) themes of temperate silvoarable
    agroforestry systems distributed across geographical, pedoclimatic
    and socio-economic gradients?

##  Bioclimatic variables (WorldClim)

*A little info about using WorldClim variables in spatial analysis in r*

BioClim variables available in WorldClim

pedoclimatic variables are derived from the monthly temperature and
rainfall values in order to generate more biologically meaningful
variables. These are often used in species distribution modelling and
related ecological modelling techniques. The pedoclimatic variables
represent annual trends (e.g., mean annual temperature, annual
precipitation) seasonality (e.g., annual range in temperature and
precipitation) and extreme or limiting environmental factors (e.g.,
temperature of the coldest and warmest month, and precipitation of the
wet and dry quarters). A quarter is a period of three months (1/4 of the
year). (see more here: <https://www.worldclim.org/data/bioclim.html>)

Don't forget that WorldClim data has a scale factor of 10 for
temperature data, so Temp = -37 is -3.7 ºC.

BIO1 = Annual Mean Temperature BIO2 = Mean Diurnal Range (Mean of
monthly (max temp - min temp)) BIO3 = Isothermality (BIO2/BIO7) (×100)
BIO4 = Temperature Seasonality (standard deviation × 100) BIO5 = Max
Temperature of Warmest Month BIO6 = Min Temperature of Coldest Month
BIO7 = Temperature Annual Range (BIO5-BIO6) BIO8 = Mean Temperature of
Wettest Quarter BIO9 = Mean Temperature of Driest Quarter BIO10 = Mean
Temperature of Warmest Quarter BIO11 = Mean Temperature of Coldest
Quarter BIO12 = Annual Precipitation BIO13 = Precipitation of Wettest
Month BIO14 = Precipitation of Driest Month BIO15 = Precipitation
Seasonality (Coefficient of Variation) BIO16 = Precipitation of Wettest
Quarter BIO17 = Precipitation of Driest Quarter BIO18 = Precipitation of
Warmest Quarter BIO19 = Precipitation of Coldest Quarter

The WorldClim data are available at different spatial resolutions
(expressed as minutes or seconds of a degree of longitude and latitude):
10 minutes, 5 minutes, 2.5 minutes, 30 seconds. The variables included
are monthly minimum and maximum temperature, precipitation, and
'pedoclimatic' variables. The original WorldClim data were at a 30
second resolution, the other data have been derived through aggregation,
by calculating the mean of groups of cells. The 10-minute spatial
resolution is about 18.6 x 18.6 = 344 km2 at the equator. Here we are
going to use 2.5-minute spatial resolution (res=2.5), this is about 4.5
km at the equator according to WorldClim
<https://www.worldclim.org/data/v1.4/cmip5_2.5m.html>

To get (projected) future climate data (CMIP5), you must provide
arguments var and res as above. Only resolutions 2.5, 5, and 10 are
currently available. In addition, you need to provide model, rcp and
year. For example, getData('CMIP5', var='tmin', res=10, rcp=85,
model='AC', year=70)

Getting WorldClim data through r

If name = 'worldclim' you must also provide arguments var, and a
resolution res. Valid variables names are 'tmin', 'tmax', 'prec' and
'bio'. Valid resolutions are 0.5, 2.5, 5, and 10 (minutes of a degree).
In the case of res=0.5, you must also provide a lon and lat argument for
a tile; for the lower resolutions global data will be downloaded. In all
cases there are 12 (monthly) files for each variable except for 'bio'
which contains 19 files.

-   getData('worldclim', var='tmin', res=0.5, lon=5, lat=45)

-   getData('worldclim', var='bio', res=10)

see also more information:
<https://worldclim.org/data/v1.4/formats.html> and on the blog post of
Emily Piche <https://emilypiche.github.io/BIO381/raster.html> and Peter
Galante <https://rsh249.github.io/spatial_bioinformatics/worldclim.html>

**Extracting/downloading WorldClim (BioClim) variables; Temperature and precipitation and merging dataset with the database**

**Dataframe with coordinates** *Loading dataset*

```{r }

database.lat.lon <- vroom::vroom("./DATASET.FROM.SCRIPT/database.lat.lon.csv") 
```

```{r Creating the simple dataframe with coordinates}
# Make a dataframe with coordinates
study.site.coords <- database.lat.lon %>%
  relocate(
    ID.P, ID.S, TITLE,
    latitude, longitude, STUDY.LOC.COUNTRY, STUDY.LOC.SPEC, STUDY.LOC.SPEC.COORDS
  ) %>%
  dplyr::select(longitude, latitude)

# latitude and longitude coordinate points converted to dataframe
latitude <- as.data.frame(study.site.coords[, 2])
longitude <- as.data.frame(study.site.coords[, 1])

# latitude and longitude coordinate points converted to vectors in order to work in the getData() function
latitude <- as.list(latitude)
longitude <- as.list(longitude)

# coordinates used for the getData() function
coords <- data.frame(
  x = longitude,
  y = latitude
)
```

*Extracting/downloading WorldClim (BioClim) variables using the getData() function from the raster package*

```{r Extracting/downloading BioClim variables, warning=FALSE, message=FALSE, eval=FALSE}
# downloading pedoclimatic variables from worldclim at a resolution of 2.5 minutes
# this is saved as a RasterStack object that contains all the 19 bioclim variables

# be aware of the warning ""getData will be removed in a future version of raster\n. Please use the geodata package instead"
# in the future the getData() function from raster might be depreciated. Instead use geodata::worldclim_global()
bioclim.rasterstack <- raster::getData("worldclim",
  var = "bio",
  res = 2.5
)

# specifying that we only need; temperature (BIO1 = Annual Mean Temperature) and precipitation (BIO12 = Annual Precipitation)
bioclim.temp.precip <- bioclim.rasterstack[[c(
  1,  # Bio 1 Annual Mean Temperature
  3,  # Bio 3 Isothermality
  4,  # Bio 4 Temperature Seasonality (Standard Deviation)
  5,  # Bio 5 Maximum Temperature of Warmest Month
  6,  # Bio 6 Minimum Temperature of Coldest Month
  7,  # Bio 7 Annual Temperature Range
  10, # Bio 10 Mean Temperature of Warmest Quarter
  12, # Bio 12 Annual Precipitation
  14, # Bio 14 Precipitation of Driest Month
  15, # Bio 15 Precipitation Seasonality (Coefficient of Variation)
  17  # Bio 17 Precipitation of Driest Quarter
)]]
names(bioclim.temp.precip) <- c(
  "bio01_Annual_mean_temp",
  "bio03_Isothermality",
  "bio04_Temp_seasonality",
  "bio05_Max_temp_warmest_month",
  "bio06_Min_temp_coldest_month",
  "bio07_Annual_temp_range",
  "bio10_Mean_temp_warmest_quarter",
  "bio12_Annual_prec",
  "bio14_Prec_driest_month",
  "bio15_Prec_seasonality",
  "bio17_Prec_driest_quarter"
)

# converting the data into a SpatialPoints dataframe by adding the same coordinate reference system
points <- sp::SpatialPoints(coords, proj4string = bioclim.temp.precip@crs)

# extracting point-specific information on temperature and precipitation from the bioclim RasterStack object using the point coordinates
temp.precip.values <- raster::extract(
  x = bioclim.temp.precip,
  y = points
) %>%
  raster::as.data.frame()


# Saving the temp.precip.values dataset with derived coordinates unto local disc
# readr::write_csv(temp.precip.values, "./DATASET.FROM.SCRIPT/temp.precip.values.csv")

# --------------------------------------------------------------------------------------------------------
# Merging (joining) the location-specific temperature and precipitation values to the coordinates (x,y)

bioclim.point.data <- cbind.data.frame(
  sp::coordinates(points),
  temp.precip.values
)

bioclim.point.data.tbl <- as_tibble(bioclim.point.data)

# ------------------------------------------------------------------------------------------------
database.coords.bioclim <-
  left_join(
    x = database.lat.lon, y = bioclim.point.data.tbl,
    by = c(
      "longitude" = "longitude",
      "latitude" = "latitude"
    )
  ) %>%
  mutate_at(c("ID.P", "ID.S"), as.numeric) %>%
  relocate(
    ID.P, ID.S, TITLE,
    STUDY.LOC.COUNTRY, STUDY.LOC.SPEC, STUDY.LOC.SPEC.COORDS,
    latitude, longitude,
    
    bio01_Annual_mean_temp,
    bio03_Isothermality,
    bio04_Temp_seasonality,
    bio05_Max_temp_warmest_month,
    bio06_Min_temp_coldest_month,
    bio07_Annual_temp_range,
    bio10_Mean_temp_warmest_quarter,
    bio12_Annual_prec,
    bio14_Prec_driest_month,
    bio15_Prec_seasonality,
    bio17_Prec_driest_quarter
  )
# ------------------------------------------------------------------------------------------------
# Saving the dataset with derived coordinates
readr::write_csv(database.coords.bioclim, "./DATASET.FROM.SCRIPT/database.coords.bioclim.csv")
# ------------------------------------
glimpse(database.coords.bioclim)
```

## Soil variables (SoilGrids - ISRIC)

*A little info about using SoilGrids variables in spatial analysis in r*

SoilGrids is a system for global digital soil mapping that uses
state-of-the-art machine learning methods to map the spatial
distribution of soil properties across the globe. SoilGrids prediction
models are fitted using over 230 000 soil profile observations from the
WoSIS database and a series of environmental covariates. Covariates were
selected from a pool of over 400 environmental layers from Earth
observation derived products and other environmental information
including climate, land cover and terrain morphology. The outputs of
SoilGrids are global soil property maps at six standard depth intervals
(according to the GlobalSoilMap IUSS working group and its
specifications) at a spatial resolution of 250 meters. Prediction
uncertainty is quantified by the lower and upper limits of a 90%
prediction interval. The additional uncertainty layer displayed at
soilgrids.org is the ratio between the inter-quantile range and the
median. The SoilGrids maps are publicly available under the CC-BY 4.0
License.

Maps of the following soil properties are available: pH, soil organic
carbon content, bulk density, coarse fragments content, sand content,
silt content, clay content, cation exchange capacity (CEC), total
nitrogen as well as soil organic carbon density and soil organic carbon
stock. SoilGrids predictions are made for the six standard depth
intervals specified in the GlobalSoilMap IUSS working group and its
specifications. The depth intervals returned are: "0-5cm", "5-15cm",
"15-30cm", "30-60cm", "60-100cm", "100-200cm".

Find out more information about the SoilGrids and GlobalSoilMap products
here: <https://www.isric.org/explore/soilgrids/faq-soilgrids> and check
ISRIC's data policy, disclaimer and citation:
<https://www.isric.org/about/data-policy>. and here:
<https://www.isric.org/sites/default/files/GlobalSoilMap_specifications_december_2015_2.pdf>

The soilsDB package contains a collection of functions for reading data
from various digital soil libraries and soil maps, including SoilGrids.
Find more information about the soilsDB package here:
<https://ncss-tech.github.io/AQP/soilDB/soilDB-Intro.html>

The fetchSoilGrids() function from soilDB obtains SoilGrids properties
information (250 m raster resolution) given a data.frame containing site
IDs, latitudes and longitudes. SoilGrids API and maps return values as
whole (integer) numbers to minimize the storage space used. These values
are converted by to produce conventional units by 'fetchSoilGrids()'

The soil properties from ISRIC SoilGrids returned by the
fetchSoilGrids() function are:

-   bdod, Bulk density of the fine earth fraction; mapped unit:
    [cg/cm\^3], conversion: 100, conventional factor unit: [kg/dm\^3]
-   cec, Cation Exchange Capacity of the soil. [mmol(c)/kg 10
    cmol(c)/kg]
-   cfvo, Volumetric fraction of coarse fragments (\> 2 mm); mapped
    unit: [cm^3/dm^3 (vol per mil)], conversion: 10, conventional factor
    unit: [cm^3/100cm^3 (vol%)]
-   clay, Proportion of clay particles (\< 0.002 mm) in the fine earth
    fraction; mapped unit: [g/kg], conversion: 10, conventional factor
    unit: [g/100g (%)]
-   nitrogen, Total nitrogen (N); mapped unit: [cg/kg], conversion: 100,
    conventional factor unit: [g/kg]
-   phh2o, Soil pH; mapped unit: [pH\*10], conversion: 10, conventional
    factor unit: pH
-   sand, Proportion of sand particles (\> 0.05 mm) in the fine earth
    fraction; mapped unit: [g/kg], conversion: 10, conventional factor
    unit: [g/100g (%)]
-   silt, Proportion of silt particles (= 0.002 mm and = 0.05 mm) in the
    fine earth fraction; mapped unit: [g/kg], conversion: 10,
    conventional factor unit: [g/100g (%)]
-   soc, Soil organic carbon content in the fine earth fraction; mapped
    unit: [dg/kg], conversion: 10, conventional factor unit: [g/kg]
-   ocd, Organic carbon density; mapped unit: [hg/m\^3], conversion: 10,
    conventional factor unit: [kg/m\^3]
-   ocs, Organic carbon stocks; mapped unit: [t/ha], conversion: 10,
    conventional factor unit: [kg/m\^2]

Read more about units and conversion factors of SoilGrids here:
<https://www.isric.org/explore/soilgrids/faq-soilgrids>

**Extracting/downloading the soil variables from SoilGrids using the
soilDB package**

*Dataframe with coordinates and ID*

```{r Creating the study site location dataframe}
# creating simple dataframe that only contains id and location (lat., lon.)
database.locations.df <-
  database.lat.lon %>%
  dplyr::select(
    ID.P, ID.S,
    latitude, longitude
  ) %>%
  dplyr::mutate_at(vars(ID.P, ID.S), as.numeric) %>%
  # conversing the ID.S + ID.P columns into only one "id" column by collapsing the two ID.S + ID.P columns
  rowwise() %>%
  mutate(ID.PAIR = paste0(sort(c(ID.P, ID.S)), collapse = " ")) %>%
  separate(ID.PAIR, " ", into = c("ID.P", "ID.S")) %>%
  mutate(ID = row_number()) %>%
  # renaming the lat. lon. columns and converting to dataframe
  dplyr::rename(
    "lat" = latitude,
    "lon" = longitude,
    "id" = ID
  ) %>%
  dplyr::mutate_at(vars(id), as.numeric) %>%
  dplyr::select(id, lat, lon) %>%
  as.data.frame()
```

*Extracting/downloading soil variables from ISRIC SoilGrids using the
fetchSoilGrids() function from the soilsDB package.*

```{r Extracting/downloading SoilGrids soil variables, warning=FALSE, message=FALSE, eval=FALSE}
# be aware that this function call takes quite some time, this is also why results are saved to local disc
SoilGrids.data <- soilDB::fetchSoilGrids(
  x = database.locations.df,
  progress = TRUE
)

# not a dataframe: class(SoilGrids.data)


# ---------------------------------------------------------------------------------------------------------------
# Saving the downloaded SoilGrids dataset to local disc
 readr::write_rds(SoilGrids.data, "./DATASET.FROM.SCRIPT/SoilGrids.data.Rds")
```

**Generating dataframe with fetched soil variables - merging with
database**

*Loading dataset*

```{r }
# Loading the downloaded SoilGrids dataset to local disc
SoilGrids.data <- readr::read_rds("./DATASET.FROM.SCRIPT/SoilGrids.data.Rds")
```

```{r Dataframe with fetched soil variables}
# selecting only relevant variables
fetchedSoilGrid.df <-
  aqp::horizons(SoilGrids.data) %>%
  dplyr::select(id, label, claymean, sandmean, siltmean, bdodmean, nitrogenmean, socmean, phh2omean, cecmean) %>%
  # notice: we exclude NA values - these are empty values from locations located in urban areas where SoilGrids do not have info
  drop_na(claymean, sandmean, siltmean, bdodmean, nitrogenmean, socmean, phh2omean, cecmean) %>%
  dplyr::mutate_at(vars(id), as.numeric) %>%
  dplyr::arrange(id)
```

*Merging (joining) the location-specific soil values to the coordinates
(x,y - with lat. lon.)*

```{r Joining location-specific soil info to the original database}
# first, we merge the SoilGrids soil data with the dataframe containing locations (using the 'id' column)
merged.soilgrids <-
  left_join(
    x = fetchedSoilGrid.df, y = database.locations.df,
    by = c("id" = "id")
  ) %>%
  relocate(id, lat, lon, label) %>%
  as_tibble()

# secondly, we join this newly formed dataset to the 'original' dataframe that contains all other information (using coordinates)
database.coords.soilgrids <-
  left_join(
    x = merged.soilgrids, y = database.lat.lon,
    by = c(
      "lat" = "latitude",
      "lon" = "longitude"
    )
  ) %>%
  mutate_at(c("ID.P", "ID.S"), as.numeric) %>%
  relocate(
    ID.P, ID.S,
    lat, lon
  )

# where there is NA values in SoilGrids data on soil variables will be empty. Hence only 107 observations have soil data

# Saving the combined dataset with biopphysical information unto local disc
readr::write_csv(database.coords.soilgrids, "./DATASET.FROM.SCRIPT/database.coords.soilgrids.csv")
```

*Merging (joining) the location-specific soil dataset to the location-specific bioclim dataset (using ID.p and ID.S)*

**Loading dataset**

```{r Loading data - SAF system trait types (agroforestry system types)}

database.coords.soilgrids <- vroom::vroom("./DATASET.FROM.SCRIPT/database.coords.soilgrids.csv") 
database.coords.bioclim <- vroom::vroom("./DATASET.FROM.SCRIPT/database.coords.bioclim.csv") 
```


```{r, eval=FALSE}
database.biophys <-
  left_join(
    x = database.coords.soilgrids, y = database.coords.bioclim,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    ),
    suffix = c("", ".y")
  ) %>%
  # avoiding the ".x" and ".y" suffix notation
  dplyr::select(-(ends_with(".y"))) %>%
  dplyr::select(-id) %>%
  relocate(
    ID.P, ID.S,
    lat, lon,
    # Pedoclimatic variables from BioClim 
    bio01_Annual_mean_temp,
    bio03_Isothermality,
    bio04_Temp_seasonality,
    bio05_Max_temp_warmest_month,
    bio06_Min_temp_coldest_month,
    bio07_Annual_temp_range,
    bio10_Mean_temp_warmest_quarter,
    bio12_Annual_prec,
    bio14_Prec_driest_month,
    bio15_Prec_seasonality,
    bio17_Prec_driest_quarter,
    # Pedoclimatic variables from SoilGrids 
    claymean, 
    sandmean, 
    siltmean, 
    bdodmean, 
    nitrogenmean, 
    socmean, 
    phh2omean, 
    cecmean
  )

# ----------------------------------------------------------------------------------------------------------------------
# Saving the combined dataset with pedoclimatic information unto local disc
readr::write_csv(database.biophys, "./DATASET.FROM.SCRIPT/database.biophys.csv")
# -----------------------------------------------------------------
glimpse(database.biophys)
```

## Global elevation data

*A little info about using global elevation in spatial analysis in r*

*Extracting elevation data*

```{r}

```

## Tree genera - biophys

-   RQ2.a - How does the system properties temperature and precipitation
    drive inter-site differences?

Creating the **trees - biophys dataset**

```{r Preparing the trees - biophys dataset, warning=FALSE, message=FALSE}
# loading the trees data from excel
trees <- read_excel("DATA/DATABASE_2.0.xlsx",
  sheet = "TREE_SPECIES"
) %>%
  # Generating broad categories of tree species - based on families, genus and hybrids
  separate(TREE.SPECIES, c("TREE.FAMILY.GROUP", "TREE.GENUS.GROUP"), "^\\S*\\K\\s+", remove = FALSE) %>%
  separate(TREE.GENUS.GROUP, c("TREE.HYBRID.SPEC.1", "TREE.HYBRID.SPEC.2"), "^\\S*\\K\\s+", remove = FALSE) %>%
  dplyr::mutate(TREE.FAMILY.GROUP = case_when(
    str_detect(TREE.FAMILY.GROUP, "(?<![:alpha:])UNSPECIFIED(?![:alpha:])") ~ "Unspecified trees",
    TRUE ~ as.character(as.character(TREE.FAMILY.GROUP))
  )) %>%
  filter(TREE.FAMILY.GROUP != "NA") %>%
  relocate(ID.P, ID.S, TREE.SPECIES, TREE.FAMILY.GROUP) %>%
  mutate_at(c("ID.P", "ID.S"), as.numeric)


# -------------------------------------------------------------------- merging trees data with the biophys database
trees.biophys.data <-
  full_join(database.biophys, trees,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  )

# ----------------------------------------------------------------------------------------------------------------------
# Saving the combined dataset with biopphysical information unto local disc
readr::write_csv(trees.biophys.data, "./DATASET.FROM.SCRIPT/trees.biophys.data.csv")
# -----------------------------------------------------------------
glimpse(trees.biophys.data)
```

## Crop species (groups) - biophys

Creating the **crops - biophys dataset**

```{r }
# loading the trees data
crops <- read_excel("DATA/DATABASE_2.0.xlsx",
  sheet = "CROP_SPECIES"
) %>%
  # Generating broad categories of crop species - based on broad output product
  # based on FAO, 2019. FAOSTAT database. http://www.fao.org/faostat/en/#data.
  mutate(CROP.OUTPUT.GROUP = case_when(
    str_detect(CROP.SPECIES, "Triticum") ~ "CEREALS CROPS",
    str_detect(CROP.SPECIES, "Zea") ~ "CEREALS CROPS",
    str_detect(CROP.SPECIES, "Hordeum") ~ "CEREALS CROPS",
    str_detect(CROP.SPECIES, "Secale") ~ "CEREALS CROPS",
    str_detect(CROP.SPECIES, "Avena") ~ "CEREALS CROPS",
    str_detect(CROP.SPECIES, "Sorghum") ~ "CEREALS CROPS",
    str_detect(CROP.SPECIES, "Fagopyrum") ~ "CEREALS CROPS",
    str_detect(CROP.SPECIES, "Glycine") ~ "PULSES",
    str_detect(CROP.SPECIES, "Pisum") ~ "PULSES",
    str_detect(CROP.SPECIES, "Cicer") ~ "PULSES",
    str_detect(CROP.SPECIES, "Lupinus") ~ "PULSES",
    str_detect(CROP.SPECIES, "Beta") ~ "SUGAR CROPS",
    str_detect(CROP.SPECIES, "Cucurbita") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "UNSPECIFIED.VEGETABLES") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "Asparagus") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "Brassica oleracea") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "Capsicum") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "Citrullus") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "Cucumis") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "Lactuca") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "Solanum lycopersicum") ~ "VEGETABLES CROPS",
    str_detect(CROP.SPECIES, "Brassica napus subsp. napus") ~ "OIL CROPS",
    str_detect(CROP.SPECIES, "Helianthus") ~ "OIL CROPS",
    str_detect(CROP.SPECIES, "Sinapsis") ~ "OIL CROPS",
    str_detect(CROP.SPECIES, "Brassica alba") ~ "OIL CROPS",
    str_detect(CROP.SPECIES, "Ribes") ~ "FRUIT CROPS",
    str_detect(CROP.SPECIES, "Rubus") ~ "FRUIT CROPS",
    str_detect(CROP.SPECIES, "Aronia") ~ "FRUIT CROPS",
    str_detect(CROP.SPECIES, "Fragaria") ~ "FRUIT CROPS",
    str_detect(CROP.SPECIES, "Solanum tuberosum") ~ "ROOTS AND TUBER CROPS",
    str_detect(CROP.SPECIES, "HERBS") ~ "OTHER CROPS FOR HUMAN CONS.",
    str_detect(CROP.SPECIES, "SP.BEEF.CATTLE") ~ "OTHER CROPS FOR HUMAN CONS.",
    str_detect(CROP.SPECIES, "SP.DAIRY.CATTLE") ~ "OTHER CROPS FOR HUMAN CONS.",
    str_detect(CROP.SPECIES, "SP.PIG") ~ "OTHER CROPS FOR HUMAN CONS.",
    str_detect(CROP.SPECIES, "Lolium") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "Vicia") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "Poa") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "Festuca") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "Medicago") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "Panicum") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "Trifolium") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "HAY.FODDER") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "UNSPECIFIED.CLOVERLAY") ~ "FORAGE CROPS",
    str_detect(CROP.SPECIES, "UNSPECIFIED.FALLOW") ~ "FALLOW"
  )) %>%
  # removing 'unreal' species
  filter(!grepl("HAY|UNSPECIFIED|HERBS|SP|NA|NS", CROP.SPECIES)) %>%
  # Generating broad categories of crop species - based on genus
  separate(CROP.SPECIES, c("CROP.GENUS.GROUP", "CROP.SPECIES.GROUP"), "^\\S*\\K\\s+", remove = FALSE)


# ------------------------------------------------------- Merging with crop groups data with biophys database
# merging crops data with the biophys database
crops.biophys.data <-
  full_join(database.biophys, crops,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  )

# ----------------------------------------------------------------------------------------------------------------------
# Saving the combined dataset with biopphysical information unto local disc
readr::write_csv(crops.biophys.data, "./DATASET.FROM.SCRIPT/crops.biophys.data.csv")
# -----------------------------------------------------------------
glimpse(crops.biophys.data)
```

## SAF system Trait types - biophys

Creating the **SAF system Trait types - biophys dataset**

```{r Merging system trait types with biophys database}
# using n_individual_study_sites because system trait types vary between sites, as some studies include multiple sites with different species compositions and arrangements.

systrait.data <-
  database.lat.lon %>%
  filter(INCLUDED == TRUE) %>%
  # renaming SAF System Trait type
  rename("SAF System Trait type" = ST.TYPE) %>%
  # selecting system trait columns
  group_by(`SAF System Trait type`) %>%
  dplyr::select(ID.P, ID.S, `SAF System Trait type`, MANAGEMENT) %>%
  mutate_at(c("ID.P", "ID.S"), as.numeric) # <- converting ID columns from characters to numeric


# ------------------------------------------------------- Merging with crop groups data with biophys database
systrait.biophys.data <-
  full_join(database.biophys, systrait.data,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  )

# ----------------------------------------------------------------------------------------------------------------------
# Saving the combined dataset with biopphysical information unto local disc
readr::write_csv(systrait.biophys.data, "./DATASET.FROM.SCRIPT/systrait.biophys.data.csv")
# -----------------------------------------------------------------
glimpse(systrait.biophys.data)
```

## Global map of study (experimental sites) locations (**see the other script: "2_generating_visualisations"**)

# Geospatial data; multivariate data (matrices)

## ES - SAF systems - pedoclimatic variables
**Ecosystem services (selected ES; including biodiversity + categories) - pedoclimatic variables**

To answer the question; **does SAF, ES research focus differ as related
to location-specific pedoclimatic variables**

Various pedoclimatic variables are **discretised** (from numeric biophys
into categorical groups). Adding grouped intervals based on continuous
soil variables. This is in order to visualise them in relation to ES for
instance.

### Descretising pedoclimatic variables 
```{r}
database.biophys <- vroom::vroom("./DATASET.FROM.SCRIPT/database.biophys.csv")
```

```{r}
biophys.discretised <- vroom::vroom("./DATASET.FROM.SCRIPT/biophys.discretised.csv") 
```

```{r eval=FALSE}
biophys.discretised <- database.biophys %>%
  # renaming SAF System Trait type
  #rename("SAF System Trait type" = ST.TYPE) %>%
  # dividing bioclim temperature data variables by conversion factor to obtain the correct values
  # not dividing SoilGrids variables because; SoilGrids API and maps return values as whole
  # (integer) numbers to minimize the storage space used. These values are converted along the operation to
  # produce conventional units by 'fetchSoilGrids()“
  # conversion factors
  dplyr::mutate(
    bio01_Annual_mean_temp = bio01_Annual_mean_temp / 10,
    bio05_Max_temp_warmest_month = bio05_Max_temp_warmest_month / 10,
    bio04_Temp_seasonality = bio04_Temp_seasonality / 1000,
    #bio05_Max_temp_warmest_month = bio05_Max_temp_warmest_month / 10,
    bio06_Min_temp_coldest_month = bio06_Min_temp_coldest_month / 10,
    bio07_Annual_temp_range = bio07_Annual_temp_range / 10,
    bio10_Mean_temp_warmest_quarter = bio10_Mean_temp_warmest_quarter / 10

  ) %>%
  # discretising using ggplot2::cut_interval(), cut_number() or cut_width()
  mutate(
    # Pedoclimatic variables from BioClim --------------------------------------------------
    bio01_Annual_mean_temp_interval = cut_width(bio01_Annual_mean_temp, 
                                                width = 5, boundary = 0), # 6 levels
    bio03_Isothermality_interval = cut_width(bio03_Isothermality, 
                                             width = 6, boundary = 0), # 6 levels
    bio04_Temp_seasonality_interval = cut_width(bio04_Temp_seasonality, 
                                                width = 2.5, boundary = 0), # 6 levels 
    bio05_Max_temp_warmest_month_interval = cut_width(bio05_Max_temp_warmest_month, 
                                                      width = 4, boundary = 0), # 6 levels 
    bio06_Min_temp_coldest_month_interval = cut_width(bio06_Min_temp_coldest_month, 
                                                      width = 8, boundary = 0), # 6 levels 
    bio07_Annual_temp_range_interval = cut_width(bio07_Annual_temp_range,
                                                 width = 7, boundary = 0), # 6 levels 
    bio10_Mean_temp_warmest_quarter_interval = cut_width(bio10_Mean_temp_warmest_quarter,
                                                         width = 3.5, boundary = 0), # 6 levels
    bio12_Annual_prec_interval = cut_width(bio12_Annual_prec, 
                                           width = 300, boundary = 0), # 6 levels
    bio14_Prec_driest_month_interval = cut_width(bio14_Prec_driest_month,
                                                 width = 20, boundary = 0), # 6 levels
    bio15_Prec_seasonality_interval = cut_width(bio15_Prec_seasonality,
                                                 width = 20, boundary = 0), # 6 levels
    bio17_Prec_driest_quarter_interval = cut_width(bio17_Prec_driest_quarter,
                                                 width = 60, boundary = 0), # 6 levels
    # Pedoclimatic variables from SoilGrids --------------------------------------------------
    clay_interval = cut_width(claymean, width = 10, boundary = 0), # 5 levels
    sand_interval = cut_width(sandmean, width = 20, boundary = 0), # 5 levels
    silt_interval = cut_width(siltmean, width = 15, boundary = 0), # 5 levels
    bd_interval = cut_width(bdodmean, width = 0.25, boundary = 0), # 5 levels
    nitrogen_interval = cut_width(nitrogenmean, width = 3, boundary = 0), # 5 levels
    soc_interval = cut_width(socmean, width = 50, boundary = 0), # 5 levels
    ph_interval = cut_width(phh2omean, width = 1, boundary = 0), # 5 levels
    cec_interval = cut_width(cecmean, width = 10, boundary = 0)
  ) %>% # 5 levels
  relocate(
    ID.P, ID.S,
    `SAF System Trait type`,
    lat, lon,
    # Pedoclimatic variables from BioClim --------------------------------------------------
    bio01_Annual_mean_temp, bio01_Annual_mean_temp_interval,
    bio03_Isothermality, bio03_Isothermality_interval,
    bio04_Temp_seasonality, bio04_Temp_seasonality_interval,
    bio05_Max_temp_warmest_month, bio05_Max_temp_warmest_month_interval,
    bio06_Min_temp_coldest_month, bio06_Min_temp_coldest_month_interval,
    bio07_Annual_temp_range, bio07_Annual_temp_range_interval,
    bio10_Mean_temp_warmest_quarter, bio10_Mean_temp_warmest_quarter_interval,
    bio12_Annual_prec, bio12_Annual_prec_interval,
    bio14_Prec_driest_month, bio14_Prec_driest_month_interval,
    bio15_Prec_seasonality, bio15_Prec_seasonality_interval,
    bio17_Prec_driest_quarter, bio17_Prec_driest_quarter_interval,
    # Pedoclimatic variables from SoilGrids --------------------------------------------------
    claymean, clay_interval,
    sandmean, sand_interval,
    siltmean, silt_interval,
    bdodmean, bd_interval,
    nitrogenmean, nitrogen_interval,
    socmean, soc_interval,
    phh2omean, ph_interval,
    cecmean, cec_interval
  )

# glimpse(biophys.categorised)

# --------------------------------------------------------------------------------- renaming ES variables

biophys.discretised.data <- biophys.discretised %>%
  rename(
    #"SAF System Trait type" = ST.TYPE,
    "Food (yield)" = ES.SUB.PROV.FOOD,
    "Feed" = ES.SUB.PROV.FEED,
    "Fibre" = ES.SUB.PROV.FIBRE,
    "Fuel (firewood etc.)" = ES.SUB.PROV.FUEL,
    "Genetic resources" = ES.SUB.PROV.GENRES,
    "Biochemicals" = ES.SUB.PROV.BIOCHEM,
    "Ornamentals" = ES.SUB.PROV.ORNA,
    "Fresh water" = ES.SUB.PROV.WATFRESH,
    "Air regulation" = ES.SUB.REGU.AIR,
    "Climate regulation" = ES.SUB.REGU.CLIMREG,
    "Carbon sequestration" = ES.SUB.REGU.CSEQREG,
    "Water regulation" = ES.SUB.REGU.WATREG,
    "Nutrient leaching regulation" = ES.SUB.REGU.NUTRILE,
    "Erosion regulation" = ES.SUB.REGU.EROREG,
    "Water purification" = ES.SUB.REGU.WATPUR,
    "Disease regulation" = ES.SUB.REGU.DISREG,
    "Pest regulation" = ES.SUB.REGU.PESTREG,
    "Polination regulation" = ES.SUB.REGU.POLREG,
    "Natural hazards regulation" = ES.SUB.REGU.NATHAZREG,
    "Soil formation" = ES.SUB.SUPP.SOILFORM,
    "Photosynthesis" = ES.SUB.SUPP.PHOTSYNT,
    "Primary production" = ES.SUB.SUPP.PRIMPROD,
    "Nutrient cycling" = ES.SUB.SUPP.NUTRICYC,
    "Water cycling" = ES.SUB.SUPP.WATCYC,
    "Cultural diversity" = ES.SUB.CULT.CULTDIV,
    "Spiritual value" = ES.SUB.CULT.SPIRITVAL,
    "Knowledge systems" = ES.SUB.CULT.KNOWLSYS,
    "Educational value" = ES.SUB.CULT.EDUVAL,
    "Inspirational value" = ES.SUB.CULT.INSPIR,
    "Aesthetic value" = ES.SUB.CULT.AESTVAL,
    "Social relations" = ES.SUB.CULT.SOCREL,
    "Sense of place" = ES.SUB.CULT.SENPLACE,
    "Cultural heritage value" = ES.SUB.CULT.CULTHERIVAL,
    "Recreational and ecotourism" = ES.SUB.CULT.RECRECOTOUR,
    "Provisioning ES" = ES.PROV,
    "Regulating ES" = ES.REGU,
    "Supporting ES" = ES.SUPP,
    "Cultural ES" = ES.CULT,
    "Ecosystem Services" = ES,
    # ----------------------------------------------------------------------------------- renaming biodiversity variables
    "Associated biodiversity" = BD, # Biodiversity -associated biodiversity (from Beillouin et al. 2021)
    "FLORA.1" = BD.SUB.FLORAMACRO,
    "FLORA.2" = BD.SUB.FLORAMICRO,
    "FAUNA.1" = BD.SUB.FAUNAMACRO,
    "FAUNA.2" = BD.SUB.FAUNAMICRO,
    "Bacteria" = BD.SUB.BACTERIA,
    "Fungi" = BD.SUB.FUNGI,
    # -------------------------------------------------------------------------- renaming categories of approach variables
    "Food" = CAT.SUB.AGRO.YIELD,
    "Biomass" = CAT.SUB.AGRO.BIOMASS,
    "Land equivalent ratio" = CAT.SUB.AGRO.LER,
    "Pests" = CAT.SUB.AGRO.PESTS,
    "Diseases" = CAT.SUB.AGRO.DISEASE,
    "Weeds" = CAT.SUB.AGRO.WEEDS,
    "Product quality" = CAT.SUB.AGRO.PRODQUAL,
    "Production stability" = CAT.SUB.AGRO.PRODSTAB,     # Product stability
    "Mechanisation" = CAT.SUB.AGRO.MECHAN,
    "Belowground competition" = CAT.SUB.AGRO.CRTRBLOW,  # Belowground (crop-tree) competition
    "Aboveground competition" = CAT.SUB.AGRO.LIGHTCMP,  # Aboveground (crop-tree) light competition
    "Nutrient use" = CAT.SUB.AGRO.NUTRIUSE,
    "Nutrient use effeciency" = CAT.SUB.AGRO.NUTRIEF,
    "Breeding and selection" = CAT.SUB.AGRO.BRDSELECT,
    "Soil quality" = CAT.SUB.ENVI.SOILQUAL,
    "Biodiversity" = CAT.SUB.ENVI.BIODIV, # Biodiversity -associated biodiversity (from Beillouin et al. 2021)
    "Carbon stock" = CAT.SUB.ENVI.CSTOCK,
    "Greenhouse gas emissions" = CAT.SUB.ENVI.GHGEMS,
    "Pollination" = CAT.SUB.ENVI.POLLCON,
    "Soil structure" = CAT.SUB.ENVI.SOILSTRUCT,
    "Chemical soil components" = CAT.SUB.ENVI.SOILCHEM,
    "Biological soil components" = CAT.SUB.ENVI.SOILBIO,
    "Soil water" = CAT.SUB.ENVI.SOILWAT,
    "Soil erosion" = CAT.SUB.ENVI.SOILEROCO,
    "Soil formations" = CAT.SUB.ENVI.SOILFORM,
    "Water quality" = CAT.SUB.ENVI.WATQUAL,
    "Water perculation" = CAT.SUB.ENVI.WATPERC,
    "Water cyclings" = CAT.SUB.ENVI.WATCYC,
    "Water availability" = CAT.SUB.ENVI.WATAVAIL,
    "Hydraulic lift" = CAT.SUB.ENVI.HYDRAULIFT,
    "Profitability" = CAT.SUB.ECON.PROFIT,       # Economic profit related
    "Management" = CAT.SUB.ECON.MANAGE,          # Economic management related
    "Market" = CAT.SUB.ECON.MARKET,              # Economic market related
    "Landscape aesthetics" = CAT.SUB.SOCUL.LANAES,
    "Recreational and ecotourisms" = CAT.SUB.SOCUL.RECRECOTOUR,
    "Spiritual values" = CAT.SUB.SOCUL.SPIRITVAL,
    "Perception of adoptation" = CAT.SUB.SOCUL.PERCEPADOPT,
    "Policy information" = CAT.SUB.SOCUL.POLICYINFR,
    "Agronomic category" = CAT.AGRO,
    "Environmental category" = CAT.ENVIRON,
    "Economic category" = CAT.ECONOM,
    "Socio-cultural category" = CAT.SOCIOCUL 
  ) %>%
  # converting flora + fauna to numeric
  mutate_at(c("FLORA.1", "FLORA.2", "FAUNA.1", "FAUNA.2"), as.numeric) %>%
  # ------------------------------------------------------------------------------------------------------------------------------
  dplyr::mutate(
    Flora = FLORA.1 + FLORA.2, # simplifying flora, fauna groups
    Fauna = FAUNA.1 + FAUNA.2
  ) %>%
  mutate(`SAF System Trait type` = recode(`SAF System Trait type`, # rename dataframe rows
    RBS = "Riparian buffer strips",
    SHB = "Line belts *",
    MIXED = "Mixed systems",
    TIMBER.ARABLE = "Timber - Arable",
    BIOMASS.ARABLE = "Biomass - Arable",
    NUT.ARABLE = "Nut - Arable",
    FRUIT.ARABLE = "Fruit - Arable",
    .keep = "all"
  )) %>%
  dplyr::mutate(ST.TYPE.PROD.TYPE = "x") %>%
  mutate(ST.TYPE.PROD.TYPE = recode(`SAF System Trait type`, # rename dataframe rows
    TIMBER.ARABLE = "Biomass",
    BIOMASS.ARABLE = "Biomass",
    NUT.ARABLE = "Food",
    FRUIT.ARABLE = "Food",
    RBS = "Multiple",
    SHB = "Multiple",
    MIXED = "Multiple",
    .keep = "all"
  ))

# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(biophys.discretised.data, "./DATASET.FROM.SCRIPT/biophys.discretised.data.csv")
# ------------------------------------------------------------------------------
glimpse(biophys.discretised.data)
```

To create (1) broad focus groups of ecosystem services as well as (2)
selected ecosystem services we use information from the *data
dimensions*; **ecosystem services**, **biodiversity** and **categories
of approach**. Following the style of *Ditzler et al. (2021)*, *Duru et
al. (2015)* and *Tamburini et al. (2020)*

**Loading dataset**
```{r }

biophys.discretised.data <- vroom::vroom("./DATASET.FROM.SCRIPT/biophys.discretised.data.csv") 
```

### Defining selected soil-related ES for soil variables

*Selected soil-related ES for plot*
```{r}
# ------------------------------------------------------------------------------------------------------------
A_selected_soil_ES_soil <- biophys.discretised.data %>%
  dplyr::select(
    # ID --------------------------
    "ID.S",
    "ID.P",
    # --------------------------------------------------------------------
#--# Biodiversity ---------------------------------------------
    "Associated biodiversity",
#--# Supporting ecosystem services ----------------------------
    "Soil formation",
    "Soil structure",
    "Water cycling",
    "Nutrient cycling",
    "Carbon stock",
    "Chemical soil components",
    "Biological soil components",
        # "Greenhouse gas emissions",
#--# Regulating ecosystem services ----------------------------
    "Erosion regulation",
    "Carbon sequestration",
    "Water regulation",
    "Water perculation",
        # "Water quality",
        # "Climate regulation",
        # "Pest regulation",
        # "Disease regulation",
        # "Polination regulation",
#--# Provisioning ecosystem services --------------------------
    "Water availability",
    "Production stability",
    "Feed",
    "Fuel (firewood etc.)",
    "Food (yield)",
        # "Land equivalent ratio",
        # "Aboveground competition",
        # "Management",
        # "Profitability",
#--# Cultural ecosystem services ------------------------------
        # "Recreational and ecotourism",
        # "Landscape aesthetics",
        # "Policy information",
    
#--# Specific pedoclimatic, soil-related ecosystem services ---- following Ellili-Bargaoui et al. (2021)
        # "Nutrient cycling",
        # "Erosion regulation",
        # "Carbon stock",
        # "Chemical soil components",
        # "Biological soil components",
        # "Water perculation",
        # "Water availability"
  ) %>%
  mutate_if(is.character, as.factor) %>%
  relocate(ID.P, ID.S) %>%
  pivot_longer(
    cols = c(-ID.P, -ID.S),
    names_to = "ES.GROUPS",
    values_to = "count"
  ) %>%
  mutate(count = as.numeric(count)) %>%
  distinct() %>%
  filter(count == 1)

# --------------------------------------------------------------------------------------------------------------------------
B_SAFTT_data <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "SAF System Trait type"
  ) %>%
  distinct()

# --------------------------------------------------------------------------------------------------------------------------
D_JOINED_A_B_selected_soil_ES_soil_SAFTT <-
  left_join(
    x = A_selected_soil_ES_soil, y = B_SAFTT_data,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`
  )
```


### Defining selected soil-related ES for bioclim variables

*Selected soil-related ES for plot*
```{r}
# ------------------------------------------------------------------------------------------------------------
A_selected_soil_ES_bioclim <- biophys.discretised.data %>%
  dplyr::select(
    # ID --------------------------
    "ID.S",
    "ID.P",
    # --------------------------------------------------------------------
#--# Biodiversity ---------------------------------------------
    "Associated biodiversity",
#--# Supporting ecosystem services ----------------------------
    "Soil formation",
    "Soil structure",
    "Water cycling",
    "Nutrient cycling",
    "Carbon stock",
        # "Chemical soil components",
        # "Biological soil components",
        # "Greenhouse gas emissions",
#--# Regulating ecosystem services ----------------------------
    "Erosion regulation",
    "Carbon sequestration",
    "Water regulation",
    "Water perculation",
    "Climate regulation",    
    "Pest regulation",
    "Disease regulation",
    "Polination regulation",
        # "Water quality",
#--# Provisioning ecosystem services --------------------------
    "Profitability",
    "Aboveground competition",
    "Water availability",
    "Production stability",
    "Feed",
    "Fuel (firewood etc.)",
    "Food (yield)",
        # "Land equivalent ratio",
        # "Management",
#--# Cultural ecosystem services ------------------------------
    "Recreational and ecotourism",
    "Landscape aesthetics",
    "Policy information",
    
#--# Specific pedoclimatic, soil-related ecosystem services ---- following Ellili-Bargaoui et al. (2021)
        # "Nutrient cycling",
        # "Erosion regulation",
        # "Carbon stock",
        # "Chemical soil components",
        # "Biological soil components",
        # "Water perculation",
        # "Water availability"
  ) %>%
  mutate_if(is.character, as.factor) %>%
  relocate(ID.P, ID.S) %>%
  pivot_longer(
    cols = c(-ID.P, -ID.S),
    names_to = "ES.GROUPS",
    values_to = "count"
  ) %>%
  mutate(count = as.numeric(count)) %>%
  distinct() %>%
  filter(count == 1)

# --------------------------------------------------------------------------------------------------------------------------
B_SAFTT_data <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "SAF System Trait type"
  ) %>%
  distinct()

# --------------------------------------------------------------------------------------------------------------------------
D_JOINED_A_B_selected_soil_ES_bioclim_SAFTT <-
  left_join(
    x = A_selected_soil_ES_bioclim, y = B_SAFTT_data,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`
  )
```


### Defining general selected ES

```{r}
# ------------------------------------------------------------------------------------------------------------------------
# Preparing ES dataset for merging by converting it to long format and merging it with SAF system traits types
A_ES <- biophys.discretised.data %>%
  dplyr::select(
    # ID --------------------------
    "ID.S",
    "ID.P",
    # --------------------------------------------------------------------
    # Biodiversity ---------------------------------------------
    "Associated biodiversity",
    # Supporting ecosystem services ----------------------------
    "Soil formation",
    "Soil structure",
    "Water cycling",
    "Nutrient cycling",
    "Greenhouse gas emissions",
    # Regulating ecosystem services ----------------------------
    "Carbon sequestration",
    "Climate regulation",
    "Water regulation",
    "Water quality",
    "Pest regulation",
    "Disease regulation",
    "Polination regulation",
    # Provisioning ecosystem services --------------------------
    "Land equivalent ratio",
    "Aboveground competition",
    "Management",
    "Profitability", 
    "Production stability",
    "Feed",
    "Fuel (firewood etc.)",
    "Food (yield)",
    # Cultural ecosystem services ------------------------------
    "Recreational and ecotourism",
    "Landscape aesthetics",
    "Policy information",
    # Specific pedoclimatic, soil-related ecosystem services ---- following Ellili-Bargaoui et al. (2021)
    "Nutrient cycling",
    "Erosion regulation",
    "Carbon stock", 
    "Chemical soil components",
    "Biological soil components",
    "Water perculation",
    "Water availability"
  ) %>%
  mutate_if(is.character, as.factor) %>%
  relocate(ID.P, ID.S) %>%
  pivot_longer(
    cols = c(-ID.P, -ID.S),
    names_to = "ES.GROUPS",
    values_to = "count"
  ) %>%
  mutate(count = as.numeric(count)) %>%
  distinct() %>%
  filter(count == 1)

# --------------------------------------------------------------------------------------------------------------------------
B_SAFTT <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "SAF System Trait type"
  ) %>%
  distinct()

# --------------------------------------------------------------------------------------------------------------------------
D_JOINED_A_B_ES_SAFTT <-
  left_join(
    x = A_ES, y = B_SAFTT,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`
  )
```


## ES - sand content
**Merging selected ecosystem services data with biophys variables; soil sand content**
```{r}
# ------------------------------------------------------------------------------------------------------------------------------
# Merging with sand content data
C_SAND <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "sand_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  dplyr::mutate(sand_interval = recode(sand_interval, # renaming dataframe rows
    "[0,20]" = "0 - 20 %",
    "(20,40]" = "20 - 40 %",
    "(40,60]" = "40 - 60 %",
    "(60,80]" = "60 - 80 %",
    "(80,100]" = "80 - 100 %",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with soil sand content
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_SAND <-
  left_join(
    x = D_JOINED_A_B_ES_SAFTT, y = C_SAND,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, sand_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  distinct()

es.systrait.sand.data <- E_JOINED_D_C_SAND
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.sand.data, "./DATASET.FROM.SCRIPT/es.systrait.sand.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.sand.data)
```




## ES - soil organic carbon *

**Merging selected soil-related ecosystem services data with biophys variables; SOC content**
```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_SOC_data <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "soc_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(soc_interval = recode(soc_interval, # renaming dataframe rows
    "[0,50]" = "0 - 50 g/kg",
    "(50,100]" = "50 - 100 g/kg",
    "(100,150]" = "100 - 150 g/kg",
    "(150,200]" = "150 - 200 g/kg",
    "(200,250]" = "200 - 250 g/kg",
    "(250,300]"  = "250 - 300 g/kg",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with soil SOC content
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_soil_selected_ES_SOC <-
  left_join(
    x = D_JOINED_A_B_selected_soil_ES_soil_SAFTT, y = C_SOC_data,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, soc_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of SOC intervals
  mutate_at(vars(soc_interval), funs(factor(., levels = unique(.)))) 

selected.soil.es.systrait.soc.data <- E_JOINED_soil_selected_ES_SOC
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(selected.soil.es.systrait.soc.data, "./DATASET.FROM.SCRIPT/selected.soil.es.systrait.soc.data.csv")
# ------------------------------------------------------------------------------
glimpse(selected.soil.es.systrait.soc.data)
```

**Merging selected ecosystem services data with biophys variables; SOC content**
```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_SOC <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "soc_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(soc_interval = recode(soc_interval, # renaming dataframe rows
    "[0,50]" = "0 - 50 g/kg",
    "(50,100]" = "50 - 100 g/kg",
    "(100,150]" = "100 - 150 g/kg",
    "(150,200]" = "150 - 200 g/kg",
    "(200,250]" = "200 - 250 g/kg",
    "(250,300]"  = "250 - 300 g/kg",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with soil SOC content
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_SOC <-
  left_join(
    x = D_JOINED_A_B_ES_SAFTT, y = C_SOC,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, soc_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of SOC intervals
  mutate_at(vars(soc_interval), funs(factor(., levels = unique(.)))) 

es.systrait.soc.data <- E_JOINED_D_C_SOC
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.soc.data, "./DATASET.FROM.SCRIPT/es.systrait.soc.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.soc.data)
```

## ES - soil bulk density *

**Merging selected soil-related ecosystem services data with biophys variables; soil bulk density**
```{r}
# --------------------------------------------------------------------------------------------------------------------

C_BULK_data <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bd_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bd_interval = recode(bd_interval, # renaming dataframe rows
    "[0.75,1]"   =   "0.75 - 1 [kg dm^-3]",
    "(1,1.25]"   =   "1 - 1.25 [kg dm^-3]",
    "(1.25,1.5]" =   "1.25 - 1.5 [kg dm^-3]",
    "(1.5,1.75]" =   "1.5 - 1.75 [kg dm^-3]",
    "(1.75,2]"   =   "1.75 - 2 [kg dm^-3]",
    .keep = "all"
  ))

# -------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with soil SOC content
# --------------------------------------------------------------------------------------------------------------------
E_JOINED_soil_selected_ES_BULK <-
  left_join(
    x = D_JOINED_A_B_selected_soil_ES_soil_SAFTT, y = C_BULK_data,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bd_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of SOC intervals
  mutate_at(vars(bd_interval), funs(factor(., levels = unique(.)))) 

selected.soil.es.systrait.bd.data <- E_JOINED_soil_selected_ES_BULK
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(selected.soil.es.systrait.bd.data,
                 "./DATASET.FROM.SCRIPT/selected.soil.es.systrait.bd.data.csv")
# ------------------------------------------------------------------------------
glimpse(selected.soil.es.systrait.bd.data)
```


**Merging selected ecosystem services data with biophys variables; soil bulk density**
```{r}
# --------------------------------------------------------------------------------------------------------------------

C_BULK <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bd_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bd_interval = recode(bd_interval, # renaming dataframe rows
    "[0.75,1]"   =   "0.75 - 1 [kg dm^-3]",
    "(1,1.25]"   =   "1 - 1.25 [kg dm^-3]",
    "(1.25,1.5]" =   "1.25 - 1.5 [kg dm^-3]",
    "(1.5,1.75]" =   "1.5 - 1.75 [kg dm^-3]",
    "(1.75,2]"   =   "1.75 - 2 [kg dm^-3]",
    .keep = "all"
  ))

# -------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with soil SOC content
# --------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_BULK <-
  left_join(
    x = D_JOINED_A_B_ES_SAFTT, y = C_BULK,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bd_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of SOC intervals
  mutate_at(vars(bd_interval), funs(factor(., levels = unique(.)))) 

es.systrait.bd.data <- E_JOINED_D_C_BULK
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.bd.data, "./DATASET.FROM.SCRIPT/es.systrait.bd.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.bd.data)
```

## ES - soil pH

**Merging selected ecosystem services data with biophys variables; soil pH**
```{r}
# --------------------------------------------------------------------------------------------------------------------

C_PH <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "ph_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(ph_interval = recode(ph_interval, # renaming dataframe rows
    "[4,5]"   =   "pH 4 - 5",
    "(5,6]"   =   "pH 5 - 6",
    "(6,7]"   =   "pH 6 - 7",
    "(7,8]"   =   "pH 7 - 8",
    "(8,9]"   =   "pH 8 - 9",
    .keep = "all"
  ))

# -------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with soil pH content
# --------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_PH <-
  left_join(
    x = D_JOINED_A_B_ES_SAFTT, y = C_PH,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, ph_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of pH intervals
  mutate_at(vars(ph_interval), funs(factor(., levels = unique(.)))) 

es.systrait.ph.data <- E_JOINED_D_C_PH
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.ph.data, "./DATASET.FROM.SCRIPT/es.systrait.ph.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.ph.data)
```



## ES - isothermality

**Merging selected ecosystem services data with biophys variables;
isothermality** Isothermality is defined as; Isothermality quantifies
how large the day-to-night temperatures oscillate relative to the
summer-to-winter (annual) oscillations. Isothermality is generally
useful for biological entities (e.g. species and organisms). Units:
Percent. Isothermality quantifies how large the day-to-night
temperatures oscillate relative to the summer-to-winter (annual)
oscillations. An isothermal value of 100 indicates the diurnal
temperature range is equivalent to the annual temperature range, while
anything less than 100 indicates a smaller level of temperature
variability within an average month relative to the year. A species
distribution may be influenced by larger or smaller temperature
fluctuations within a month relative to the year and this predictor is
useful for ascertaining such information.

```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_ISOT <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bio03_Isothermality_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bio03_Isothermality_interval = recode(bio03_Isothermality_interval, # rename dataframe rows
    "[12,18]" = "12 - 18 [%]",
    "(18,24]" = "18 - 24 [%]",
    "(24,30]"  = "24 - 30 [%]",
    "(30,36]" = "30 - 36 [%]",
    "(36,42]" = "36 - 42 [%]",
    "(42,48]" = "42 - 48 [%]",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with temperature
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_ISOTs <-
  left_join(
    x = D_JOINED_A_B_ES_SAFTT, y = C_ISOT,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bio03_Isothermality_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of temperature intervals
  mutate_at(vars(bio03_Isothermality_interval), funs(factor(., levels = unique(.))))

es.systrait.isot.data <- E_JOINED_D_C_ISOTs
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.isot.data, "./DATASET.FROM.SCRIPT/es.systrait.isot.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.isot.data)
```

## ES - temperature seasonality

**Merging selected ecosystem services data with biophys variables;
temperature seasonality** Temperature seasonality is defined as; the
amount of temperature variation over a given year (or averaged years)
based on the standard deviation (variation) of monthly temperature
averages. Units: Temperature (degrees Celsius)

```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_TEMP_s <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bio04_Temp_seasonality_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bio04_Temp_seasonality_interval = recode(bio04_Temp_seasonality_interval, # rename dataframe rows
    "[2.5,5]" = "2.5 - 5 [°C]",
    "(5,7.5]" = "5 - 7.5 [°C]",
    "(7.5,10]"  = "7.5 - 10 [°C]",
    "(10,12.5]" = "10 - 12.5 [°C]",
    "(12.5,15]" = "12.5 - 15 [°C]",
    "(15,17.5]" = "15 - 17.5 [°C]",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with temperature
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_TEMP_s <-
  left_join(
    x = D_JOINED_A_B_ES_SAFTT, y = C_TEMP_s,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bio04_Temp_seasonality_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of temperature intervals
  mutate_at(vars(bio04_Temp_seasonality_interval), funs(factor(., levels = unique(.))))

es.systrait.temp.seasonality.data <- E_JOINED_D_C_TEMP_s
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.temp.seasonality.data, "./DATASET.FROM.SCRIPT/es.systrait.temp.seasonality.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.temp.seasonality.data)
```

## ES - precipitation seasonality

Precipitation seasonality is defined as; a
measure of the variation in monthly precipitation totals over the course
of the year. This index is the ratio of the standard deviation of the
monthly total precipitation to the mean monthly total precipitation
(also known as the coefficient of variation) and is expressed as a
percentage. Units: Percent

**Merging selected ecosystem services data with biophys variables; precipitation seasonality** 
```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_PREC_s <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bio15_Prec_seasonality_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bio15_Prec_seasonality_interval = recode(bio15_Prec_seasonality_interval, # rename dataframe rows
    "[0,20]" = "0 - 20 [%]",
    "(20,40]" = "20 - 40 [%]",
    "(40,60]"  = "40 - 60 [%]",
    "(60,80]" = "60 - 80 [%]",
    "(80,100]" = "80 - 100 [%]",
    "(100,120]" = "100 - 120 [%]",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with precipitation
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_PREC_s <-
  left_join(
    x = D_JOINED_A_B_ES_SAFTT, y = C_PREC_s,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bio15_Prec_seasonality_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of precipitation intervals
  mutate_at(vars(bio15_Prec_seasonality_interval), funs(factor(., levels = unique(.))))

es.systrait.prec.seasonality.data <- E_JOINED_D_C_PREC_s
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.prec.seasonality.data, "./DATASET.FROM.SCRIPT/es.systrait.prec.seasonality.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.prec.seasonality.data)
```

## ES - annual mean temperature

**Merging selected ecosystem services data with biophys variables; temperature**
```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_TEMP <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bio01_Annual_mean_temp_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bio01_Annual_mean_temp_interval = recode(bio01_Annual_mean_temp_interval, # rename dataframe rows
    "[-10,-5]" = "-10 - -5 [°C]",
    "(-5,0]" = "-5 - 0 [°C]",
    "(0,5]"  = "0 - 5 [°C]",
    "(5,10]" = "5 - 10 [°C]",
    "(10,15]" = "10 - 15 [°C]",
    "(15,20]" = "15 - 20 [°C]",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with temperature
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_TEMP <-
  left_join(
    x = D_JOINED_A_B_ES_SAFTT, y = C_TEMP,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bio01_Annual_mean_temp_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of temperature intervals
  mutate_at(vars(bio01_Annual_mean_temp_interval), funs(factor(., levels = unique(.))))

es.systrait.temp.data <- E_JOINED_D_C_TEMP
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.temp.data, "./DATASET.FROM.SCRIPT/es.systrait.temp.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.temp.data)
```

## ES - annual precipitation

**Merging selected ecosystem services data with biophys variables; precipitation**
```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_PREC <- biophys.discretised.data %>%
    dplyr::select(
    "ID.P",
    "ID.S",
    
    "bio12_Annual_prec_interval") %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bio12_Annual_prec_interval = recode(bio12_Annual_prec_interval,               # rename dataframe rows
                                "[0,300]" = "0 - 300 [mm]",
                                "(300,600]" = "300 - 600 [mm]",
                                "(600,900]"  = "600 - 900 [mm]",
                                "(900,1.2e+03]" = "900 - 1200 [mm]",
                                "(1.2e+03,1.5e+03]" = "1200 - 1500 [mm]",
                                "(1.5e+03,1.8e+03]" = "1500 - 1800 [mm]",
                                .keep = "all"))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with precipitation
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_PREC <- 
  left_join(x = D_JOINED_A_B_ES_SAFTT, y = C_PREC,
            by = c("ID.P" = "ID.P",
                   "ID.S" = "ID.S")) %>%
  distinct() %>%
  relocate(ID.P, ID.S,
           ES.GROUPS, `SAF System Trait type`, bio12_Annual_prec_interval) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of precipitation interval levels
  mutate_at(vars(bio12_Annual_prec_interval), funs(factor(., levels=unique(.)))) 

es.systrait.prec.data <- E_JOINED_D_C_PREC
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.prec.data, "./DATASET.FROM.SCRIPT/es.systrait.prec.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.prec.data)
```

## ES - temperature extremes (bio10) *

Bio10 = mean temperature warmest quarter

**Merging selected ecosystem services data with biophys variables; temperature extremes**
```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_TEMP_bio10 <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bio10_Mean_temp_warmest_quarter_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bio10_Mean_temp_warmest_quarter_interval = recode(bio10_Mean_temp_warmest_quarter_interval, 
    "[7,10.5]" = "7 - 10.5 [°C]",
    "(10.5,14]" = "10.5 - 14 [°C]",
    "(14,17.5]"  = "14 - 17.5 [°C]",
    "(17.5,21]" = "17.5 - 21 [°C]",
    "(21,24.5]" = "21 - 24.5 [°C]",
    "(24.5,28]" = "24.5 - 28 [°C]",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with temperature
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_TEMP_bio10 <-
  left_join(
    x = D_JOINED_A_B_selected_soil_ES_bioclim_SAFTT, y = C_TEMP_bio10,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bio10_Mean_temp_warmest_quarter_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of temperature intervals
  mutate_at(vars(bio10_Mean_temp_warmest_quarter_interval), funs(factor(., levels = unique(.))))

es.systrait.temp.bio10.data <- E_JOINED_D_C_TEMP_bio10
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.temp.bio10.data, "./DATASET.FROM.SCRIPT/es.systrait.temp.bio10.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.temp.bio10.data)
```


**Merging selected ecosystem services data with biophys variables; temperature extremes**
```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_TEMP_bio10 <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bio10_Mean_temp_warmest_quarter_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bio10_Mean_temp_warmest_quarter_interval = recode(bio10_Mean_temp_warmest_quarter_interval, 
    "[7,10.5]" = "7 - 10.5 [°C]",
    "(10.5,14]" = "10.5 - 14 [°C]",
    "(14,17.5]"  = "14 - 17.5 [°C]",
    "(17.5,21]" = "17.5 - 21 [°C]",
    "(21,24.5]" = "21 - 24.5 [°C]",
    "(24.5,28]" = "24.5 - 28 [°C]",
    .keep = "all"
  ))

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with temperature
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_TEMP_bio10 <-
  left_join(
    x = D_JOINED_A_B_selected_soil_ES_bioclim_SAFTT, y = C_TEMP_bio10,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bio10_Mean_temp_warmest_quarter_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of temperature intervals
  mutate_at(vars(bio10_Mean_temp_warmest_quarter_interval), funs(factor(., levels = unique(.))))

es.systrait.temp.bio10.data <- E_JOINED_D_C_TEMP_bio10
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.temp.bio10.data, "./DATASET.FROM.SCRIPT/es.systrait.temp.bio10.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.temp.bio10.data)
```

## ES - precipitation/draught extremes (bio17) *

Bio17 = precipitation driest quarter

**Merging selected ecosystem services data with biophys variables; precipitation extremes**
```{r warning=FALSE, message=FALSE}
# --------------------------------------------------------------------------------------------------------------------------

C_PREC_bio17 <- biophys.discretised.data %>%
  dplyr::select(
    "ID.P",
    "ID.S",
    "bio17_Prec_driest_quarter_interval"
  ) %>%
  # to avoid exhaustion of R memory
  distinct() %>%
  mutate(bio17_Prec_driest_quarter_interval = recode(bio17_Prec_driest_quarter_interval, 
    "[0,60]" = "0 - 60 [mm]",
    "(60,120]" = "60 - 120 [mm]",
    "(120,180]"  = "120 - 180 [mm]",
    "(180,240]" = "180 - 240 [mm]",
    "(240,300]" = "240 - 300 [mm]", # does not exist, so a dummy row is added further down to have 6 intervals
    "(300,360]" = "300 - 360 [mm]",
    .keep = "all"
  )) 

# --------------------------------------------------------------------------------------------------------------------------
# Final dataset of selected ecosystem services (including biodiversity and categories data) with temperature
# --------------------------------------------------------------------------------------------------------------------------
E_JOINED_D_C_PREC_bio17 <-
  left_join(
    x = D_JOINED_A_B_selected_soil_ES_bioclim_SAFTT, y = C_PREC_bio17,
    by = c(
      "ID.P" = "ID.P",
      "ID.S" = "ID.S"
    )
  ) %>%
  distinct() %>%
  relocate(
    ID.P, ID.S,
    ES.GROUPS, `SAF System Trait type`, bio17_Prec_driest_quarter_interval
  ) %>%
  mutate_if(is.character, as.factor) %>%
  # convert to factor, sort the facet plots in order of temperature intervals
  mutate_at(vars(bio17_Prec_driest_quarter_interval), funs(factor(., levels = unique(.)))) %>%   
# Adding dummy row for the range (240,300] because there is no such incedence and it looks ugly in the plot
  add_row(bio17_Prec_driest_quarter_interval = "240 - 300 [mm]",
          `SAF System Trait type` = "Timber - Arable", 
          ES.GROUPS = "Food (yield)",
          ID.P = 3,
          ID.S = 1,
          count = 1)

es.systrait.prec.bio17.data <- E_JOINED_D_C_PREC_bio17 
# ------------------------------------------------------------------------------------------------------------------------------
# Saving the dataset
readr::write_csv(es.systrait.prec.bio17.data, "./DATASET.FROM.SCRIPT/es.systrait.prec.bio17.data.csv")
# ------------------------------------------------------------------------------
glimpse(es.systrait.prec.bio17.data)
```






















# Projected future climates and relation to ES research for silvoarable agroforestry in Europe

## Get current climate data

```{r}
file_path <- paste0(dirname(here::here()), "/silvoarable_review/DATASET.FROM.SCRIPT/")

# geodata::worldclim_global(var = 'bio', res = 2.5, path = file_path)

raster::getData(name = 'worldclim', 
                var = 'bio', 
                # resolution 10 in order to get the right data types for processing
                res = 10, 
                path = file_path)
```

## Get projected future climate data

```{r}
raster::getData(
  # data from the CMIP5 models - resolution 10 in order to get the right data types for processing
  name = "CMIP5", var = "bio", res = 10, 
  # RCP 4.5 assumes a peak decline in GHG emissions around 2040, followed by decreasing emission
  rcp = 45, 
  # single model results of the IPSL-CM5A-LR’ model
  model = "IP",
  # get projections for the period 2061-2080 (alternative is 50).
  year = 70, 
  path = file_path
)
```

## Load and process isothermality data

```{r}
iso_T <- stars::read_stars(paste0(file_path, "wc10/bio3.bil"))

iso_T_70 <- stars::read_stars(paste0(file_path, "cmip5/10m/ip45bi703.tif"))
```

## Plotting Europe map

## Global map of study (experimental sites) locations

Creating and plotting the **global map**

```{r}
world <- map_data("world") %>%
  filter(region != "Antarctica") # get the world map coordinates using map_data() function from the tidyverse suit of packages



world.study.sites <-
  ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "white", fill = "lightgray", size = 0.1
  ) +
  geom_point(
    data = database.lat.lon,
    aes(longitude, latitude, color = ST.TYPE),
    alpha = 1, size = 0.2
  ) +
  geom_rect(
    aes(xmin = -15, xmax = 35, ymin = 35.5, ymax = 59),
    color = "black", fill = NA, inherit.aes = FALSE
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 1), legend.position = "top",
    plot.margin = margin(t = 0, r = 5, b = 0, l = 5, unit = "pt")
  ) +
  # scale_colour_discrete_qualitative(palette = "Dynamic", nmax = 8) +
  guides(color = guide_legend(override.aes = list(size = 1.5))) +
  coord_sf(ylim = c(20, 70), xlim = c(-140, 140)) +
  labs( # title = "Locations of study sites for peer-reviewed silvoarable agroforestry",
    # subtitle = paste("n (individual study sites) = ", n_individual_study_sites),
    # caption = paste0("Total number of independent studies = ", n_independent_studies),
    x = "Longitude",
    y = "Latitude"
  )
# ggtitle("A")

world.study.sites
```

Creating map that is **zoomed on Europe** Now what if we wanted to use
the *Mollweide projection* again as target projection

```{r}
# At first, we set the target CRS and transform the whole worldmap dataset to that CRS.
worldmap <- ne_countries(
  scale = "medium", type = "map_units",
  returnclass = "sf"
)
target_crs <- "+proj=moll"
worldmap_trans <- st_transform(worldmap, crs = target_crs)

# Next, we specify the display window in WGS84 coordinates as longitude / latitude degrees.

disp_win_wgs84 <- st_sfc(st_point(c(-20, 30)), st_point(c(45, 73)),
  crs = 4326
)

# These coordinates can be transformed to our target CRS via st_transform.

disp_win_trans <- st_transform(disp_win_wgs84, crs = target_crs)

# We can extract the coordinates from these points and pass them as limits for the x and y scale respectively. Note also that I set datum = target_crs which makes sure that the graticule is displayed in the target CRS. Otherwise ggplot2 by default displays a WGS84 graticule.

disp_win_coord <- st_coordinates(disp_win_trans)

 Europe_map <- 
   ggplot() +
  geom_sf(data = worldmap_trans) +
  coord_sf(
    xlim = disp_win_coord[, "X"], ylim = disp_win_coord[, "Y"],
    datum = target_crs, expand = FALSE
  ) +
  theme_bw()
```

```{r}
south_america_map <- rnaturalearth::ne_countries(continent = "south america", returnclass = "sf")
# The precision has to be set to a value > 0 to resolve internal boundaries.
st_precision(south_america_map) <- 1e9 # Required to
south_america_map <- st_union(south_america_map)

{
par(mar = c(0,0,0,0))
plot(south_america_map)
}
```

## Get map of Europe

```{r}

world_map <- ne_countries(scale = "medium", returnclass = "sf")
Europe_map <- world[which(world$continent == "Europe"),]

glimpse(Europe_map)

Europe_map <- Europe_map %>%
  dplyr::filter(!sovereignt == "Russia")

# The precision has to be set to a value > 0 to resolve internal boundaries.
st_precision(Europe_map) <- 1e15 # Required to
Europe_map <- st_union(Europe_map)

ggplot(Europe_map) +
  geom_sf() +
  coord_sf(xlim = c(-25,50), ylim = c(35,70), expand = FALSE)
```

## Crop climate raster data to area of Europe

To crop a raster with the stars package, square brackets [] will work as
crop operator (see here for details).

```{r}
iso_T_Eu <- iso_T[Europe_map]
# CRS for projected T and south america map are the same (EPSG 4326) but the
# proj4string includes more details for annual_T_70. Thus, they have to
# be made identical before cropping.
st_crs(iso_T_70) <- st_crs(Europe_map)
iso_T_70_Eu <- iso_T_70[Europe_map]
```

```{r}
nbreaks <- 9
# colors.
temp_colors <- colorRampPalette(c("#fff7ec", "#fee8c8", "#fdd49e", "#fdbb84", "#fc8d59", "#ef6548", "#d7301f", "#b30000", "#7f0000"))

{
  par(mfrow = c(1, 2))
  plot(iso_T_70_Eu, main = "Isothermality - 1960-1990",
       nbreaks = nbreaks,
       col = temp_colors(nbreaks - 1))
  plot(main = "Isothermality - RCP 4.5 projection for 2061-2080",
       iso_T_70_Eu, nbreaks = nbreaks,
       col = temp_colors(nbreaks - 1)) 
}
```
