---
title: "3_META_MODEL_FITTING"
author: "M.K.K. Lindhardt"
date: "2024-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


################################################################################
Sources for developing approaches in meta-analysis: 

https://wviechtb.github.io/metafor/reference/rma.mv.html 
https://www.taylorfrancis.com/chapters/edit/10.1201/9780429061240-8/multivariate-meta-analysis-ding-geng-din-chen-karl-peace 
https://training.cochrane.org/handbook/current/chapter-06 
https://training.cochrane.org/handbook/current/chapter-10


It is not uncommon for studies to have multiple outcome measures which lead to statistical multivariate analysis. The results of studies with multiple outcomes and/or endpoints are typically synthesized via conventional univariate meta-analysis (UMA) on each outcome separately, ignoring the correlations between the outcomes. The impact of ignoring the within-study correlation has been explored extensively in the statistical literature, with issues including overestimated variance of the pooled eﬀect size and biased estimates, which in turn may inﬂuence statistical inferences. In this case, multivariate meta-analysis should be used to synthesize multiple outcomes while taking into account their correlations, often resulting in superior parameter estimation. With study-level moderators or predictors, multivariate meta-regression can also be developed in parallel with multivariate regression techniques.

Meta-Analysis with R Package metafor. Rearrange the Data Format. In order to use the metafor package for multivariate meta-analysis, meta-data
should be rearranged accordingly. With this rearranged data format, we then construct a list of the variance-covariance matrices of the observed outcomes for the ﬁve studies to create a block diagonal matrix, V, for metafor.
With the rearranged data and variance-covariance matrix, V, we now ﬁt the ﬁxed-eﬀects meta-analysis model using metafor with the option of
method="FE". To ﬁt the random-eﬀects multivariate meta-analysis model using the metafor package, we simply change the option to method="REML". We note that this reproduces the results, with this random-effects model, we can also test the difference between

#####################################################

Response Variables are measured on a continuous scale, where each individual outcome is a measurement of a numerical quantity
Chosen Effect Size:  

Formulas to estimate effects (and their standard errors)?

#####################################################
Step-by-Step Framework for Meta-Analysis

1) Data Preparation
Clean and transform the data.
Standardize location information and create unique identifiers.
Classify locations by continent.
--- DESCRIPTIVE_VIZ: Exploratory data analysis and visualization.
Standardize measures of variation and convert SE to SD (save this version).
Impute missing values (silvo_se, control_se, silvo_n, control_n).
Convert SE to SD in the imputed dataset (save this version).
Calculate effect sizes (ROM) for both non-imputed and imputed datasets.
Compare the imputed and non-imputed datasets using descriptive statistics (mean, SD, median) and tests (density plots, boxplots, t-tests, Kolmogorov-Smirnov test).

2) Meta-Analysis Model Fitting
Use multivariate/multilevel mixed-effects models (rma.mv).
Fit models on both non-imputed and imputed datasets.
Compare results side by side, including effect sizes, confidence intervals, and heterogeneity statistics.

3) Sensitivity Analysis and Diagnostics
Perform sensitivity analysis, including leave-one-out analysis.
Conduct trim-and-fill analysis to check for publication bias.

4) Moderator Analysis
Split the dataset by moderators and conduct analyses on both imputed and non-imputed data.
Consider meta-regression to formally test for differences in moderator effects between the datasets.

5) Visualizations
Create comprehensive visual summaries, including caterpillar plots, forest plots, and geographical maps of study locations.

#############
# STEP 0
##########################################################################################################################################
PREPARING SCRIPT AND READ IN THE DATA
##########################################################################################################################################

```{r}
# Clean workspace
rm(list = ls())
```


## Loading required packages and libraries

```{r Loading other needed packages, warning=FALSE}
# Suppress warnings to avoid clutter in the console output
suppressWarnings({

  # Load multiple add-on packages using pacman::p_load for efficiency
  pacman::p_load(
    # Data Manipulation / Transformation
    tidyverse,        # Comprehensive collection of R packages for data science
    readr,            # Read and write csv 
    future,           # Parallel processing
    future.apply,     # Parallel processing
    ###################################################################################################################
    # Data Visualization
    ggplot2,          # Data visualization package (part of tidyverse)
    patchwork,        # ggplot2 API for sequentially building up a plot
    ###################################################################################################################
    # Meta-Analysis
    metafor,          # For conducting meta-analysis, effect sizes, and response ratios
    clubSandwich,     # Cluster-robust variance estimators for ordinary and weighted least squares linear regression models
    ###################################################################################################################
    # Exploratory Data Analysis (EDA)
    DataExplorer,     # For exploratory data analysis
    SmartEDA,         # For smart exploratory data analysis
    ###################################################################################################################
    # Project Management and Code Styling
    here,             # Easy file referencing using the top-level directory of a file project
    styler            # For code formatting and styling
  )
  
  # If encountering issues with raster::getData(), consider using geodata::worldclim_global() from the geodata package
})

###################################################################################################################
# Set preferences for conflicting functions
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("group_by", "dplyr") 
conflict_prefer("summarise", "dplyr")
conflict_prefer("mutate", "dplyr")
conflict_prefer("arrange", "dplyr")
conflict_prefer("locale", "readr")
```


Loading the two datasets (imputed and non-imputed)

```{r Loading database, warning=FALSE, message=FALSE}
# Set the working directory automatically using 'here'
setwd(here::here())

# Suppress warnings to avoid clutter in the console output
suppressWarnings({

# Define file paths
file_meta_data_ROM_non_imp <- here::here("DATA", "OUTPUT_FROM_R", "meta_data_ROM_non_imp.csv")
file_meta_data_ROM_imp <- here::here("DATA", "OUTPUT_FROM_R", "meta_data_ROM_imp.csv")

# Read in the non-imputed dataset
non_imp_dataset <- read_csv(file_meta_data_ROM_non_imp) %>%
  as.data.frame()

# Read in the imputed dataset
imp_dataset <- read_csv(file_meta_data_ROM_imp) %>%
  as.data.frame()
})
```



#############
# STEP 1
##########################################################################################################################################
PERFORMING MULTIVARIATE/MULTILEVEL LINEAR (MIXED-EFFECTS) MODELLING 
##########################################################################################################################################

```{r}
# Set up the parallel processing plan
plan(multisession, workers = parallel::detectCores() - 1)

# Function to fit the multivariate random-effects model
fit_meta_model <- function(data, dataset_name) {
  # Start time tracking
  start_time <- Sys.time()
  cat("Starting model fitting for", dataset_name, "...\n")

  # Check if the required columns are present
  if (!all(c("yi", "vi", "id_article", "response_variable", "exp_id") %in% colnames(data))) {
    stop("Dataset must contain 'yi', 'vi', 'id_article', 'response_variable', and 'exp_id' columns.")
  }

  # Fit the multivariate random-effects model with additional random effect for 'exp_id'
  model <- tryCatch({
    rma.mv(
      yi = yi,                # Effect size estimates
      V = vi,                 # Variance-covariance matrix of the effect sizes
      random = list(
        ~ 1 | id_article,                   # Random effect for between-study variability
        ~ 1 | id_article/response_variable, # Nested random effect for within-study variability
        ~ 1 | exp_id                       # Random effect for experiment-level variability
      ),
      data = data,            # Meta-analysis data
      verbose = TRUE,         # Enable verbose output for tracking progress
      control = list(
        optimizer = "optim",         # Use 'optim' optimizer
        optim.method = "BFGS",       # Use BFGS optimization method
        iter.max = 1000,             # Maximum number of iterations
        rel.tol = 1e-8               # Convergence tolerance
      ),
      method = "ML"           # Use Maximum Likelihood estimation
    )
  }, error = function(e) {
    cat("Error in model fitting for", dataset_name, ":", e$message, "\n")
    return(NULL)
  })

  # End time tracking
  end_time <- Sys.time()
  cat("Model fitting completed for", dataset_name, "in", round(difftime(end_time, start_time, units = "mins"), 2), "minutes.\n")

  return(model)
}

# Apply the function to both non-imputed and imputed datasets
cat("\nFitting model for Non-Imputed Data:\n")
model_non_imp <- fit_meta_model(non_imp_dataset, "Non-Imputed Data")

cat("\nFitting model for Imputed Data:\n")
model_imp <- fit_meta_model(imp_dataset, "Imputed Data")

# Check and compare model summaries
if (!is.null(model_non_imp)) {
  cat("\nSummary of Non-Imputed Model:\n")
  print(summary(model_non_imp))
}

if (!is.null(model_imp)) {
  cat("\nSummary of Imputed Model:\n")
  print(summary(model_imp))
}

# Calculate the time taken for both models
cat("\nTotal time taken for model fitting:\n")
total_time <- Sys.time() - start_time
print(total_time)
```

```{r}
summary(model_non_imp)
summary(model_imp)

```
```{r}
# Fixed effects estimates
fixed_effects_non_imp <- coef(model_non_imp)
fixed_effects_imp <- coef(model_imp)

print(fixed_effects_non_imp)
print(fixed_effects_imp)

```

```{r}
# Variance components (random effects)
var_components_non_imp <- model_non_imp$sigma2
var_components_imp <- model_imp$sigma2

cat("Variance components for Non-Imputed Model:", var_components_non_imp, "\n")
cat("Variance components for Imputed Model:", var_components_imp, "\n")

```

```{r}
# Heterogeneity statistics
# Function to calculate I2 for a given model
calculate_I2 <- function(model) {
  # Extract variance components
  tau2 <- sum(model$sigma2) # Sum of all random effect variance components
  sigma2 <- mean(model$vi)  # Mean of the within-study variances

  # Calculate I^2
  I2 <- tau2 / (tau2 + sigma2)
  return(I2 * 100) # Convert to percentage
}

# Calculate I^2 for non-imputed and imputed models
I2_non_imp <- calculate_I2(model_non_imp)
I2_imp <- calculate_I2(model_imp)

cat("I^2 for Non-Imputed Model:", round(I2_non_imp, 2), "%\n")
cat("I^2 for Imputed Model:", round(I2_imp, 2), "%\n")

```

```{r}
# Residuals
residuals_non_imp <- residuals(model_non_imp)
residuals_imp <- residuals(model_imp)

# Fitted values
fitted_values_non_imp <- fitted(model_non_imp)
fitted_values_imp <- fitted(model_imp)

# Print a few residuals and fitted values
head(data.frame(Residuals_Non_Imp = residuals_non_imp, Fitted_Non_Imp = fitted_values_non_imp))
head(data.frame(Residuals_Imp = residuals_imp, Fitted_Imp = fitted_values_imp))
```

```{r}
# Log-Likelihood, AIC, and BIC
logLik_non_imp <- logLik(model_non_imp)
logLik_imp <- logLik(model_imp)

AIC_non_imp <- AIC(model_non_imp)
AIC_imp <- AIC(model_imp)

BIC_non_imp <- BIC(model_non_imp)
BIC_imp <- BIC(model_imp)

cat("Non-Imputed Model - Log-Likelihood:", logLik_non_imp, "AIC:", AIC_non_imp, "BIC:", BIC_non_imp, "\n")
cat("Imputed Model - Log-Likelihood:", logLik_imp, "AIC:", AIC_imp, "BIC:", BIC_imp, "\n")

```

```{r}
# Confidence intervals
conf_int_non_imp <- confint(model_non_imp)
conf_int_imp <- confint(model_imp)

print(conf_int_non_imp)
print(conf_int_imp)

```

```{r}
# Forest plot for non-imputed data
cat("\nForest Plot for Non-Imputed Model:\n")
forest(model_non_imp, showweights = TRUE, xlab = "Effect Size (yi)")

# Forest plot for imputed data
cat("\nForest Plot for Imputed Model:\n")
forest(model_imp, showweights = TRUE, xlab = "Effect Size (yi)")

```

```{r}
# Standardized residuals
rstandard_non_imp <- rstandard(model_non_imp)
rstandard_imp <- rstandard(model_imp)

# Plot standardized residuals
plot(rstandard_non_imp, main = "Standardized Residuals - Non-Imputed Data")
plot(rstandard_imp, main = "Standardized Residuals - Imputed Data")

```

```{r}
# Extract key results into a data frame
results_df <- data.frame(
  Dataset = c("Non-Imputed", "Imputed"),
  Fixed_Effect = c(fixed_effects_non_imp, fixed_effects_imp),
  Tau2 = c(var_components_non_imp[1], var_components_imp[1]),
  I2 = c(I2_non_imp, I2_imp),
  AIC = c(AIC_non_imp, AIC_imp),
  BIC = c(BIC_non_imp, BIC_imp)
)

print(results_df)

```

