---
title: "4_SENSITIVITY_MODERATOR_ANALYSIS"
author: "M.K.K. Lindhardt"
date: "2024-11-17"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



################################################################################
Sources for developing approaches in meta-analysis: 

https://wviechtb.github.io/metafor/reference/rma.mv.html 
https://www.taylorfrancis.com/chapters/edit/10.1201/9780429061240-8/multivariate-meta-analysis-ding-geng-din-chen-karl-peace 
https://training.cochrane.org/handbook/current/chapter-06 
https://training.cochrane.org/handbook/current/chapter-10


It is not uncommon for studies to have multiple outcome measures which lead to statistical multivariate analysis. The results of studies with multiple outcomes and/or endpoints are typically synthesized via conventional univariate meta-analysis (UMA) on each outcome separately, ignoring the correlations between the outcomes. The impact of ignoring the within-study correlation has been explored extensively in the statistical literature, with issues including overestimated variance of the pooled eﬀect size and biased estimates, which in turn may inﬂuence statistical inferences. In this case, multivariate meta-analysis should be used to synthesize multiple outcomes while taking into account their correlations, often resulting in superior parameter estimation. With study-level moderators or predictors, multivariate meta-regression can also be developed in parallel with multivariate regression techniques.

Meta-Analysis with R Package metafor. Rearrange the Data Format. In order to use the metafor package for multivariate meta-analysis, meta-data
should be rearranged accordingly. With this rearranged data format, we then construct a list of the variance-covariance matrices of the observed outcomes for the ﬁve studies to create a block diagonal matrix, V, for metafor.
With the rearranged data and variance-covariance matrix, V, we now ﬁt the ﬁxed-eﬀects meta-analysis model using metafor with the option of
method="FE". To ﬁt the random-eﬀects multivariate meta-analysis model using the metafor package, we simply change the option to method="REML". We note that this reproduces the results, with this random-effects model, we can also test the difference between

#####################################################

Response Variables are measured on a continuous scale, where each individual outcome is a measurement of a numerical quantity
Chosen Effect Size:  

Formulas to estimate effects (and their standard errors)?

#####################################################
Step-by-Step Framework for Meta-Analysis

1) Data Preparation
Clean and transform the data.
Standardize location information and create unique identifiers.
Classify locations by continent.
--- DESCRIPTIVE_VIZ: Exploratory data analysis and visualization.
Standardize measures of variation and convert SE to SD (save this version).
Impute missing values (silvo_se, control_se, silvo_n, control_n).
Convert SE to SD in the imputed dataset (save this version).
Calculate effect sizes (ROM) for both non-imputed and imputed datasets.
Compare the imputed and non-imputed datasets using descriptive statistics (mean, SD, median) and tests (density plots, boxplots, t-tests, Kolmogorov-Smirnov test).

2) Meta-Analysis Model Fitting
Use multivariate/multilevel mixed-effects models (rma.mv).
Fit models on both non-imputed and imputed datasets.
Compare results side by side, including effect sizes, confidence intervals, and heterogeneity statistics.

3) Sensitivity Analysis and Diagnostics
Perform sensitivity analysis, including leave-one-out analysis.
Conduct trim-and-fill analysis to check for publication bias.

4) Moderator Analysis
Split the dataset by moderators and conduct analyses on both imputed and non-imputed data.
Consider meta-regression to formally test for differences in moderator effects between the datasets.

5) Visualizations
Create comprehensive visual summaries, including caterpillar plots, forest plots, and geographical maps of study locations.

#############
# STEP 0
##########################################################################################################################################
PREPARING SCRIPT AND READ IN THE DATA
##########################################################################################################################################

```{r}
# Clean workspace
rm(list = ls())
```


## Loading required packages and libraries

```{r Loading other needed packages, warning=FALSE}
# Suppress warnings to avoid clutter in the console output
suppressWarnings({

  # Load multiple add-on packages using pacman::p_load for efficiency
  pacman::p_load(
    # Data Manipulation / Transformation
    tidyverse,        # Comprehensive collection of R packages for data science
    readr,            # Read and write csv 
    dlookr,           # Diagnose, explore, and transform data with dlookr
    skimr,            # Provides easy summary statistics about variables in data frames, tibbles, data tables and vectors
    janitor,          # For cleaning and renaming data columns
    readxl,           # To read Excel files
    vroom,            # Fast reading of large datasets from local disk
    missForest,       # Random Forest method for imputing missing data
    mice,             # For dealing with missing data by creating multiple imputations for multivariate missing data
    missRanger,       # Fast missing value imputation by chained random forest
    conflicted,       # An alternative conflict resolution strategy
    future,           # Parallel processing
    future.apply,     # Parallel processing
    Matrix,           # Matrix operations
    progressr,
    ###################################################################################################################
    # Data Visualization
    ggplot2,          # Data visualization package (part of tidyverse)
    patchwork,        # ggplot2 API for sequentially building up a plot
    ###################################################################################################################
    # Spatial Data
    tidygeocoder,     # Unified interface for performing both forward and reverse geocoding queries
    raster,           # For spatial data analysis, especially BioClim variables from WorldClim
    sp,               # For spatial data classes and methods
    sf,               # For simple features in R, handling vector data
    rnaturalearth,    # For world map data
    rnaturalearthdata, 
    ###################################################################################################################
    # Meta-Analysis
    metafor,          # For conducting meta-analysis, effect sizes, and response ratios
    clubSandwich,     # Cluster-robust variance estimators for ordinary and weighted least squares linear regression models
    ###################################################################################################################
    # Exploratory Data Analysis (EDA)
    DataExplorer,     # For exploratory data analysis
    SmartEDA,         # For smart exploratory data analysis
    ###################################################################################################################
    # Project Management and Code Styling
    here,             # Easy file referencing using the top-level directory of a file project
    styler            # For code formatting and styling
  )
  
  # If encountering issues with raster::getData(), consider using geodata::worldclim_global() from the geodata package
})

###################################################################################################################
# Set preferences for conflicting functions
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("group_by", "dplyr") 
conflict_prefer("summarise", "dplyr")
conflict_prefer("mutate", "dplyr")
conflict_prefer("arrange", "dplyr")
conflict_prefer("locale", "readr")
conflict_prefer("extract", "raster")
conflict_prefer("intersect", "base")
```


Loading the datasets

```{r Loading database, warning=FALSE, message=FALSE}
# Set the working directory automatically using 'here'
setwd(here::here())


# Define your working directory using 'here'
dir <- here::here("DATA", "OUTPUT_FROM_R", "SAVED_OBJECTS_FROM_R")

# Load datasets 

## NON-IMPUTED
non_imp_dataset <- readRDS(here::here("DATA", "OUTPUT_FROM_R", "SAVED_OBJECTS_FROM_R", "non_imp_data_rom.rds"))

## IMPUTED
imp_dataset <- readRDS(here::here("DATA", "OUTPUT_FROM_R", "SAVED_OBJECTS_FROM_R", "imp_data_rom.rds"))

```

```{r}
imp_dataset |> glimpse()
```

```{r}
# Check the unique response variables and sub-response variables
unique(imp_dataset$response_variable)
# unique(imp_dataset$sub_response_variable)
```
```{r}
response_summary <- 
  imp_dataset |> 
  count(response_variable) |> 
  arrange(desc(n))

# Bar plot for response variable counts
response_var_summary_plot <- 
  response_summary |> 
  ggplot(aes(x = reorder(response_variable, -n), y = n)) +
  geom_bar(stat = "identity", fill = "#0072B2") +
  coord_flip() +
  labs(title = "Count of Observations per Response Variable",
       x = "Response Variable",
       y = "Count of Observations") +
  theme_minimal()

response_var_summary_plot
```

```{r}
# Check the unique response variables and sub-response variables
unique_articles <- unique(imp_dataset$id_article)
unique_response_vars <- unique(imp_dataset$response_variable)

print(unique_response_vars)
print(length(unique_articles))  # Total number of unique articles

```

```{r}
# Count the number of unique articles for each response variable
article_summary <- imp_dataset %>%
  group_by(response_variable) %>%
  summarise(unique_articles = n_distinct(id_article)) %>%
  arrange(desc(unique_articles))

article_summary

```
```{r}
# Bar plot for unique article counts by response variable
article_summary_plot <-
  article_summary |> 
  ggplot(aes(x = reorder(response_variable, -unique_articles), y = unique_articles)) +
  geom_bar(stat = "identity", fill = "#E69F00") +
  coord_flip() +
  labs(title = "Count of Unique Articles per Response Variable",
       x = "Response Variable",
       y = "Number of Unique Articles") +
  theme_minimal()

article_summary_plot
```

```{r}
# Sample one observation per article per response variable
sampled_data <- imp_dataset %>%
  group_by(id_article, response_variable) %>%
  slice_head(n = 1)

sampled_data
```
```{r}
# Boxplot of effect sizes by response variable (unique articles)
boxplot_effec_size_response_variable <- sampled_data |> 
  ggplot(aes(x = response_variable, 
             y = yi, 
             fill = response_variable)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Boxplot of Effect Sizes (yi) by Response Variable (Unique Articles)",
       x = "Response Variable",
       y = "Effect Size (yi)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

boxplot_effec_size_response_variable
```

The boxplot shows the distribution of effect sizes (yi) for each response variable. In the context of this meta-analysis, the effect size (yi) represents the standardized difference between the treatment group (agroforestry systems) and the control group (monoculture).

Biodiversity: the effect sizes are mostly positive, suggesting that agroforestry systems might have a higher effect on biodiversity compared to monoculture.
Crop Yield: the effect sizes are genrally negative, indicating potential yield reductions in agroforestry systems.
Soil Quality: tend to show positive effect sizes, suggesting an improvement in these aspects under agroforestry systems.
Product Quality: 
Pest and Diseases:
Greenhouse gas emissions:

The variation in the effect sizes for each response variable indicates heterogeneity, showing that the impact of agroforestry varies across different studies and contexts. However, the effect sizes in this plot do not account for potential study-level covariates or moderators, which could influence the observed differences. It will be important to consider the sampling variance (vi) associated with each effect size when interpreting these results, as larger studies (with smaller variance) provide more reliable estimates.





```{r}
# Calculate mean effect size and number of unique articles for each response variable
response_summary <- sampled_data %>%
  group_by(response_variable) %>%
  summarise(
    mean_yi = mean(yi, na.rm = TRUE),
    num_articles = n_distinct(id_article)
  )

# Merge the summary data back into the sampled data for ordering
sampled_data <- sampled_data %>%
  left_join(response_summary, by = "response_variable")

# Plot histogram of effect sizes with additional improvements
ggplot(sampled_data, aes(x = yi, fill = response_variable)) +
  geom_histogram(binwidth = 0.05, alpha = 0.7, color = "black") +
  # mean yi line
  geom_vline(data = response_summary, aes(xintercept = mean_yi, color = response_variable),
             linetype = "dashed", size = 0.8) +
  facet_wrap(~ reorder(response_variable, -num_articles), scales = "free_y") +
  labs(
    title = "Distribution of Effect Sizes (yi) by Response Variable (Unique Articles)",
    x = "Effect Size (yi)",
    y = "Frequency"
  ) +
  scale_fill_brewer(palette = "Set2") +
  scale_color_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 16)
  )

```

```{r}
# Calculate mean effect size and number of unique articles for each response variable
response_summary <- sampled_data %>%
  group_by(response_variable) %>%
  summarise(
    mean_yi = mean(yi, na.rm = TRUE),
    num_articles = n_distinct(id_article)
  ) %>%
  arrange(desc(num_articles))

# Merge the summary data back into the sampled data for ordering
sampled_data <- sampled_data %>%
  left_join(response_summary, by = "response_variable") %>%
  mutate(response_variable = factor(response_variable, levels = response_summary$response_variable))

# Reshape the data to long format for plotting
long_data <- sampled_data %>%
  select(id_article, response_variable, silvo_mean, control_mean) %>%
  pivot_longer(cols = c(silvo_mean, control_mean), 
               names_to = "group", 
               values_to = "mean_value") %>%
  mutate(group = recode(group, "silvo_mean" = "Silvo (Agroforestry)", "control_mean" = "Control (Monoculture)"))

# Calculate mean values for each group and response variable
group_summary <- long_data %>%
  group_by(response_variable, group) %>%
  summarise(
    mean_value = mean(mean_value, na.rm = TRUE),
    .groups = "drop"
  )


# Plot the distribution of mean values for Silvo and Control groups with free x-scales
ggplot(long_data, aes(x = mean_value, fill = group)) +
  geom_histogram(binwidth = 5, alpha = 0.7, color = "black", position = "identity") +
  geom_vline(data = group_summary, aes(xintercept = mean_value, color = group),
             linetype = "dashed", size = 0.8) +
  facet_wrap(~ response_variable, scales = "free_x") +  # Set scales to "free_x"
  labs(
    title = "Distribution of Mean Values for Silvo (Agroforestry) and Control (Monoculture)",
    x = "Mean Value",
    y = "Frequency"
  ) +
  scale_fill_manual(values = c("#0072B2", "#D55E00")) +
  scale_color_manual(values = c("#0072B2", "#D55E00")) +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 16)
  )

```

```{r}
# List of response variables and moderators
response_variables <- c("Biodiversity", "Crop yield", "Water quality", "Pest and Disease", 
                        "Soil quality", "Greenhouse gas emission", "Product quality")

moderators <- c("tree_type", "crop_type", "age_system", "season", 
                "soil_texture", "no_tree_per_m", "tree_height", "alley_width")
```

```{r}
# Define a generic function for counting unique values
count_unique_response_and_moderator <- function(data, response_col, moderators) {
  
  # Count of unique response variables for each moderator
  unique_response_var_per_moderator <- data %>%
    select(all_of(response_col), all_of(moderators)) %>%
    pivot_longer(cols = all_of(moderators), names_to = "moderator", values_to = "moderator_value") %>%
    filter(!is.na(moderator_value)) %>%
    group_by(moderator) %>%
    summarise(unique_responses = n_distinct(.data[[response_col]]), .groups = "drop")

  # Count of unique moderator levels for each response variable
  unique_moderators_per_response_var <- data %>%
    select(all_of(response_col), all_of(moderators)) %>%
    pivot_longer(cols = all_of(moderators), names_to = "moderator", values_to = "moderator_value") %>%
    filter(!is.na(moderator_value)) %>%
    group_by(.data[[response_col]]) %>%
    summarise(unique_moderators = n_distinct(moderator_value), .groups = "drop")

  # Return both summaries as a list
  list(
    unique_response_var_per_moderator = unique_response_var_per_moderator,
    unique_moderators_per_response_var = unique_moderators_per_response_var
  )
}

```

```{r}
# Define your response column and list of moderators
response_col <- "response_variable"
moderators <- c("tree_type", "crop_type", "age_system", "season", 
                "soil_texture", "no_tree_per_m", "tree_height", "alley_width")

# Apply the function to the non-imputed dataset
non_imp_summary <- count_unique_response_and_moderator(non_imp_dataset, response_col, moderators)

# Apply the function to the imputed dataset
imp_summary <- count_unique_response_and_moderator(imp_dataset, response_col, moderators)

# View the summaries
print(non_imp_summary$unique_response_var_per_moderator)
print(non_imp_summary$unique_moderators_per_response_var)

print(imp_summary$unique_response_var_per_moderator)
print(imp_summary$unique_moderators_per_response_var)

```


Visualize the Summary Data

Plot 1: Number of Unique Response Variables per Moderator
```{r}
# Define your response column and list of moderators
response_col <- "response_variable"
moderators <- c("tree_type", "crop_type", "age_system", "season", 
                "soil_texture", "no_tree_per_m", "tree_height", "alley_width")

# Generate the summaries using the function
results <- count_unique_response_and_moderator(non_imp_dataset, response_col, moderators)

# Extract the specific summary for plotting
unique_response_var_per_moderator <- results$unique_response_var_per_moderator

# Bar plot for unique response variables per moderator
unique_response_var_per_moderator |> 
  ggplot(aes(x = reorder(moderator, unique_responses), y = unique_responses, fill = moderator)) +
  geom_bar(stat = "identity", color = "black") +
  coord_flip() +
  labs(
    title = "Number of Unique Response Variables per Moderator",
    x = "Moderator",
    y = "Number of Unique Response Variables"
  ) +
  scale_fill_brewer(palette = "Set3") +
  theme_minimal() +
  theme(legend.position = "none")

```


Plot 2: Number of Unique Moderators per Response Variable
```{r}
# Define your response column and list of moderators
response_col <- "response_variable"
moderators <- c("tree_type", "crop_type", "age_system", "season", 
                "soil_texture", "no_tree_per_m", "tree_height", "alley_width")

# Generate summaries
results <- count_unique_response_and_moderator(non_imp_dataset, response_col, moderators)

# Extract the specific summary for plotting
unique_moderators_per_response_var <- results$unique_moderators_per_response_var


# Bar plot for unique moderators per response variable
unique_moderators_per_response_var |> 
  ggplot(aes(x = reorder(response_variable, unique_moderators), y = unique_moderators, fill = response_variable)) +
  geom_bar(stat = "identity", color = "black") +
  coord_flip() +
  labs(
    title = "Number of Unique Moderators per Response Variable",
    x = "Response Variable",
    y = "Number of Unique Moderator Levels"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(legend.position = "none")
```

Alternative Visualization: Heatmap
comprehensive view of the data in a heatmap to show the interaction between moderators and response variables

```{r}
# Create a heatmap data frame
heatmap_data <- non_imp_dataset %>%
  select(response_variable, all_of(moderators)) %>%
  pivot_longer(cols = all_of(moderators), names_to = "moderator", values_to = "moderator_value") %>%
  filter(!is.na(moderator_value)) %>%
  group_by(response_variable, moderator) %>%
  summarise(count = n_distinct(moderator_value), .groups = "drop")

# Heatmap plot
ggplot(heatmap_data, aes(x = moderator, y = response_variable, fill = count)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(
    title = "Heatmap of Unique Moderator Levels by Response Variable",
    x = "Moderator",
    y = "Response Variable",
    fill = "Count of Unique Levels"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

```




############
# STEP 6
##########################################################################################################################################
SENSITIVITY ANALYSIS (Leave-One-Out Sensitivity Analysis)
##########################################################################################################################################

non_imp_dataset 
imp_dataset 



Performing a Leave-One-Out (LOO) sensitivity analysis on the full datasets and full model can be justified. This analysis would complement the LOO performed on each individual meta-regression model and provide additional insights. Here’s a detailed explanation:

Justification for LOO Sensitivity Analysis on Full Model
Assess Overall Model Robustness:

Applying LOO sensitivity analysis to the full model allows you to evaluate the robustness of the overall meta-analytic findings. It helps identify if specific studies have a disproportionate influence on the aggregated results.
This is particularly useful when you have concerns about the quality or heterogeneity of the included studies. If the removal of a single study drastically changes the overall effect size or model fit, this study might need further investigation.
Comparison Between Datasets (Imputed vs Non-Imputed):

Conducting LOO on both non_imp_dataset and imp_dataset allows you to compare the stability of the results between datasets with and without imputation.
It helps validate the imputation process by checking whether imputed data introduces or mitigates influential observations.
Insights Across All Moderators:

The LOO analysis on individual meta-regression models examines the impact of leaving out observations within each moderator level. However, performing LOO on the full model assesses the overall influence of each study across all moderators combined.
This provides a broader view of the dataset and highlights influential studies that might not be detected when focusing only on specific moderators.
Pros and Cons of LOO on the Full Model
Pros:
Comprehensive Sensitivity Check: It provides a complete picture of the influence of each study across the entire dataset.
Validation of Imputed Data: It allows you to test the robustness of results obtained from imputed datasets, helping ensure that imputation did not introduce bias.
Identification of Influential Studies: It can detect influential studies that have a large impact on the overall meta-analytic effect size or variance components.
Cons:
Computational Intensity: The LOO analysis on the full model can be computationally expensive, especially with large datasets and complex random-effects structures.
Potential Overlap: If individual meta-regression models are already analyzed with LOO, the full-model LOO might yield redundant information. However, this risk is mitigated by focusing on the broader dataset perspective.
Complex Interpretation: If a study is found to be influential in the full model LOO but not in specific meta-regression models, interpretation can be challenging. It may indicate complex interactions or issues with the overall model specification.
Implementation Plan
You can implement the LOO sensitivity analysis on the full model using both imputed and non-imputed datasets. Here’s how to proceed:

Fit the Full Model on each dataset (non_imp_dataset, imp_dataset, non_imp_dataset_imputed, imp_dataset_imputed).
Run LOO Sensitivity Analysis using a function that iteratively removes one study at a time and refits the model.
Summarize the Results: Report changes in the effect size, heterogeneity statistics (I²), and model fit metrics (AIC, BIC, LogLik).

Why Split by Response Variable?
Distinct Moderation Effects: Response variables (e.g., ecosystem services like greenhouse gas emissions, soil quality, etc.) likely differ in how moderators (e.g., tree type, crop type, season) influence them. Splitting allows the model to isolate the unique relationships within each response variable.

Interpretability: Without splitting, estimates may conflate effects across distinct response variables, making it difficult to attribute results to a specific ecosystem service.

Appropriate LOO Sensitivity: Running LOO sensitivity on mixed-response datasets can be misleading, as the omitted studies might affect multiple response variables differently, introducing biases or artifacts.

Fits the Cabbage Approach: This methodology inherently groups analyses by discrete categories, and splitting by response variable mirrors that logic.

```{r}

```

