---
title: "4_SENSITIVITY_MODERATOR_ANALYSIS"
author: "M.K.K. Lindhardt"
date: "2024-11-17"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



################################################################################
Sources for developing approaches in meta-analysis: 

https://wviechtb.github.io/metafor/reference/rma.mv.html 
https://www.taylorfrancis.com/chapters/edit/10.1201/9780429061240-8/multivariate-meta-analysis-ding-geng-din-chen-karl-peace 
https://training.cochrane.org/handbook/current/chapter-06 
https://training.cochrane.org/handbook/current/chapter-10


It is not uncommon for studies to have multiple outcome measures which lead to statistical multivariate analysis. The results of studies with multiple outcomes and/or endpoints are typically synthesized via conventional univariate meta-analysis (UMA) on each outcome separately, ignoring the correlations between the outcomes. The impact of ignoring the within-study correlation has been explored extensively in the statistical literature, with issues including overestimated variance of the pooled eﬀect size and biased estimates, which in turn may inﬂuence statistical inferences. In this case, multivariate meta-analysis should be used to synthesize multiple outcomes while taking into account their correlations, often resulting in superior parameter estimation. With study-level moderators or predictors, multivariate meta-regression can also be developed in parallel with multivariate regression techniques.

Meta-Analysis with R Package metafor. Rearrange the Data Format. In order to use the metafor package for multivariate meta-analysis, meta-data
should be rearranged accordingly. With this rearranged data format, we then construct a list of the variance-covariance matrices of the observed outcomes for the ﬁve studies to create a block diagonal matrix, V, for metafor.
With the rearranged data and variance-covariance matrix, V, we now ﬁt the ﬁxed-eﬀects meta-analysis model using metafor with the option of
method="FE". To ﬁt the random-eﬀects multivariate meta-analysis model using the metafor package, we simply change the option to method="REML". We note that this reproduces the results, with this random-effects model, we can also test the difference between

#####################################################

Response Variables are measured on a continuous scale, where each individual outcome is a measurement of a numerical quantity
Chosen Effect Size:  

Formulas to estimate effects (and their standard errors)?

#####################################################
Step-by-Step Framework for Meta-Analysis

1) Data Preparation
Clean and transform the data.
Standardize location information and create unique identifiers.
Classify locations by continent.
--- DESCRIPTIVE_VIZ: Exploratory data analysis and visualization.
Standardize measures of variation and convert SE to SD (save this version).
Impute missing values (silvo_se, control_se, silvo_n, control_n).
Convert SE to SD in the imputed dataset (save this version).
Calculate effect sizes (ROM) for both non-imputed and imputed datasets.
Compare the imputed and non-imputed datasets using descriptive statistics (mean, SD, median) and tests (density plots, boxplots, t-tests, Kolmogorov-Smirnov test).

2) Meta-Analysis Model Fitting
Use multivariate/multilevel mixed-effects models (rma.mv).
Fit models on both non-imputed and imputed datasets.
Compare results side by side, including effect sizes, confidence intervals, and heterogeneity statistics.

3) Sensitivity Analysis and Diagnostics
Perform sensitivity analysis, including leave-one-out analysis.
Conduct trim-and-fill analysis to check for publication bias.

4) Moderator Analysis
Split the dataset by moderators and conduct analyses on both imputed and non-imputed data.
Consider meta-regression to formally test for differences in moderator effects between the datasets.

5) Visualizations
Create comprehensive visual summaries, including caterpillar plots, forest plots, and geographical maps of study locations.

#############
# STEP 0
##########################################################################################################################################
PREPARING SCRIPT AND READ IN THE DATA
##########################################################################################################################################

```{r}
# Clean workspace
rm(list = ls())
```


## Loading required packages and libraries

```{r Loading other needed packages, warning=FALSE}
# Suppress warnings to avoid clutter in the console output
suppressWarnings({

  # Load multiple add-on packages using pacman::p_load for efficiency
  pacman::p_load(
    # Data Manipulation / Transformation
    tidyverse,        # Comprehensive collection of R packages for data science
    readr,            # Read and write csv 
    dlookr,           # Diagnose, explore, and transform data with dlookr
    skimr,            # Provides easy summary statistics about variables in data frames, tibbles, data tables and vectors
    janitor,          # For cleaning and renaming data columns
    readxl,           # To read Excel files
    vroom,            # Fast reading of large datasets from local disk
    missForest,       # Random Forest method for imputing missing data
    mice,             # For dealing with missing data by creating multiple imputations for multivariate missing data
    missRanger,       # Fast missing value imputation by chained random forest
    conflicted,       # An alternative conflict resolution strategy
    future,           # Parallel processing
    future.apply,     # Parallel processing
    ###################################################################################################################
    # Data Visualization
    ggplot2,          # Data visualization package (part of tidyverse)
    patchwork,        # ggplot2 API for sequentially building up a plot
    ###################################################################################################################
    # Spatial Data
    tidygeocoder,     # Unified interface for performing both forward and reverse geocoding queries
    raster,           # For spatial data analysis, especially BioClim variables from WorldClim
    sp,               # For spatial data classes and methods
    sf,               # For simple features in R, handling vector data
    rnaturalearth,    # For world map data
    rnaturalearthdata, 
    ###################################################################################################################
    # Meta-Analysis
    metafor,          # For conducting meta-analysis, effect sizes, and response ratios
    clubSandwich,     # Cluster-robust variance estimators for ordinary and weighted least squares linear regression models
    ###################################################################################################################
    # Exploratory Data Analysis (EDA)
    DataExplorer,     # For exploratory data analysis
    SmartEDA,         # For smart exploratory data analysis
    ###################################################################################################################
    # Project Management and Code Styling
    here,             # Easy file referencing using the top-level directory of a file project
    styler            # For code formatting and styling
  )
  
  # If encountering issues with raster::getData(), consider using geodata::worldclim_global() from the geodata package
})

###################################################################################################################
# Set preferences for conflicting functions
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("group_by", "dplyr") 
conflict_prefer("summarise", "dplyr")
conflict_prefer("mutate", "dplyr")
conflict_prefer("arrange", "dplyr")
conflict_prefer("locale", "readr")
conflict_prefer("extract", "raster")
```


Loading the datasets

```{r Loading database, warning=FALSE, message=FALSE}
# Set the working directory automatically using 'here'
setwd(here::here())

# Define the output directory using 'here'
output_dir <- here::here("DATA", "OUTPUT_FROM_R", "SAVED_OBJECTS_FROM_R")

# Load all datasets and models
non_imp_dataset <- readRDS(file.path(output_dir, "non_imp_dataset.rds"))
imp_dataset <- readRDS(file.path(output_dir, "imp_dataset.rds"))
non_imp_dataset_imputed <- readRDS(file.path(output_dir, "non_imp_dataset_imputed.rds"))
imp_dataset_imputed <- readRDS(file.path(output_dir, "imp_dataset_imputed.rds"))
```




2. Splitting by Moderator Subset Data
Define the moderators and create a function to split the dataset by each moderator.

```{r}
# Define the list of moderators
moderators <- c("tree_type", "crop_type", "age_system", "season", 
                "soil_texture", "no_tree_per_m", "tree_height", "alley_width")

# Function to split the dataset by a specific moderator
split_data_by_moderator <- function(data, moderator) {
  cat("\nSplitting dataset by moderator:", moderator, "\n")
  split_data <- split(data, data[[moderator]])
  return(split_data)
}



```


3. Generating V_matrix for Each Moderator Subset Split Dataset
Generate the variance-covariance matrix for each subset of data.

```{r}
# Function to create V_matrix for each subset
generate_v_matrix_for_subset <- function(subset_data, V_matrix) {
  subset_indices <- rownames(subset_data)
  V_subset <- V_matrix[subset_indices, subset_indices]
  return(V_subset)
}

```


4. Running rma.mv for Each Subset Moderator Dataset
Fit the multivariate random-effects model for each level of the moderator.

```{r}
# Function to fit the rma.mv model for each level of the moderator
fit_models_for_moderator_levels <- function(split_data, V_matrix, dataset_name) {
  model_results <- list()

  for (level in names(split_data)) {
    cat("\nFitting model for level:", level, "in", dataset_name, "\n")

    # Subset data for the current level
    subset_data <- split_data[[level]]

    # Skip if there are insufficient data points
    if (nrow(subset_data) < 5) {
      cat("Insufficient data for level:", level, "- Skipping model fitting.\n")
      next
    }

    # Generate V_matrix for the subset
    V_subset <- generate_v_matrix_for_subset(subset_data, V_matrix)

    # Fit the model
    model <- tryCatch({
      rma.mv(
        yi = subset_data$yi,
        V = V_subset,
        mods = ~ 1,
        random = list(
          ~ 1 | id_article,
          ~ 1 | id_article/response_variable,
          ~ 1 | exp_id
        ),
        data = subset_data,
        method = "ML"
      )
    }, error = function(e) {
      cat("Error in model fitting for level:", level, "-", e$message, "\n")
      return(NULL)
    })

    # Store the model results
    if (!is.null(model)) {
      model_results[[level]] <- list(
        model = model,
        summary = summary(model),
        aic = AIC(model),
        bic = BIC(model),
        logLik = logLik(model)
      )
      cat("Model fitting complete for level:", level, "\n")
    }
  }

  return(model_results)
}

```

