---
title: "old_code_notes_to be_saved_but_not_used"
output: html_document
date: "2024-07-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#############
# STEP 2
##########################################################################################################################################
CALCULATING INITIAL EFFECT SIZES AND EVALUATING NUMBER OF STUDIES TO INCLUDE (SENSITIVITY ANALYSIS)
##########################################################################################################################################

continuous data, where each individual’s outcome is a measurement of a numerical quantity;

```{r}
## Calculate Effect Sizes and Variances for Multiple Parameters
# Log-transformed response ratios (lnRR) and corresponding variances
data_lnRR <- database_clean %>%
  filter(
    !is.na(silvo_mean) & !is.na(control_mean) & !is.na(silvo_n) & 
    !is.na(control_n) & !is.na(silvo_se) & !is.na(control_se)
  ) %>%
  mutate(
    lnRR = log(silvo_mean / control_mean),
    var_lnRR = (silvo_se^2 / (silvo_n * silvo_mean^2)) + 
               (control_se^2 / (control_n * control_mean^2)),
    slab = paste(Id_article, ", ", Study_Year_Start)
  ) %>%
  filter(
    !is.nan(lnRR) & !is.infinite(lnRR) & 
    !is.nan(var_lnRR) & !is.infinite(var_lnRR) & 
    var_lnRR > 0
  ) %>%
  relocate(Id_article, Response_variable, Sub_response_variable, silvo_mean, silvo_se, control_mean, control_se, lnRR, var_lnRR)

# Check the structure of the meta_data
glimpse(meta_data)
```

```{r}
meta_data %>% dplyr::glimpse() 
meta_data %>% View()
```


```{r}
## Meta-Analysis for Each Response Variable
results <- list()  # List to store results for each response variable
response_vars <- unique(meta_data$Response_variable)  # Get unique response variables

# Loop through each response variable
for (response in response_vars) {
  # Filter data for the current response variable and remove rows with non-positive or missing variances
  data_response <- filter(meta_data, Response_variable == response & var_lnRR > 0 & !is.na(var_lnRR))
  
  # Check if there's enough data to perform meta-analysis
  if (nrow(data_response) > 1) {
    # Fit random-effects model using log-transformed response ratios and variances
    res <- rma(yi = lnRR, vi = var_lnRR, data = data_response, method = "REML")
    
    # Store the results in the list
    results[[response]] <- res
    
    # Print summary of the meta-analysis
    cat("\nResponse Variable:", response, "\n")
    print(summary(res))
    
    # Print predicted pooled effect size and corresponding CI/PI
    print(predict(res, transf = exp, digits = 2))
    
    # Generate forest plot
    forest(res, xlab = "Log Response Ratio (lnRR)", slab = data_response$slab, main = response, cex = 0.8, cex.lab = 1.2)
    
    # Generate funnel plot to check for publication bias
    funnel(res)
  } else {
    cat("\nResponse Variable:", response, "\n")
    cat("Not enough data to perform meta-analysis.\n")
  }
}

## Sensitivity Analysis for Each Response Variable
# Loop through each response variable to perform influence diagnostics
for (response in response_vars) {
  res <- results[[response]]  # Get results for the current response variable
  
  if (!is.null(res)) {
    # Perform influence diagnostics
    inf <- influence(res)
    
    # Plot influence diagnostics
    plot(inf, main = paste("Influence Diagnostics for", response))
    
    # Print summary statistics for interpretation
    cat("\nResponse Variable:", response, "\n")
    cat("Overall Effect Size (lnRR):", res$b, "\n")
    cat("95% Confidence Interval:", confint(res)$ci.lb, "to", confint(res)$ci.ub, "\n")
    cat("Heterogeneity (Q):", res$QE, "\n")
    cat("I^2:", res$I2, "%\n")
  }
}
```







################################
MANUAL CALCULATION OF EFFECT SIZE MEASURE
################################

log-transformed response ratio (lnRR)
```{r}
# Calculate Effect Sizes (lnRR) and Variances for Multiple Parameters
meta_data_lnRR <- meta_data |> 
  # Group by article ID and response variable
  group_by(id_article, response_variable) |> 
  # Summarise to get means per group and include study_year_start
  summarise(
    silvo_mean = mean(silvo_mean, na.rm = TRUE),   # Mean of silvo_mean per group
    control_mean = mean(control_mean, na.rm = TRUE), # Mean of control_mean per group
    silvo_se = mean(silvo_se, na.rm = TRUE),         # Mean of silvo_se per group
    control_se = mean(control_se, na.rm = TRUE),     # Mean of control_se per group
    silvo_n = mean(silvo_n, na.rm = TRUE),           # Mean of silvo_n per group
    control_n = mean(control_n, na.rm = TRUE),       # Mean of control_n per group
    study_year_start = first(study_year_start),      # Include study_year_start
    .groups = 'drop'
  ) |> 
  # Calculate log response ratio (lnRR) and its variance (var_lnRR)
  mutate(
    # Compute log response ratio
    lnRR = log(silvo_mean / control_mean), 
    # Variance of lnRR combining silvo and control variances
    var_lnRR = (silvo_se^2 / (silvo_n * silvo_mean^2)) +  
               (control_se^2 / (control_n * control_mean^2)),  
    # Create label for plotting
    slab = paste(id_article, ", ", study_year_start)  
  ) |> 
  # Filter out rows with NaN, infinite, or non-positive variance values
  filter(
    !is.nan(lnRR) & !is.infinite(lnRR) & 
    !is.nan(var_lnRR) & !is.infinite(var_lnRR) & 
    var_lnRR > 0
  ) |> 
  # Reorder columns for better readability and organization
  relocate(id_article, response_variable, silvo_mean, control_mean, lnRR, var_lnRR, slab)
```

```{r}
# Display a glimpse of the final dataset for verification
meta_data_lnRR |> glimpse()
```


Step 1: Inspect the Variances

```{r}
# Remove rows with missing values in filtered_meta_data_ROM
filtered_meta_data_ROM_clean <- filtered_meta_data_ROM[complete.cases(filtered_meta_data_ROM), ]

# Check for missing values again to confirm
missing_values_updated <- sapply(filtered_meta_data_ROM_clean, function(x) sum(is.na(x)))
print("Missing Values in updated filtered_meta_data_ROM:")
print(missing_values_updated)

filtered_meta_data_ROM_clean
```


```{r}
# Check for missing values
meta_data_RR |> as.data.frame() |> summary()

# Check for extreme values in variances
hist(meta_data_RR$vi, breaks = 10, main = "Histogram of Variances", xlab = "Variance of yi")
```
```{r}
# Simplify the model by removing the multilevel structure
simple_res <- rma.mv(yi = yi, 
                     V = V_matrix, 
                     random = ~ 1 | id_article, 
                     data = meta_data_RR)
```


```{r}
# List of optimizers to try
optimizers <- c("nlminb", "optim", "Nelder-Mead", "BFGS", "CG", "L-BFGS-B", "SANN", "Brent")

# Function to fit the random-effects model with different optimizers
fit_model <- function() {
  for (opt in optimizers) {
    res <- tryCatch({
      message(paste("Trying optimizer:", opt))
      rma.mv(yi = yi, 
             V = V_matrix, 
             random = ~ 1 | id_article/response_variable, 
             data = meta_data_RR,
             verbose = TRUE,
             control = list(iter.max = 1000, rel.tol = 1e-8, optimizer = opt))
    }, error = function(e) {
      message(paste("Optimizer", opt, "failed with error:", e$message))
      NULL
    })
    if (!is.null(res)) {
      if (res$convergence == 0) {
        message(paste("Optimizer", opt, "converged successfully."))
        return(res)
      }
    }
  }
  stop("None of the optimizers converged.")
}

```


```{r}
# Fit the random-effects model using the filtered data
res_filtered <- rma.mv(yi = yi, 
                       V = V_matrix, 
                       random = ~ 1 | id_article/response_variable, 
                       data = filtered_meta_data_ROM,
                       verbose = TRUE,
                       control = list(iter.max = 1000, rel.tol = 1e-8))


```



```{r}
# Plot Cook's Distance
plot(cooks_d, type="h", ylab="Cook's Distance", xlab="Observed Outcome",
     main="Cook's Distance Plot")
abline(h=4/(nrow(filtered_meta_data_ROM_clean)-length(coef(res))), lty=2, col="red")

# Plot Studentized Residuals
plot(student_resid$resid, type="h", ylim=c(-3, 3),
     ylab="Studentized Residuals", xlab="Observed Outcome",
     main="Studentized Residuals Plot")
abline(h=c(-2, 2), lty=2, col="red")

# Plot DFBETAS for each coefficient
dfbetas_df <- as.data.frame(dfbetas_vals)
for (i in 1:ncol(dfbetas_df)) {
  plot(dfbetas_df[, i], type="h", ylab=paste("DFBETAS for Coef", i), xlab="Observed Outcome",
       main=paste("DFBETAS for Coefficient", i))
  abline(h=c(-2, 2), lty=2, col="red")
}

# Plot Hat Values
plot(hat_vals, type="h", ylab="Hat Values", xlab="Observed Outcome",
     main="Hat Values Plot")
abline(h=2*mean(hat_vals), lty=2, col="red")
```



























```{r}
# Step 1: Inspect the Variances
# Inspect the distribution of var_lnRR
summary(meta_data_lnRR$var_lnRR)
hist(meta_data_lnRR$var_lnRR, 
     breaks = 30, 
     main = "Histogram of Variances", 
     xlab = "Variance of lnRR")
```
```{r}
# Create the density plot for var_lnRR
ggplot(meta_data_lnRR, aes(x = var_lnRR)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of Variances", 
       x = "Variance of lnRR", 
       y = "Density") +
  theme_minimal()
```














































##########################################################################################################################################
Using the build-inn metafor function "escalc" to calculate RoM" for the log transformed ratio of means
##########################################################################################################################################

```{r}
###################################################################
# Ensure standard deviations are valid for escalc
meta_data <- meta_data %>%
   # Ensure standard deviations are valid for escalc
    filter(silvo_se > 0 & control_se > 0) 
    # Ensure means are greater than 0 to avoid infinite values
    #filter(silvo_mean > 0 & control_mean > 0) 

###################################################################
# Use the escalc function to compute effect sizes and variances
meta_data_es <- metafor::escalc(
  # "ROM" for the log transformed ratio of means
  measure = "ROM", 
  m1i = silvo_mean, m2i = control_mean, 
  sd1i = silvo_se, sd2i = control_se, 
  n1i = silvo_n, n2i = control_n, 
  data = meta_data, 
  slab = paste(id_article, response_variable, sep = ", ")) %>% 
  # Reorder columns for better readability and organization
  relocate(id_article, response_variable, silvo_mean, control_mean, yi, vi)

###################################################################
# Summarize effect sizes and variances per study and response variable
summary_meta_es <- meta_data_es %>%
  group_by(id_article, response_variable) %>%
  summarize(
    mean_yi = mean(yi, na.rm = TRUE),  # Average effect size
    mean_vi = mean(vi, na.rm = TRUE),  # Average variance
    n = n()  # Number of observations
  ) %>%
  dplyr::ungroup()
###################################################################
# Join the summarized data back to the original data
meta_data_es <- meta_data_es %>%
  left_join(summary_meta_es, by = c("id_article", "response_variable"))

###################################################################
# View the computed effect sizes and variances
meta_data_es %>% glimpse()
```

##############################################################################
Building generic function to derive ROM, to be used for each response variable 
##########################################################################################################################################


```{r}
# Define a generic function to compute effect sizes and variances
compute_effect_sizes <- function(data, response_var) {
  # Filter data for the specified response variable
  filtered_data <- data %>%
    filter(response_variable == response_var)
  
  # Ensure standard deviations are valid for escalc
  filtered_data <- filtered_data %>%
    # Ensure standard deviations are valid for escalc
    filter(silvo_se > 0 & control_se > 0) 
    # Ensure means are greater than 0 to avoid infinite values
    #filter(silvo_mean > 0 & control_mean > 0) 
  
  # Use the escalc function to compute effect sizes and variances
  effect_sizes <- metafor::escalc(
    # "ROM" specifies the log-transformed ratio of means as the effect size measure
    measure = "ROM",  
    # Mean of the silvoarable treatment
    m1i = silvo_mean,  
    # Mean of the control treatment
    m2i = control_mean,  
    # Standard error of the silvoarable treatment
    sd1i = silvo_se,  
    # Standard error of the control treatment
    sd2i = control_se,  
    # Sample size of the silvoarable treatment
    n1i = silvo_n,  
    # Sample size of the control treatment
    n2i = control_n,  
     # Data frame containing the relevant columns
    data = filtered_data, 
    # Create a label combining id_article and response_variable
    slab = paste(id_article, response_variable, sep = ", ")  
  ) %>% 
  # Reorder columns for better readability and organization
  relocate(id_article, response_variable, silvo_mean, control_mean, yi, vi)
  
  # Summarize effect sizes and variances per study and response variable
  summarized_effect_sizes <- effect_sizes %>%
    group_by(id_article, response_variable) %>%
    summarize(
      mean_yi = mean(yi, na.rm = TRUE),  # Average effect size
      mean_vi = mean(vi, na.rm = TRUE),  # Average variance
      n = n()  # Number of observations
    ) %>%
    ungroup()
  
  # Join summarized data back to the original data
  combined_data <- effect_sizes %>%
    left_join(summarized_effect_sizes, by = c("id_article", "response_variable"))
  
  # Return the combined data with summarized effect sizes and variances
  return(combined_data)
}
```

```{r}
# Define the list of response variables
response_variables <- c(
  "Biodiversity",
  "Crop yield",
  "Greenhouse gas emission",
  "Pest and Disease",
  "Product quality",
  "Soil quality",
  "Soil water content",
  "Water quality"
)

# Initialize an empty list to store results
results <- list()

# Loop through each response variable and compute effect sizes
for (response_var in response_variables) {
  cat("Computing effect sizes for:", response_var, "\n")
  results[[response_var]] <- compute_effect_sizes(meta_data, response_var)
}

# Combine all results into a single data frame for easier analysis and visualization
combined_results <- bind_rows(results)
# Warnings: Some 'yi' and/or 'vi' values equal to +-Inf. Recoded to NAs --> happens when either control_mean or solvi_mean is 0
# Computing effect sizes for: Biodiversity 
# Advarsel: Some 'yi' and/or 'vi' values equal to +-Inf. Recoded to NAs.
# Computing effect sizes for: Crop yield 
# Computing effect sizes for: Greenhouse gas emission 
# Computing effect sizes for: Pest and Disease 
# Advarsel: Some 'yi' and/or 'vi' values equal to +-Inf. Recoded to NAs.
# Computing effect sizes for: Product quality 
# Computing effect sizes for: Soil quality 
# Computing effect sizes for: Soil water content 
# Computing effect sizes for: Water quality 
# Advarsel: Some 'yi' and/or 'vi' values equal to +-Inf. Recoded to NAs.

combined_results %>% glimpse()
```

```{r}
# Access the computed effect sizes for a specific response variable
# Convert results for a specific response variable to tibble and view
biodiversity_results <- as_tibble(results[["Biodiversity"]]) %>% 
  # Reorder columns for better readability and organization
  relocate(id_article, response_variable, silvo_mean, control_mean, mean_yi, mean_vi)

biodiversity_results
```



################################
B) Sensitivity Analyses to assess the potential impact of missing outcome data.

```{r}

```


#############
# STEP 4
##########################################################################################################################################
PERFORMING META-ANALYSIS USING THE GENERAL FUNCTION rma IN metafor
##########################################################################################################################################

Linear (Mixed-Effects) Models

```{r}
# Remove duplicates to keep one row per combination of id_article and response_variable
biodiversity_results_forrest <- biodiversity_results %>%
  distinct(id_article, response_variable, 
           .keep_all = TRUE)
```

```{r}
# Biodiversity

res_meta_biodiversity <- rma(mean_yi, mean_vi, 
                             data = biodiversity_results_forrest) 


```
```{r}
# Create a forest plot for the meta-analysis of Biodiversity
forest(res_meta_biodiversity, 
       xlab = "Log Response Ratio (lnRR)",  # Label for the x-axis
       slab = biodiversity_results_forrest$id_article,  # Study labels
       main = "Forest Plot for Biodiversity",  # Title of the plot
       cex = 0.8,  # Size of the text
       cex.lab = 1.2)  # Size of the axis labels

# Create a funnel plot for the meta-analysis of Biodiversity
funnel(res_meta_biodiversity, 
  main = "Funnel Plot for Biodiversity")  # Title of the plot
```

```{r}
positive_definite_check <- is_positive_definite(V_matrix)
print(paste("V_matrix is positive definite:", positive_definite_check))

# Error
# The error encountered, "uendelig eller manglende værdier i 'x'" (infinite or missing values in 'x'), indicates that V_matrix contains infinite or missing values, which can cause issues with positive definiteness checks and model fitting. To address this, we need to clean V_matrix by handling these infinite or missing values.

# But I will do this on the original filtered_meta_data_ROM data!
```



```{r}
# Check for missing values in specified columns and remove those rows
filtered_meta_data_ROM_clean <- filtered_meta_data_ROM %>%
  filter(!is.na(silvo_mean) & !is.na(silvo_se) & !is.na(silvo_sd) &
         !is.na(control_mean) & !is.na(control_se) & !is.na(control_sd) &
         !is.na(yi) & !is.na(vi))
```

Proportion of excluded studies:

```{r}
filtered_meta_data_ROM_clean vs filtered_meta_data_ROM
```



##########################################################################
UPDATING THE V_matrix
##########################################################################

```{r}
# Create a new V_matrix corresponding to the updated filtered_meta_data_ROM aka. filtered_meta_data_ROM_clean
# Assuming the original V_matrix has row and column names that match id_obs or another unique identifier
# of filtered_meta_data_ROM

# Create a list to store variance-covariance matrices for each study
V_list <- list()

# Loop through each unique study ID in the dataset
for (study in unique(filtered_meta_data_ROM_clean$id_article)) {
  # Subset the data for the current study
  study_data <- filtered_meta_data_ROM_clean[filtered_meta_data_ROM_clean$id_article == study, ]
  
  # Check if the current study has more than one outcome
  if (nrow(study_data) > 1) {
    # Create a diagonal matrix with the variances (vi) of the outcomes
    V <- diag(study_data$vi)
    
    # Assume a constant correlation of 0.5 between outcomes within the same study
    corr <- 0.5
    
    # Loop through the rows of the matrix to set the off-diagonal elements
    for (i in 1:nrow(V)) {
      for (j in 1:nrow(V)) {
        # Set the off-diagonal elements to the product of the correlation and the square root of the product of the corresponding variances
        if (i != j) {
          V[i, j] <- corr * sqrt(V[i, i] * V[j, j])
        }
      }
    }
    # Add the variance-covariance matrix to the list for the current study
    V_list[[as.character(study)]] <- V
  } else {
    # If there is only one outcome, the variance is just the variance of the single outcome
    V_list[[as.character(study)]] <- study_data$vi
  }
}

# Combine the matrices into a block-diagonal matrix
V_matrix_clean <- bldiag(V_list)
```

```{r}
# Ensure the V_matrix_clean is still valid and clean
positive_definite_check <- is_positive_definite(V_matrix_clean)
print(paste("V_matrix_clean is positive definite:", positive_definite_check))
```



